---
title: "Introduction"
author: "Connor Quiroz, Juliano Palacios-Abrantes, and  Jessica A. Gephart"
date: "2026-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



# Read in data

```{r}
# Read in aquatic animal protein reliance data
aquatic_reliance <- read_csv("../data/miscellaneous/fao_aquatic_reliance_source.csv")

# Read in consumption data
consumption_full <- read_parquet("../data/miscellaneous/consumption_full.parquet")
```
# When did aquaculture surpass capture fisheries?

```{r}
# Calculate the aggregate protein reliance by habitat and method
agg_rel_method <- aquatic_reliance %>%
  group_by(
  year,
  data_source,
  method,
  food_group
  ) %>%
  # Calculate annual global aquatic animal protein consumption
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(perc_animal_protein = 100 *
           (food_group_total_protein_t / sum(food_group_total_protein_t))) %>%
  filter(food_group == "aquatic")

agg_rel_method %>%
  filter(method != "unknown")
```


# What percent of global-animal source protein consumption do marine caught aquatic foods make up?

```{r}
# Calculate the aggregate protein reliance by habitat and method
agg_rel_habitat_method <- aquatic_reliance %>%
  group_by(
  year,
  data_source,
  method,
  habitat,
  food_group
  ) %>%
  # Calculate annual global aquatic animal protein consumption
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(perc_animal_protein = 100 *
           (food_group_total_protein_t / sum(food_group_total_protein_t))) %>%
  filter(food_group == "aquatic") %>%
  # Remove overlapping FBS years
  filter(year == 2023)

agg_rel_habitat_method %>%
  mutate(perc_animal_protein = round(perc_animal_protein, 1)) %>%
  arrange(-perc_animal_protein)
```

# What proportion of marine captures has increased 
```{r}
# How many countries in each region source a majority of their marine capture foods
# from foreign sources?
 consumption_full %>%
  filter(method == "capture",
         habitat == "marine") %>%
  group_by(year, consumer_iso3c, consumption_source) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(consumption_source == "foreign") %>%
  add_region("consumer_iso3c", "region") %>%
  group_by(region) %>%
  count(prop_consumption > 0.5)

# What are some example countries with a majority in marine capture consumption?
 consumption_full %>%
  filter(method == "capture",
         habitat == "marine") %>%
  group_by(year, consumer_iso3c, consumption_source) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(consumption_source == "foreign") %>%
  add_region("consumer_iso3c", "region") %>%
  filter(region %in% c("North America", "Europe", "Africa")) %>%
   select(-consumption_live_t) %>%
   View()
```

North America, Europe, and Africa have particularly increased their foreign consumption over the years, exceeding domestic consumption in some countries like Canada, Seychelles, and Austria.
