---
title: "00 — Preprocess Data"
author: "Connor Quiroz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
```

# Overview

This notebook defines four functions that build the exposure, sensitivity, and
adaptive capacity (E/S/AC) components used to calculate climate risk. Each
function handles one job:

| Function | Runs once? | Purpose |
|---|---|---|
| `preprocess_climate_data()` | ✅ Yes | Read DBEM files → interpolate missing taxa → write parquet |
| `build_exposure()` | Per portfolio | Join climate projections to consumption; compute % stock change |
| `build_sensitivity()` | Per portfolio | Compute aquatic animal reliance by portfolio type |
| `build_adaptive_capacity()` | Per portfolio | Load and join all AC indicators |

`build_e_s_ac_data()` orchestrates the three per-portfolio steps and writes the
final joined dataset to `../output/`.

---

# Step 1 — Climate data preprocessing (run once)

Reads all DBEM species files, joins species names, converts EEZ strings to
ISO3C codes, and writes a single joined CSV. Because this step is
portfolio-agnostic it only needs to run once per session.

```{r preprocess-climate-data}
preprocess_climate_data <- function(dir) {

  file_names <- list.files(
    paste0(dir, "exposure/dbem_delta_mcp/"),
    pattern = "\\.csv",
    full.names = TRUE
  )

  species_names <- fread(paste0(dir, "exposure/dbem_spp_list.csv"))

  df <- map_dfr(file_names, function(f) {
    df_i <- fread(f)
    if (nrow(df_i) == 0) return(NULL)

    df_i %>%
      left_join(species_names, by = "taxon_key") %>%
      mutate(taxon_name = str_to_lower(taxon_name)) %>%
      select(taxon_name, eez_name, ssp, mean_mcp_delta_2030) %>%
      pivot_wider(names_from = ssp, values_from = mean_mcp_delta_2030, values_fn = mean)
  })

  # Recode EEZ territory strings to ISO3C — territories are assigned to their
  # sovereign nation
  df <- df %>%
    mutate(eez_name = case_when(
      eez_name == "Tokelau (New Zealand)"              ~ "NZL",
      eez_name == "Svalbard Isl. (Norway)"             ~ "NOR",
      eez_name == "Chagos Archipelago (UK)"            ~ "GBR",
      eez_name == "Christmas Isl. (Australia)"         ~ "AUS",
      eez_name == "High seas"                          ~ "NEI",
      eez_name == "Iran (Sea of Oman)"                 ~ "IRN",
      eez_name == "Mayotte (France)"                   ~ "FRA",
      eez_name == "Mozambique Channel Isl. (France)"   ~ "FRA",
      eez_name == "New Caledonia (France)"             ~ "FRA",
      eez_name == "Norfolk Isl. (Australia)"           ~ "AUS",
      eez_name == "Réunion (France)"                   ~ "FRA",
      eez_name == "Saudi Arabia (Persian Gulf)"        ~ "SAU",
      eez_name == "Wallis & Futuna Isl. (France)"      ~ "FRA",
      eez_name == "Yemen (Arabian Sea)"                ~ "YEM",
      TRUE ~ countrycode(eez_name, origin = "country.name", destination = "iso3c")
    )) %>%
    rename(eez_iso3c = eez_name, sciname = taxon_name)

  write_csv(df, "../output/future_climate_joined.csv")
  message("✅ future_climate_joined.csv written to ../output/")

  invisible(df)
}
```

---

# Step 2 — Exposure (per portfolio)

Joins climate projections to 2019 consumption data, fills missing species
values up through the taxonomic hierarchy, then computes each country's
weighted % change in available stock under SSP1-2.6 and SSP5-8.5.

**Portfolio differences:**

- `"foreign"` — keeps only rows where `consumer_iso3c != producer_iso3c` and
  applies a three-way case_when for the catch-change multiplier (handles
  positive, negative, and zero separately).
- `"total"` — retains the full consumption portfolio and uses a simpler
  `1 + (x / 100)` multiplier.

```{r build-exposure}
# Helper: classify each sciname to its taxonomic level and extract species epithet
build_taxa_levels <- function(sciname_df) {
  sciname_df %>%
    mutate(
      taxa_level = case_when(
        sciname == kingdom   ~ "kingdom",
        sciname == phylum    ~ "phylum",
        sciname == superclass ~ "superclass",
        sciname == class     ~ "class",
        sciname == order     ~ "order",
        sciname == family    ~ "family",
        sciname == subfamily ~ "subfamily",
        sciname == genus     ~ "genus",
        str_detect(sciname, " ") ~ "species"
      ),
      species = if_else(str_count(sciname, " ") == 1, word(sciname, 2), NA_character_)
    ) %>%
    relocate(species, .before = genus) %>%
    select(-common_name, -isscaap)
}

# Helper: for each higher taxonomic level, compute mean SSP values per EEZ
# so that species with missing projections can borrow from relatives
build_taxa_lookup <- function(df) {
  tax_levels <- c("genus", "subfamily", "family", "order", "class", "phylum")

  unique_climate <- df %>%
    select(eez_iso3c, sciname_hs_modified, ssp126, ssp585,
           genus, subfamily, family, order, class, phylum, kingdom, superclass) %>%
    distinct()

  map_dfr(tax_levels, function(level) {
    unique_climate %>%
      filter(!is.na(ssp126) | !is.na(ssp585)) %>%
      distinct(eez_iso3c, sciname_hs_modified, .keep_all = TRUE) %>%
      group_by(eez_iso3c, !!sym(level)) %>%
      summarise(
        avg_ssp126 = mean(ssp126, na.rm = TRUE),
        avg_ssp585 = mean(ssp585, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      rename(taxa_name = !!sym(level)) %>%
      mutate(taxa_level = level)
  })
}

# Helper: walk up the taxonomic ladder and fill NA SSP values from the nearest
# available ancestor within the same EEZ
fill_with_higher_taxa <- function(df, lookup_table) {
  tax_levels <- c("genus", "subfamily", "family", "order", "class", "phylum")

  for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level)

    df <- df %>%
      select(-taxa_level) %>%
      left_join(
        lookup_table %>% filter(taxa_level == level),
        by = c("eez_iso3c", join_by_map)
      ) %>%
      mutate(
        ssp126 = coalesce(ssp126, avg_ssp126),
        ssp585 = coalesce(ssp585, avg_ssp585)
      ) %>%
      select(-avg_ssp126, -avg_ssp585)
  }

  df %>% select(-taxa_level)
}

build_exposure <- function(dir, portfolio = c("foreign", "total")) {
  portfolio <- match.arg(portfolio)

  consumption <- read_parquet(paste0(dir, "example_consumption_eez_2024_12_06.parquet")) %>%
    mutate(sciname_hs_modified = coalesce(sciname_hs_modified, sciname))

  sciname_taxa <- read_csv(paste0(dir, "sciname.csv"), show_col_types = FALSE) %>%
    build_taxa_levels()

  climate_by_spp_eez <- read_csv("../output/future_climate_joined.csv", show_col_types = FALSE) %>%
    drop_na() %>%
    group_by(eez_iso3c, sciname) %>%
    summarise(across(c(ssp126, ssp585), ~mean(.x, na.rm = TRUE)), .groups = "drop")

  consumption_2019 <- consumption %>%
    filter(year == 2019) %>%
    left_join(sciname_taxa, by = c("sciname_hs_modified" = "sciname")) %>%
    left_join(climate_by_spp_eez, by = c("eez_iso3c", "sciname_hs_modified" = "sciname"))

  interpolated <- fill_with_higher_taxa(consumption_2019, build_taxa_lookup(consumption_2019)) %>%
    mutate(
      taxa_level = case_when(
        sciname_hs_modified == kingdom    ~ "kingdom",
        sciname_hs_modified == phylum     ~ "phylum",
        sciname_hs_modified == superclass ~ "superclass",
        sciname_hs_modified == class      ~ "class",
        sciname_hs_modified == order      ~ "order",
        sciname_hs_modified == family     ~ "family",
        sciname_hs_modified == subfamily  ~ "subfamily",
        sciname_hs_modified == genus      ~ "genus",
        str_detect(sciname_hs_modified, " ") ~ "species"
      ),
      species = if_else(
        str_count(sciname_hs_modified, " ") == 1,
        word(sciname_hs_modified, 2),
        NA_character_
      )
    )

  write_parquet(interpolated, "../output/interpolated_future_data.parquet")

  future_long <- interpolated %>%
    pivot_longer(c(ssp126, ssp585), names_to = "scenario", values_to = "future_change_catch")

  # Portfolio-specific row filter + catch-change multiplier
  # Note: the two portfolios intentionally use different multiplier formulas —
  # the foreign portfolio handles positive/negative/zero separately to preserve
  # directionality; the total portfolio uses a single linear conversion.
  stock_pre <- if (portfolio == "foreign") {
    future_long %>%
      filter(
        consumer_iso3c != producer_iso3c,
        !is.na(live_weight_t),
        !is.na(future_change_catch)
      ) %>%
      mutate(future_change_catch = case_when(
        future_change_catch > 0 ~ (future_change_catch / 100) + 1,
        future_change_catch < 0 ~ 1 - ((future_change_catch / 100) * -1),
        future_change_catch == 0 ~ 1
      ))
  } else {
    future_long %>%
      filter(!is.na(live_weight_t), !is.na(future_change_catch)) %>%
      mutate(future_change_catch = 1 + (future_change_catch / 100))
  }

  stock_pre %>%
    mutate(change_in_stock = live_weight_t * future_change_catch) %>%
    group_by(consumer_iso3c, scenario) %>%
    summarise(
      change_in_stock = sum(change_in_stock),
      live_weight_t   = sum(live_weight_t),
      .groups = "drop"
    ) %>%
    mutate(pct_change = 100 * ((change_in_stock - live_weight_t) / live_weight_t))
}
```

---

# Step 3 — Sensitivity (per portfolio)

Computes aquatic animal reliance as a share of total animal protein supply.

- `"foreign"` — sums across all consumption sources (domestic + foreign) to
  get total reliance.
- `"total"` — retains only the `foreign` consumption source row (no
  aggregation needed).

```{r build-sensitivity}
build_sensitivity <- function(dir, portfolio = c("foreign", "total")) {
  portfolio <- match.arg(portfolio)

  supply_importance <- read_csv(
    paste0(dir, "sensitivity/fao_aquatic_reliance_source.csv"),
    show_col_types = FALSE
  )

  base <- supply_importance %>%
    filter(year == 2019, habitat == "marine", method == "capture", food_group == "aquatic")

  if (portfolio == "foreign") {
    base %>%
      group_by(iso3c) %>%
      summarise(aa_reliance_pct = sum(prop_animal_protein), .groups = "drop")
  } else {
    base %>%
      filter(consumption_source == "foreign") %>%
      select(iso3c, aa_reliance_pct = prop_animal_protein)
  }
}
```

---

# Step 4 — Adaptive capacity (portfolio-agnostic)

Loads all ten AC indicators across the four IPCC dimensions (Assets,
Flexibility, Learning, Social Organization), standardises column names, and
returns a single joined data frame keyed on `consumer_iso3c`.

The `consuming_countries` argument (a vector of ISO3C codes) is passed in
explicitly so this function has no hidden dependency on objects created in
earlier steps.

```{r build-adaptive-capacity}
build_adaptive_capacity <- function(dir, consuming_countries) {

  cc_df <- tibble(consumer_iso3c = consuming_countries, artis = 1)

  # country_coverage() returns the ARTIS countries that are missing from a
  # given AC dataset — useful for diagnosing coverage gaps
  country_coverage <- function(data, join_by) {
    full_join(cc_df, data %>% mutate(covered = 1), by = c("consumer_iso3c" = join_by)) %>%
      filter(is.na(covered), artis == 1) %>%
      pull(consumer_iso3c)
  }

  # --- Assets ---
  sanitation <- read_csv(paste0(dir, "adaptive capacity/assets/sanitation.csv"), show_col_types = FALSE) %>%
    filter(Indicator_Code == "SH_STA_BASS_ZS", Time_Period == 2019) %>%
    rename(sanitation = Value, consumer_iso3c = Geography_Code) %>%
    select(consumer_iso3c, sanitation)

  gdp <- read_csv(paste0(dir, "adaptive capacity/assets/gdp.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, gdp = `2019`) %>%
    filter(!is.na(gdp))

  trade_gdp <- read_csv(paste0(dir, "adaptive capacity/assets/trade_gdp.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Code` == "NE.TRD.GNFS.ZS") %>%
    select(consumer_iso3c = `Country Code`, gdp_trade = `2019`) %>%
    filter(!is.na(gdp_trade))

  # --- Flexibility ---
  life_expectancy <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/life_expectancy_at_birth.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator Code` == "SP.DYN.LE00.IN") %>%
    select(consumer_iso3c = `Country Code`, life_expectancy = `2019`) %>%
    filter(!is.na(life_expectancy))

  supermarkets <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/supermarkets_per_100000.csv"),
    show_col_types = FALSE
  ) %>%
    mutate(consumer_iso3c = countrycode(Region, origin = "country.name", destination = "iso3c")) %>%
    select(consumer_iso3c, supermarkets = `2019`) %>%
    filter(!is.na(supermarkets))

  total_labor   <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_labor_force.csv"), show_col_types = FALSE)
  total_pop     <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_population.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, total_pop_2019 = `2019`)

  prop_labor <- total_labor %>%
    left_join(total_pop, by = c("Country Code" = "consumer_iso3c")) %>%
    select(consumer_iso3c = `Country Code`, labor_2019 = `2019`, total_pop_2019) %>%
    mutate(prop_labor = labor_2019 / total_pop_2019) %>%
    filter(!is.na(prop_labor)) %>%
    select(consumer_iso3c, prop_labor)

  # --- Learning ---
  hci <- read_csv(paste0(dir, "adaptive capacity/learning/HCIData.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Name` == "Human Capital Index (HCI) (scale 0-1)") %>%
    select(consumer_iso3c = `Country Code`, hci = `2018`) %>%
    filter(!is.na(hci))

  # --- Social organization ---
  gov_effectiveness <- read_csv(
    paste0(dir, "adaptive capacity/social organization/government_effectiveness_percentile.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator ID` == "WB.WWGI.GE.PER.RNK") %>%
    select(consumer_iso3c = `Economy ISO3`, gov_effectiveness = `2019`) %>%
    filter(!is.na(gov_effectiveness))

  fsc <- read_csv(
    paste0(dir, "adaptive capacity/social organization/food-systems-dashboard-2025-03-04.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Start Year` == 2019 | `End Year` == 2019) %>%
    rename(fsc = Value, consumer_iso3c = ISO3) %>%
    select(consumer_iso3c, fsc)

  rol <- read_csv(
    paste0(dir, "adaptive capacity/social organization/WB-WWGI.csv"),
    show_col_types = FALSE
  ) %>%
    filter(Indicator == "Rule of Law: Percentile Rank") %>%
    select(consumer_iso3c = `Economy ISO3`, rol = `2019`)

  # Join all AC indicators
  gdp %>%
    left_join(trade_gdp,        by = "consumer_iso3c") %>%
    left_join(sanitation,       by = "consumer_iso3c") %>%
    left_join(supermarkets,     by = "consumer_iso3c") %>%
    left_join(life_expectancy,  by = "consumer_iso3c") %>%
    left_join(prop_labor,       by = "consumer_iso3c") %>%
    left_join(hci,              by = "consumer_iso3c") %>%
    left_join(gov_effectiveness, by = "consumer_iso3c") %>%
    left_join(fsc,              by = "consumer_iso3c") %>%
    left_join(rol,              by = "consumer_iso3c")
}
```

---

# Orchestrator — `build_e_s_ac_data()`

Calls `build_exposure()`, `build_sensitivity()`, and
`build_adaptive_capacity()` for a given portfolio type, joins the results, and
writes the output parquet.

```{r build-e-s-ac-data}
build_e_s_ac_data <- function(portfolio = c("foreign", "total"), dir) {
  portfolio <- match.arg(portfolio)
  cat(sprintf("\n--- Building E/S/AC data: %s portfolio ---\n", portfolio))

  exposure     <- build_exposure(dir, portfolio)
  sensitivity  <- build_sensitivity(dir, portfolio)

  consuming_countries <- exposure %>% distinct(consumer_iso3c) %>% pull()
  adaptive_cap <- build_adaptive_capacity(dir, consuming_countries)

  e_s_ac <- adaptive_cap %>%
    right_join(sensitivity,  by = c("consumer_iso3c" = "iso3c")) %>%
    left_join(exposure,      by = "consumer_iso3c")

  out_path <- sprintf("../output/e_s_ac_data_%s.parquet", portfolio)
  write_parquet(e_s_ac, out_path)
  cat(sprintf("✅ Saved %s\n", out_path))

  invisible(e_s_ac)
}
```
