---
title: "00 — Preprocess Data"
author: "Connor Quiroz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(rfishbase)
```

FIXIT: Warning: There were 250 warnings in `mutate()` check out where those 250 warnings are coming from and see whether that is okay or not.

# Overview

This notebook defines four functions that build the exposure, sensitivity, and
adaptive capacity (E/S/AC) components used to calculate climate risk. Each
function handles one job:

| Function | Runs once? | Purpose |
|---|---|---|
| `preprocess_climate_data()` | ✅ Yes | Read DBEM files → interpolate missing taxa → write parquet |
| `build_exposure()` | Per portfolio | Join climate projections to consumption; compute % stock change |
| `build_sensitivity()` | Per portfolio | Compute aquatic animal reliance by portfolio type |
| `build_adaptive_capacity()` | Per portfolio | Load and join all AC indicators |

`build_e_s_ac_data()` orchestrates the three per-portfolio steps and writes the
final joined dataset to `../output/`.

---

# Step 1 — Climate data preprocessing (run once)

Reads all DBEM species files, joins species names, converts EEZ strings to
ISO3C codes, and writes a single joined CSV. Because this step is
portfolio-agnostic it only needs to run once per session.

```{r preprocess-climate-data}
preprocess_climate_data <- function(dir) {

  file_names <- list.files(
    paste0(dir, "exposure/dbem_delta_mcp/"),
    pattern = "\\.csv",
    full.names = TRUE
  )

  species_names <- fread(paste0(dir, "exposure/dbem_spp_list.csv"))


  df <- map_dfr(file_names, function(f) {
    df_i <- fread(f)
    if (nrow(df_i) == 0) return(NULL)

    df_i %>%
      left_join(species_names, by = "taxon_key") %>%
      mutate(taxon_name = str_to_lower(taxon_name),
             file_found_in = as.numeric(sub(".*_(\\d+)\\.csv$", "\\1", f))) %>% # variable that stores which file name X species was found in (will be used to ensure that species are found in their correct file names and not multiple files)
      select(taxon_name, eez_name, ssp, mean_mcp_delta_2030, file_found_in) %>%
      pivot_wider(names_from = ssp, values_from = mean_mcp_delta_2030, values_fn = mean)
  })
  
  # Save files to .RDS (for quality checks) FIXIT - maybe delete later
  saveRDS(file_names, "../data/exposure/utils/file_names.rds")
  saveRDS(species_names, "../data/exposure/utils/species_names.rds")

  # Recode EEZ territory strings to ISO3C — territories are assigned to their
  # sovereign nation
  df <- df %>%
    mutate(eez_name = case_when(
      eez_name == "Tokelau (New Zealand)"              ~ "NZL",
      eez_name == "Svalbard Isl. (Norway)"             ~ "NOR",
      eez_name == "Chagos Archipelago (UK)"            ~ "GBR",
      eez_name == "Christmas Isl. (Australia)"         ~ "AUS",
      eez_name == "High seas"                          ~ "NEI",
      eez_name == "Iran (Sea of Oman)"                 ~ "IRN",
      eez_name == "Mayotte (France)"                   ~ "FRA",
      eez_name == "Mozambique Channel Isl. (France)"   ~ "FRA",
      eez_name == "New Caledonia (France)"             ~ "FRA",
      eez_name == "Norfolk Isl. (Australia)"           ~ "AUS",
      eez_name == "Réunion (France)"                   ~ "FRA",
      eez_name == "Saudi Arabia (Persian Gulf)"        ~ "SAU",
      eez_name == "Wallis & Futuna Isl. (France)"      ~ "FRA",
      eez_name == "Yemen (Arabian Sea)"                ~ "YEM",
      TRUE ~ countrycode(eez_name, origin = "country.name", destination = "iso3c")
    )) %>%
    rename(eez_iso3c = eez_name, sciname = taxon_name)

  write_csv(df, "../output/future_catch_data_joined/future_catch_joined.csv")
  message("✅ future_catch_joined.csv written to ../output/")

  invisible(df)
}

# FIXIT: find some way to get soverign country out of string
# dplyr::mutate(!!country_col_name := stringr::str_remove(.data[[country_col_name]], "\\(.+\\)$")) 
```

---

# Step 2 — Exposure (per portfolio)

Joins climate projections to 2019 consumption data, fills missing species
values up through the taxonomic hierarchy, then computes each country's
weighted % change in available stock under SSP1-2.6 and SSP5-8.5.

**Portfolio differences:**

- `"foreign"` — keeps only rows where `consumer_iso3c != producer_iso3c` and
  applies a three-way case_when for the catch-change multiplier (handles
  positive, negative, and zero separately).
- `"total"` — retains the full consumption portfolio and uses a simpler
  `1 + (x / 100)` multiplier.

```{r build-exposure}
# Helper: classify each sciname to its taxonomic level and extract species epithet
build_taxa_levels <- function(sciname_df) {
  sciname_df %>%
    mutate(
      taxa_level = case_when(
        sciname == kingdom   ~ "kingdom",
        sciname == phylum    ~ "phylum",
        sciname == superclass ~ "superclass",
        sciname == class     ~ "class",
        sciname == order     ~ "order",
        sciname == family    ~ "family",
        sciname == subfamily ~ "subfamily",
        sciname == genus     ~ "genus",
        str_detect(sciname, " ") ~ "species"
      ),
      species = if_else(str_count(sciname, " ") == 1, word(sciname, 2), NA_character_)
    ) %>%
    relocate(species, .before = genus) %>%
    select(-common_name, -isscaap)
}

# Helper: for each higher taxonomic level, compute mean SSP values per EEZ
# so that species with missing projections can borrow from relatives
generate_mcp_averages <- function(df,
                                  method = c("eez", "realm", "region", "global"),
                                  averaging = c("unweighted", "weighted"),
                                  ecoregion,
                                  num_species) {
  
  method <- match.arg(method)
  averaging <- match.arg(averaging)
  tax_levels <- c("genus", "family", "order", "class", "phylum", "kingdom")
  
  x <- map_dfr(tax_levels, function(level) {
    
    # Based on function input, group by function method input
    group_cols <- switch(
      method,
      eez   = c("eez_iso3c", level),
      realm = c("realm", level),
      region = c("region", level),
      global = c(level)
    )
    
    if (averaging == "weighted") { # Compute weighted avearages
      
      df2 <- df %>%
        filter(!is.na(ssp126) | !is.na(ssp585)) %>%
        group_by(sciname, eez_iso3c) %>%
        summarize(live_weight_t = sum(live_weight_t)) %>%
        ungroup() %>%
        left_join(
          ecoregion %>% select(iso3c, realm),
          by = c("eez_iso3c" = "iso3c")
        ) %>%
        add_region("eez_iso3c", "region")
      
      # Join ssp estimates back onto df dataset
      df2 <- df2 %>%
        left_join(
          df %>%
            add_region("eez_iso3c", "region") %>%
            left_join(ecoregion, by = c("eez_iso3c" = "iso3c")) %>%
            filter(!is.na(ssp126)) %>%
            distinct(sciname, across(all_of(c("eez_iso3c", group_cols[length(group_cols)]))), ssp126, ssp585),
          by = c("sciname", "eez_iso3c")
        )
      
      # Allow averages to only include at least X species (so that one species doesn't dominate).
      # Number of minimum species used for gapfilling depends on user input
      df2 <- df2 %>%
        group_by(across(all_of(c(group_cols[1], level)))) %>%
        mutate(n = n_distinct(sciname)) %>%   # Fixit: Currently set to minimum number of species. Do we want this to be minimum number of eez / iso3c pairs? That would be much easier to satisfy.
        ungroup() %>%
        filter(n >= num_species)

      
      out <- df2 %>%
        group_by(across(all_of(group_cols))) %>%
        mutate(total_live_weight_group = sum(live_weight_t)) %>%
        
        summarize(
          # Weighted averages
          
          !!paste0(method, "_avg_ssp126") := weighted.mean(ssp126, live_weight_t, na.rm = TRUE),
          !!paste0(method, "_avg_ssp585") := sum(ssp585 * (live_weight_t / total_live_weight_group)),

          # Weighted Coefficients of variation
          !!paste0(method, "_cv_ssp126") := {
            
            wm <- weighted.mean(ssp126, live_weight_t, na.rm = TRUE)

            wvar <- sum(live_weight_t * (ssp126 - wm)^2, na.rm = TRUE) /
              sum(live_weight_t, na.rm = TRUE)

            sqrt(wvar) / wm
            
          },
          
          !!paste0(method, "_cv_ssp585") := {
            
            wm <- weighted.mean(ssp585, live_weight_t, na.rm = TRUE)

            wvar <- sum(live_weight_t * (ssp585 - wm)^2, na.rm = TRUE) /
              sum(live_weight_t, na.rm = TRUE)

            sqrt(wvar) / wm
            
          }
        ) %>%
        ungroup()
      
    } else { # Compute unweighted averages
      
      df2 <- df %>%
      distinct() %>%
      left_join(
        ecoregion %>% select(iso3c, realm),
        by = c("eez_iso3c" = "iso3c")
      ) %>%
      filter(!is.na(ssp126) | !is.na(ssp585)) %>%
      add_region("eez_iso3c", "region") %>%
      distinct(eez_iso3c, sciname, .keep_all = TRUE)
      
      
      # Allow averages to only include at least 2 species (so that one species doesn't dominate)
      df2 <- df2 %>%
        group_by(across(all_of(c(group_cols[1], level)))) %>%
        mutate(n = n()) %>%   # number of rows in each pair
        ungroup() %>%
        filter(n >= num_species)
      
      out <- df2 %>%
        group_by(across(all_of(group_cols))) %>%
        summarise(
          !!paste0(method, "_avg_ssp126") := mean(ssp126, na.rm = TRUE),
          !!paste0(method, "_avg_ssp585") := mean(ssp585, na.rm = TRUE),
          
          !!paste0(method, "_cv_ssp126") :=
            sd(ssp126, na.rm = TRUE) / mean(ssp126, na.rm = TRUE),
          
          !!paste0(method, "_cv_ssp585") :=
            sd(ssp585, na.rm = TRUE) / mean(ssp585, na.rm = TRUE)
        ) %>%
        ungroup()
      
    }
    
    out %>%
      rename(taxa_name = all_of(level)) %>%
      mutate(taxa_level = level)
    
  })
  
  return(x)
  
}

# Helper: compile averages across all gap filling combinations into a list
# that will be used to gap fill NA MCP projections in consumption data
build_taxa_lookup <- function(df, ecoregion, num_species) {
  tax_levels <- c("genus", "family", "order", "class", "phylum", "kingdom")

  # For each taxa in an EEZ, compute the average ssp126 / 585 value for all species within in the taxa
  
  # ---------------------------------------------------------------------------
  # EEZ-level averages
  # ---------------------------------------------------------------------------
  
  eez_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "eez",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  eez_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "eez",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)

  # ---------------------------------------------------------------------------
  # Realm-level averages
  # ---------------------------------------------------------------------------
  
  realm_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "realm",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  realm_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "realm",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  # ---------------------------------------------------------------------------
  # Region-level averages
  # ---------------------------------------------------------------------------
  region_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "region",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  region_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "region",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  # ---------------------------------------------------------------------------
  # Global-level averages
  # ---------------------------------------------------------------------------
  
  global_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "global",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  global_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "global",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  return(list(
    eez_gapfills_unwgt = eez_gapfills_unwgt,
    eez_gapfills_wgt = eez_gapfills_wgt,
    region_gapfills_unwgt = region_gapfills_unwgt,
    region_gapfills_wgt = region_gapfills_wgt,
    realm_gapfills_unwgt = realm_gapfills_unwgt,
    realm_gapfills_wgt = realm_gapfills_wgt,
    global_gapfills_unwgt = global_gapfills_unwgt,
    global_gapfills_wgt = global_gapfills_wgt
  ))
      
}

# Helper: walk up the taxonomic ladder and fill NA SSP values from the nearest
# available ancestor within the same EEZ
fill_higher_taxa_helper <- function(df,
                                    gap_filling_method,
                                    averaging,
                                    ecoregion,
                                    lookup_table,
                                    level,
                                    join_by_map) {
  
  suffix <- if (averaging == "weighted") "wgt" else "unwgt"
  selector <- paste0(gap_filling_method,"_gapfills_",suffix)
  
  if (level == "genus") {
    
    if (gap_filling_method == "realm") {
      df <- df %>%
        left_join(ecoregion %>% select(iso3c, realm), by = c("eez_iso3c" = "iso3c"))
    } else if (gap_filling_method == "region") {
      df <- df %>%
        add_region("eez_iso3c", "region")
    }
    
  }
  
  # Based on function input, group by function method input
  join_cols <- switch(
    gap_filling_method,
    eez   = c("eez_iso3c", join_by_map),
    realm = c("realm", join_by_map),
    region = c("region", join_by_map),
    global = c(join_by_map)
  )
  
  df <- df %>%
    select(-taxa_level) %>%
    ## Join data to EEZ interpolated data
    left_join(
      lookup_table[[selector]] %>% filter(taxa_level == level),
      by = join_cols) %>%
    ## Conduct EEZ gapfills
    ### Only assign the gapfill method if there was actually a gapfill occurring
    ### Only assign the taxonomic level for which the gapfill was performed at
    ### if a gapfill occurred
    mutate(
      gapfill_level = case_when( ## Define at which taxa level gap filled data was gotten from
        is.na(ssp126) & !is.na(!!sym(paste0(gap_filling_method,"_avg_ssp126"))) ~ level, # Data coverage between ssp126/585 are the same, so we only need to check one variable
        TRUE                                              ~ gapfill_level
      ),
      gapfill_method = case_when(
        is.na(ssp126) & !is.na(!!sym(paste0(gap_filling_method,"_avg_ssp126"))) ~ paste0(gap_filling_method,"_gapfill"),
        TRUE                                              ~ gapfill_method
      ),
      ssp126 = coalesce(ssp126, !!sym(paste0(gap_filling_method,"_avg_ssp126"))),
      ssp585 = coalesce(ssp585, !!sym(paste0(gap_filling_method,"_avg_ssp585"))),
      cv_ssp126 = coalesce(cv_ssp126, !!sym(paste0(gap_filling_method,"_cv_ssp126"))),
      cv_ssp585 = coalesce(cv_ssp585, !!sym(paste0(gap_filling_method,"_cv_ssp585")))
    ) %>%
    select(-!!sym(paste0(gap_filling_method,"_avg_ssp126")),
           -!!sym(paste0(gap_filling_method,"_avg_ssp585")),
           -!!sym(paste0(gap_filling_method,"_cv_ssp126")),
           -!!sym(paste0(gap_filling_method,"_cv_ssp585")))
  
  return(df)
  
}

# Helper: conduct gap filling based on selected method
fill_with_higher_taxa <- function(df,
                                  gap_filling_method,
                                  averaging,
                                  ecoregion,
                                  num_species) {
  
  if(!file.exists(paste0("../data/exposure/lookup_tables/",num_species,"_species/lookup_table.rds"))) {
    
    # Create directory if doesn't already exist
    if (!dir.exists(paste0("../data/exposure/lookup_tables/",num_species,"_species/"))) {
      dir.create(paste0("../data/exposure/lookup_tables/",num_species,"_species/"))
    }
    
    # Build lookup_table
    lookup_table <- build_taxa_lookup(df = df, ecoregion = ecoregion, num_species = num_species)
    saveRDS(lookup_table, paste0("../data/exposure/lookup_tables/",num_species,"_species/lookup_table.rds")) # Save lookup table
     
  } else {
    
    lookup_table <- readRDS(paste0("../data/exposure/lookup_tables/",num_species,"_species/lookup_table.rds"))
    
  }
  
  cat(sprintf("✔️ Lookup table loaded in\n"))
  
  
  # Set taxonomic levels to iterate through
  tax_levels <- c("genus", "family", "order", "class", "phylum", "kingdom")

  # Create empty variables that will be filled as gap filling happens
  out_df <- df %>% mutate(gapfill_level = NA_character_,
                      gapfill_method = NA_character_,
                      cv_ssp126 = NA,
                      cv_ssp585 = NA)

  # ---------------------------------------------------------------------------
  # EEZ-level gap filling
  # ---------------------------------------------------------------------------
  if (gap_filling_method != "hierarchical" & gap_filling_method != "diagonal") {
   
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level)
      
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "eez",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    
    }
  
  # ---------------------------------------------------------------------------
  # Realm-level gap filling
  # ---------------------------------------------------------------------------
  
  if (gap_filling_method == "realm") {
    
    for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level) 
    
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "realm",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    }
    
  }
    
    
  # ---------------------------------------------------------------------------
  # Region-level gap filling
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "region") {
    
    for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level) 
    
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "region",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    }
    
  }
    
  # ---------------------------------------------------------------------------
  # Realm --> Region --> Global gapfilling 
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "thorough") {
    
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level) 
      
      # 1. Attempt realm-level gap fills
      out_df <- fill_higher_taxa_helper(out_df,
                                    gap_filling_method = "realm",
                                    averaging = averaging,
                                    lookup_table = lookup_table,
                                    ecoregion = ecoregion,
                                    level = level,
                                    join_by_map = join_by_map)
      
    }
    
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level) 
      
      # 2. Attempt region-level gap fills
      out_df <- fill_higher_taxa_helper(out_df,
                                    gap_filling_method = "region",
                                    averaging = averaging,
                                    lookup_table = lookup_table,
                                    ecoregion = ecoregion,
                                    level = level,
                                    join_by_map = join_by_map)
      
    }
    
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level) 
      
      # 2. Attempt region-level gap fills
      out_df <- fill_higher_taxa_helper(out_df,
                                    gap_filling_method = "global",
                                    averaging = averaging,
                                    lookup_table = lookup_table,
                                    ecoregion = ecoregion,
                                    level = level,
                                    join_by_map = join_by_map)
      
    }
  
  }
    
  }

  # ---------------------------------------------------------------------------
  # Hierarchical gapfilling 
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "hierarchical") {
    
    for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level)

    # 1. Attempt EEZ-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "eez",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)

    # 2. Attempt realm-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "realm",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)

    # 3. Attempt region-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "region",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)

    # 4. Attempt global-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "global",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    }
    
  }
  

  
  # ---------------------------------------------------------------------------
  # Diagonal gapfilling 
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "diagonal") {
    
    # 1. Initialize the 7x4 matrix
    pattern <- matrix(data = c(1,3,6,10,2,5,9,14,4,8,13,18,7,12,
                               17,21,11,16,20,23,15,19,22,24), nrow = 6, ncol = 4,
                      byrow = TRUE)
    
    colnames(pattern) <- c("eez", "realm", "region", "global")
    rownames(pattern) <- c("genus", "family", "order", "class", "phylum", "kingdom")
    
    
    lookup <- expand.grid(row = rownames(pattern), col = colnames(pattern))
    lookup$value <- as.vector(pattern)
    
    lookup <- lookup[order(lookup$value), ]
    
    lookup$col <- as.character(lookup$col)
    lookup$row <- as.character(lookup$row)
    
    for (i in 1:nrow(lookup)) {
      
      # Set join by map key (e.g., join by EEZ & genus, EEZ & family, realm & phylum, etc)
      join_by_map <- setNames("taxa_name", lookup[,1][i])
      
      out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = lookup$col[lookup$value == i],
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = lookup$row[lookup$value == i],
                                  join_by_map = join_by_map)
      
    }
    
  }

  # Post gapfilling, delineate whether the gapfilling
  # was not needed or was unable to be performed.
  # Gapfill levels / methods are otherwise left unchanged.
  
  out_df <- out_df %>%
    mutate(gapfill_method = case_when(
      !is.na(ssp126) & is.na(gapfill_method) ~ "Gapfill not needed",
      is.na(ssp126) ~ "Unsuccessful gapfill",
      TRUE ~ gapfill_method 
    ),
    gapfill_level = case_when(
      !is.na(ssp126) & is.na(gapfill_level) ~ "Gapfill not needed",
      is.na(ssp126) ~ "Unsuccessful gapfill",
      TRUE ~ gapfill_level 
    ),
    averaging = case_when(!(gapfill_method == "Gapfill not needed" |
                          gapfill_method == "Unsuccessful gapfill") ~ averaging,
                          TRUE ~ NA_character_) # define whether weighted or unweighted averaging was used.
    ) %>% 
    select(-taxa_level) # not needed in final output
  
  return(out_df)
}

build_exposure <- function(dir, portfolio, gap_filling_method, averaging, consumption, num_species) {
  # Load in file names and species names acquired from the sau mcp data
  file_names <- readRDS("../data/exposure/utils/file_names.rds")
  species_names <- readRDS("../data/exposure/utils/species_names.rds")
  
  ## Read in Juliano's mcp data
  climate_by_spp_eez <- read_csv("../output/future_catch_data_joined/future_catch_joined.csv", show_col_types = FALSE)
  
  # File integrity diagnostics of Juliano's climate data
  
  filename_keys <- as.integer(str_extract(basename(file_names), "\\d+"))
  
  # Scan all files once, collecting all needed info in a single pass
  file_scan <- map_dfr(file_names, function(f) {
    df_i <- fread(f, colClasses = list(character = c("eez_name", "ssp")))
    filename_key <- as.integer(str_extract(basename(f), "\\d+"))
    
    if (nrow(df_i) == 0) {
      return(tibble(filename_key = filename_key, actual_key = NA_integer_, n_rows = 0L, found_in_file = basename(f)))
    }
    
    tibble(filename_key = filename_key, actual_key = unique(df_i$taxon_key), n_rows = nrow(df_i), found_in_file = basename(f))
  })
  
  # File has data, but no accompanying species key
  data_but_no_key <- climate_by_spp_eez %>%
    filter(is.na(sciname)) %>%
    group_by(sciname) %>%
    distinct(file_found_in)
  
  # No file at all
  no_file <- species_names %>% 
    filter(!taxon_key %in% filename_keys) %>%
    as_tibble() %>%
    mutate(issue = "no file", found_in_file = NA_character_)
  
  # Empty files
  empty_files <- file_scan %>%
    filter(n_rows == 0) %>%
    left_join(as_tibble(species_names), by = c("filename_key" = "taxon_key")) %>%
    mutate(issue = "empty file") %>%
    select(taxon_key = filename_key, taxon_name, issue, found_in_file)
  
  # Mismatched keys — also check if no_file species are hiding in these files
  mismatched <- file_scan %>%
    filter(n_rows > 0, filename_key != actual_key) %>%
    left_join(as_tibble(species_names), by = c("actual_key" = "taxon_key")) %>%
    mutate(issue = "mismatched key") %>%
    select(taxon_key = actual_key, taxon_name, issue, found_in_file)
  
  # Check if any no_file species are hiding inside mismatched files
  no_file_found <- no_file %>%
    filter(taxon_key %in% mismatched$taxon_key) %>%
    left_join(mismatched %>% select(taxon_key, found_in_file), by = "taxon_key") %>%
    mutate(issue = "no file — found in mismatched file")
  
  no_file_not_found <- no_file %>%
    filter(!taxon_key %in% mismatched$taxon_key)
  
  # Get the files which have no data themselves and contain files from another species id
  mismatched_origin_file <- gsub(c("percentage_change_"), "", mismatched$found_in_file) 
  mismatched_origin_file <- as.numeric (gsub(".csv", "", mismatched_origin_file))
  
  mismatched_origin_file <- species_names %>%
    filter(
      taxon_key %in% 
             species_names$taxon_key[species_names$taxon_key %in% mismatched_origin_file]
    )
  
  # Correct files
  correct <- species_names %>%
    as_tibble() %>%
    filter(
      taxon_key %in% filename_keys,
      !taxon_key %in% empty_files$taxon_key,
      !taxon_key %in% mismatched$taxon_key,
      !taxon_key %in% mismatched_origin_file$taxon_key
    ) %>%
    mutate(issue = "correct", found_in_file = NA_character_)
  
  # Summary
  cat("Exposure data quality__________________________________________________________________", "\n")
  cat("File has data, but no accompanying species key                                     ", nrow(data_but_no_key), "\n")
  cat("No file (truly missing):                                                           ", nrow(no_file_not_found), "\n")
  cat("File unavailable, but data found in other file:                                    ", nrow(mismatched), "\n")
  cat("File available, but contains another species data (original data is nowhere else): ", nrow(mismatched_origin_file), "\n")
  cat("Empty file:                                                                        ", nrow(empty_files), "\n")
  cat("Incorrect:                                                                         ", nrow(no_file_not_found) + nrow(data_but_no_key) + nrow(no_file_found) + nrow(empty_files) + nrow(mismatched) + nrow(mismatched_origin_file), "\n")
  cat("Correct:                                                                           ", nrow(correct) - nrow(data_but_no_key), "\n")
  cat("Total accounted for:                                                               ", nrow(no_file_not_found) + nrow(no_file_found) + nrow(empty_files) + nrow(mismatched) + nrow(mismatched_origin_file) + nrow(correct), "\n")
  cat("Total in species_names:                                                            ", nrow(species_names), "\n")
  
  ## Remove file_found_in variable (only needed for data coverage diagnostics)
  ## Also drop NAs so that no na EEZs are joined to potential NAs in consumption
  ## data with EEZs
  
  ## Read in Juliano's mcp data
  climate_by_spp_eez <- climate_by_spp_eez %>%
    select(-file_found_in) %>%
    drop_na()
  
  # Read data
  
  ## Read in ARTIS sciname data
  sciname_taxa <- read_csv(paste0(dir, "sciname.csv"), show_col_types = FALSE) %>%
    build_taxa_levels()
  
  ## Read in ecoregion data
  ### Load in ecoregion data (i.e., average mcp values across ecoregion data)
  
  ### Read in ecoregions for largest soverign shape
  ecoregion_unique <- read.csv("../data/exposure/gap_filling/spatially_joined_ecoregion_to_EEZ/ecoregion_clean.csv") %>%
    select(-FREQUENCY) %>%
    rename(iso3c = "ISO_SOV1",
           shape_area = "MAX_Shape_Area")
  
  # Read in ecoregion data (all shapes associated with soverign)
ecoregion_full <- read.csv("../data/exposure/gap_filling/spatially_joined_ecoregion_to_EEZ/ecoregion.csv") %>%
  rename("iso3c" = "ISO_SOV1",
         ecoregion = "ECOREGION",
         province = "PROVINCE",
         realm = "REALM",
         shape_area = "Shape_Area")
  
  ### Join largest shape for the respective soverign country to their ecoregion names
  ecoregion <- left_join(ecoregion_unique, ecoregion_full, by = c("iso3c", "shape_area")) %>%
    filter(realm != "")
  
  ### Assign countries not associated with a marine ecoregion to "landlocked" code

  landlocked_countries <- consumption %>%
    distinct(consumer_iso3c) %>%
    drop_na() %>%
    add_region(col = "consumer_iso3c", region.col.name = "region") %>%
    filter(!consumer_iso3c %in% ecoregion$iso3c) %>%
    mutate(ecoregion = paste0("Landlocked, ", region),
           province = paste0("Landlocked, ", region),
           realm = paste0("Landlocked, ", region)) %>%
    select(-region) %>%
    rename(iso3c = consumer_iso3c)
  
  # Clean ecoregion data
  ecoregion <- bind_rows(ecoregion, landlocked_countries) %>%
    arrange(iso3c) %>%
    select(-shape_area)
  
  
  #   Correct Juliano's unaligned species names to Fishbase / SeaLifeBase naming
  ##  Problem: mcp data has (1) updated taxa names; and (2) out of date taxa names
  ##  We can't just correct all Juliano's data through Fishbase / SeaLifeBase
  ##  naming as they also have out of date items that are unaligned with ARTIS.
  ### Step 1: See which of Juliano's data when joined to 2019 ARTIS consumption data,
  ### is not joined

  #### Get list of unique ARTIS scientific names
  consumption_species_list <- consumption %>% # Filter to only species names
  distinct(sciname)

  #### Which of Juliano's species are not matching to ARTIS?
  unaligned_names <- correct$taxon_name[!str_to_lower(correct$taxon_name) %in%
                                          consumption_species_list$sciname]
  
  # Obtain fishbase / sealifebase corrections tables 
  if (!file.exists("../data/exposure/fb_slb_taxa_corrections/fishbase_corrections.rds")) {
    for (server in c("fishbase", "sealifebase")) {
    # Acquire full fishbase taxonomy
    taxa <- load_taxa(server = server, version = "25.04")
    
    # Get synonyms table
    syn <- synonyms(server = server, version = "25.04") %>%
      filter(Status == "accepted name" |
             Status == "synonym")
    
    # Join synonyms to corrected taxa names
    naming_corrections <- left_join(
      syn %>% filter(!(SpecCode == 0 | SpecCode == 1)) %>% select(synonym, SpecCode),
      taxa %>% select(SpecCode, Species)
    )
    
    # Save to RDS
    saveRDS(naming_corrections, paste0("../data/exposure/fb_slb_taxa_corrections/",server,"_corrections.rds"))
    
    }
  }
  
  # Load in Fishbase / Sealifebase taxa corrections
  fb_corrections <- readRDS("../data/exposure/fb_slb_taxa_corrections/fishbase_corrections.rds")
  slb_corrections <- readRDS("../data/exposure/fb_slb_taxa_corrections/sealifebase_corrections.rds")
  
  fb_slb_corrections <- bind_rows(fb_corrections, slb_corrections) # Join fb/slb
  
  # Join unaligned names to corrections
  corrections <- sapply(unaligned_names, function (x) unique(fb_slb_corrections$Species[x == fb_slb_corrections$synonym])) %>% unname()
  
  #### Create a separate dataframe of the catch projections that contain corrected names.
  #### We will use the Juliano's original naming as well as 
  #### the corrected one for two joins to ARTIS
  #### the first join will use Juliano's presently correct data, the second join
  #### will join Juliano's previously incorrect names (now corrected)
  
  ##### Create lookup table
  ##### (str_to_lower had to be done here becaue fb/slb lookup tables are uppercase)
  correction_lookups <- setNames(corrections, unaligned_names %>% str_to_lower())
  
  ##### Create data separate data frame
  climate_by_spp_eez_corrected <- climate_by_spp_eez %>%
  mutate(sciname = ifelse(
    sciname %in% names(correction_lookups),
    correction_lookups[sciname],
    NA
  )) %>%
    drop_na(sciname)

  ## Problem: mcp data has (1) updated taxa names; and (2) out of date taxa names
  ### Combine Juliano's pre and post-corrected naming data.
  climate_by_spp_eez_combined <- bind_rows(climate_by_spp_eez, climate_by_spp_eez_corrected) %>%
    group_by(eez_iso3c, sciname) %>%
    summarise(across(c(ssp126, ssp585), ~mean(.x, na.rm = TRUE))) %>%
    ungroup()
  
  ### For each species, average mcp projections across territories with an EEZ
  
  ### Join consumption data to catch projections
    consumption_mpc <- consumption %>%
    left_join(sciname_taxa, by = c("sciname" = "sciname")) %>%
      # Clean only artis sciname that didn't have data in sciname table
    left_join(climate_by_spp_eez_combined, by = c("eez_iso3c", "sciname")) %>%
      mutate(
        genus = case_when(
          sciname == "crassostrea gigas" ~ "crassostrea", TRUE ~ genus
        ),
        subfamily = case_when(
          sciname == "crassostrea gigas" ~ "crassostreinae", TRUE ~ subfamily
        ),
        family = case_when(
          sciname == "crassostrea gigas" ~ "ostreidae", TRUE ~ family
        ),
        order = case_when(
          sciname == "crassostrea gigas" ~ "ostreidae", TRUE ~ order
        ),
        class = case_when(
          sciname == "crassostrea gigas" ~ "bivalvia", TRUE ~ class
        ),
        phylum = case_when(
          sciname == "crassostrea gigas" ~ "mollusca", 
          sciname == "petromyzon marinus" ~ "chordata",
          TRUE ~ phylum,
        ),
        kingdom = case_when(
          sciname == "crassostrea gigas" ~ "animalia", TRUE ~ kingdom
        )
      )
  
  #### Use gap filling if user input says to, otherwise avoid
  if (gap_filling_method == "none") {
    
    x <- consumption_mpc
    
  } else {
    
    # fixit: maybe don't need lookup_table as a input here? but in the embedded
    # function
    x <- fill_with_higher_taxa(df = consumption_mpc,
                               gap_filling_method = gap_filling_method,
                               averaging = averaging,
                               ecoregion = ecoregion,
                               num_species = num_species)
    
    
  }

  ### Add taxa levels to data
  interpolated <- x %>%
    mutate(
      taxa_level = case_when(
        sciname == kingdom    ~ "kingdom",
        sciname == phylum     ~ "phylum",
        sciname == superclass ~ "superclass",
        sciname == class      ~ "class",
        sciname == order      ~ "order",
        sciname == family     ~ "family",
        sciname == subfamily  ~ "subfamily",
        sciname == genus      ~ "genus",
        str_detect(sciname, " ") ~ "species"
      ),
      species = if_else(
        str_count(sciname, " ") == 1,
        word(sciname, 2),
        NA_character_
      )
    )
  
# Save interpolated consumption data to csv if gapfilling was done
if (gap_filling_method != "none") {
  
  
  if (!file.exists(paste0("../output/consumption_projections_gap_filled/", num_species, "_species/consumption_projections_gap_filled_", portfolio, "_", averaging, "_", gap_filling_method, "_gapfill.parquet"))) {
    
    # Create directory if doesn't already exist
    if (!dir.exists(paste0("../output/consumption_projections_gap_filled/", num_species, "_species/"))) {
      
      dir.create(paste0("../output/consumption_projections_gap_filled/", num_species, "_species/"))
      
    }
    
    out_path <- paste0("../output/consumption_projections_gap_filled/", num_species, "_species/consumption_projections_gap_filled_", portfolio, "_", averaging, "_", gap_filling_method, "_gapfill.parquet")
  write_parquet(interpolated, out_path)
    
  }

}
  
  # Pivot to longer
  future_long <- interpolated %>%
    pivot_longer(c(ssp126, ssp585), names_to = "scenario", values_to = "future_change_catch")

  # Portfolio-specific row filter + catch-change multiplier
  # Note: the two portfolios intentionally use different multiplier formulas —
  # the foreign portfolio handles positive/negative/zero separately to preserve
  # directionality; the total portfolio uses a single linear conversion.
  stock_pre <- if (portfolio == "foreign_portfolio") {
    future_long %>%
      filter(
        consumer_iso3c != producer_iso3c,
        !is.na(live_weight_t),
        !is.na(future_change_catch)
      ) %>%
      mutate(future_change_catch = case_when(
        future_change_catch > 0 ~ (future_change_catch / 100) + 1,
        future_change_catch < 0 ~ 1 - ((future_change_catch / 100) * -1),
        future_change_catch == 0 ~ 1
      ))
  } else {
    future_long %>%
      filter(!is.na(live_weight_t), !is.na(future_change_catch)) %>%
      mutate(future_change_catch = 1 + (future_change_catch / 100))
  }

  # For each country, calculate the % change in aquatic food supply by calculating
  # a weighted average of each consumed taxa's projected percentage change in 
  # catch potential catch multiplied by by their respective proportional live
  # weight tonnage of their total consumed weight.
  
  # catch projection percentages by their proportional live weight tonnage
  # and converting to  
  exposure <- stock_pre %>%
    mutate(change_in_stock = live_weight_t * future_change_catch) %>%
    group_by(consumer_iso3c, scenario) %>% # FIXIT: changed from consumer_iso3c to producer_iso3c - switch back
    summarise(
      change_in_stock = sum(change_in_stock),
      live_weight_t   = sum(live_weight_t)
    ) %>%
    ungroup() %>%
    mutate(pct_change = 100 * ((change_in_stock - live_weight_t) / live_weight_t)) # FIXIT: Use a pipe here if I am testing the expsoure
    # On producer_iso3c
    # rename(consumer_iso3c = producer_iso3c) # FIXIT: Remove this as well when done testing
  
  # Return exposure dataframe
  return(exposure)
}
```

---

# Step 3 — Sensitivity (per portfolio)

Computes aquatic animal reliance as a share of total animal protein supply.

- `"foreign"` — sums across all consumption sources (domestic + foreign) to
  get total reliance.
- `"total"` — retains only the `foreign` consumption source row (no
  aggregation needed).

```{r build-sensitivity}
build_sensitivity <- function(dir) {

  supply_importance <- read_csv(
    paste0(dir, "sensitivity/fao_aquatic_reliance_source.csv"),
    show_col_types = FALSE
  )

  base <- supply_importance %>%
    filter(year == 2019, habitat == "marine", method == "capture", food_group == "aquatic")
  
  # Calculate reliance on foreign imported marine capture foods
  foreign <- base %>%
      filter(consumption_source == "foreign") %>%
      select(iso3c, foreign_reliance_pct = prop_animal_protein)
  
  # Calculate reiance on marine capture
  mc <- base %>%
      group_by(iso3c) %>%
      summarise(mc_reliance_pct = sum(prop_animal_protein)) %>%
      ungroup()
  
  # combine foreign / marine capture reliance data
  foreign_mc <- left_join(foreign, mc, by = "iso3c")
  
  return(foreign_mc)

}
```

---

# Step 4 — Adaptive capacity (portfolio-agnostic)

Loads all ten AC indicators across the four IPCC dimensions (Assets,
Flexibility, Learning, Social Organization), standardises column names, and
returns a single joined data frame keyed on `consumer_iso3c`.

The `consuming_countries` argument (a vector of ISO3C codes) is passed in
explicitly so this function has no hidden dependency on objects created in
earlier steps.

```{r build-adaptive-capacity}
build_adaptive_capacity <- function(dir, consuming_countries) {

  cc_df <- tibble(consumer_iso3c = consuming_countries, artis = 1)

  # country_coverage() returns the ARTIS countries that are missing from a
  # given AC dataset — useful for diagnosing coverage gaps
  country_coverage <- function(data, join_by) {
    full_join(cc_df, data %>% mutate(covered = 1), by = c("consumer_iso3c" = join_by)) %>%
      filter(is.na(covered), artis == 1) %>%
      pull(consumer_iso3c)
  }

  # --- Assets ---
  sanitation <- read_csv(paste0(dir, "adaptive capacity/assets/sanitation.csv"), show_col_types = FALSE) %>%
    filter(Indicator_Code == "SH_STA_BASS_ZS", Time_Period == 2019) %>%
    rename(sanitation = Value, consumer_iso3c = Geography_Code) %>%
    select(consumer_iso3c, sanitation)

  gdp <- read_csv(paste0(dir, "adaptive capacity/assets/gdp.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, gdp = `2019`) %>%
    filter(!is.na(gdp))

  trade_gdp <- read_csv(paste0(dir, "adaptive capacity/assets/trade_gdp.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Code` == "NE.TRD.GNFS.ZS") %>%
    select(consumer_iso3c = `Country Code`, gdp_trade = `2019`) %>%
    filter(!is.na(gdp_trade))

  # --- Flexibility ---
  life_expectancy <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/life_expectancy_at_birth.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator Code` == "SP.DYN.LE00.IN") %>%
    select(consumer_iso3c = `Country Code`, life_expectancy = `2019`) %>%
    filter(!is.na(life_expectancy))

  supermarkets <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/supermarkets_per_100000.csv"),
    show_col_types = FALSE
  ) %>%
    mutate(consumer_iso3c = countrycode(Region, origin = "country.name", destination = "iso3c")) %>%
    select(consumer_iso3c, supermarkets = `2019`) %>%
    filter(!is.na(supermarkets))

  total_labor   <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_labor_force.csv"), show_col_types = FALSE)
  total_pop     <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_population.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, total_pop_2019 = `2019`)

  prop_labor <- total_labor %>%
    left_join(total_pop, by = c("Country Code" = "consumer_iso3c")) %>%
    select(consumer_iso3c = `Country Code`, labor_2019 = `2019`, total_pop_2019) %>%
    mutate(prop_labor = labor_2019 / total_pop_2019) %>%
    filter(!is.na(prop_labor)) %>%
    select(consumer_iso3c, prop_labor)

  # --- Learning ---
  hci <- read_csv(paste0(dir, "adaptive capacity/learning/HCIData.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Name` == "Human Capital Index (HCI) (scale 0-1)") %>%
    select(consumer_iso3c = `Country Code`, hci = `2018`) %>%
    filter(!is.na(hci))

  # --- Social organization ---
  gov_effectiveness <- read_csv(
    paste0(dir, "adaptive capacity/social organization/government_effectiveness_percentile.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator ID` == "WB.WWGI.GE.PER.RNK") %>%
    select(consumer_iso3c = `Economy ISO3`, gov_effectiveness = `2019`) %>%
    filter(!is.na(gov_effectiveness))

  fsc <- read_csv(
    paste0(dir, "adaptive capacity/social organization/food-systems-dashboard-2025-03-04.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Start Year` == 2019 | `End Year` == 2019) %>%
    rename(fsc = Value, consumer_iso3c = ISO3) %>%
    select(consumer_iso3c, fsc)

  rol <- read_csv(
    paste0(dir, "adaptive capacity/social organization/WB-WWGI.csv"),
    show_col_types = FALSE
  ) %>%
    filter(Indicator == "Rule of Law: Percentile Rank") %>%
    select(consumer_iso3c = `Economy ISO3`, rol = `2019`)

  # Join all AC indicators
  gdp %>%
    left_join(trade_gdp,        by = "consumer_iso3c") %>%
    left_join(sanitation,       by = "consumer_iso3c") %>%
    left_join(supermarkets,     by = "consumer_iso3c") %>%
    left_join(life_expectancy,  by = "consumer_iso3c") %>%
    left_join(prop_labor,       by = "consumer_iso3c") %>%
    left_join(hci,              by = "consumer_iso3c") %>%
    left_join(gov_effectiveness, by = "consumer_iso3c") %>%
    left_join(fsc,              by = "consumer_iso3c") %>%
    left_join(rol,              by = "consumer_iso3c")
}
```

---

# Orchestrator — `build_e_s_ac_data()`

Calls `build_exposure()`, `build_sensitivity()`, and
`build_adaptive_capacity()` for a given portfolio type, joins the results, and
writes the output parquet.

```{r build-e-s-ac-data}
build_e_s_ac_data <- function(portfolio = c("foreign_portfolio", "total_portfolio"),
                              dir,
                              gap_filling_method = c("none", "region", "realm", "thorough", "hierarchical", "diagonal"),
                              averaging = c("unweighted", "weighted", "none"),
                              consumption,
                              num_species) {
  
  gap_filling_method <- match.arg(gap_filling_method)
  averaging <- match.arg(averaging)
  portfolio <- match.arg(portfolio)
  cat(sprintf("\n--- Building E/S/AC data: %s portfolio ---\n", portfolio))
  cat(sprintf("%s gap filling, %s averaging; %d minimum species\n", gap_filling_method, averaging, num_species))

  exposure     <- build_exposure(dir, portfolio, gap_filling_method = gap_filling_method, averaging = averaging, consumption = consumption, num_species = num_species)
  sensitivity  <- build_sensitivity(dir)

  consuming_countries <- exposure %>% distinct(consumer_iso3c) %>% pull()
  adaptive_cap <- build_adaptive_capacity(dir, consuming_countries)

  e_s_ac <- adaptive_cap %>%
    right_join(sensitivity,  by = c("consumer_iso3c" = "iso3c")) %>%
    left_join(exposure,      by = "consumer_iso3c")
  
  # Create directory if doesn't already exist
  if (!dir.exists(paste0("../output/e_s_ac/", num_species, "_species/"))) {
    
    dir.create(paste0("../output/e_s_ac/", num_species, "_species/"))
    
  }
  
  if (gap_filling_method == "none") {
    out_path <- paste0("../output/e_s_ac/", num_species, "_species/e_s_ac_", portfolio, "_", gap_filling_method, "_gapfill.csv") # FIXIT: added producer, need to remove "producer" after testing
  } else {
    out_path <- paste0("../output/e_s_ac/",num_species,"_species/e_s_ac_", portfolio, "_", averaging, "_", gap_filling_method, "_gapfill.csv") # FIXIT: added producer, need to remove "producer" after testing
  }

  
  write_csv(e_s_ac, out_path)
  cat(sprintf("✅ Saved %s\n", out_path))

  invisible(e_s_ac)
}
```
