---
title: "01 — Climate Risk Analysis"
author: "Connor Quiroz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(gtools)
```

# Objectives

This notebook:

1. Runs `00_Preprocess_Data.Rmd` to load all data-building functions into the
   environment.
2. Builds E/S/AC datasets for two portfolio types: a country's **foreign
   imports** and its **entire consumption** portfolio.
3. Calculates climate risk for each portfolio.
4. Runs a sensitivity analysis on adaptive capacity components *(work in
   progress)*.

---

# Setup

> **User input:** Set `file_path` to the root of your input data folder.
> Everything else runs automatically.

## Calculate the risk index based on a country’s exposure, sensitivity, and adaptive capacity components.

```{r load-working-directory}
# Set working directory for pulling in data easily path here

## Connor's laptop path
file_path <- "C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/data/"

## Jessica's path:
# file_path <- ""
```


```{r load-functions}
# Knitting 00_Preprocess_Data.Rmd loads all helper functions for generating the exposure, sensitiviyty, adaptive capacity.
knitr::knit("00_Preprocess_Data.Rmd", output = tempfile())

source("../utils/create_map.R") ## For creating maps
source("../utils/calculate_vulnerability.R") ## For calculating vulnerability
source("../utils/calculate_simulated_vulnerability.R") ## For calculating simulated vulnerability (sensitivity analysis)
# source("../utils/sens_sensitivity_analysis.R")  ## Sensitivity analysis on the sensitivity measure (FIXIT: work in progress)
source("../utils/exp_sensitivity_analysis.R") ## Sensitivity analysis on the exposure measure
source("../utils/sens_ac_sensitivity_analysis.R") ## Sensitivity analysis on the exposure measure
source("../utils/number_unique_combs.R")  # Calculate the different unique combinations given the number of features (Used in sensitivity analysis)
```

---

# Climate data preprocessing

Run once per session — reads DBEM species files and writes
`future_climate_joined.csv` to `../output/`. Subsequent calls to
`build_e_s_ac_data()` read from that file, so re-running this is only needed
if the underlying DBEM inputs change.

```{r preprocess-climate}
# preprocess_climate_data(dir = file_path)
```

---

# Build Exposure / Sensitivity / Adaptive capacity datasets

This pulls from the *00_Preprocess_Data.Rmd* file

```{r build-e_s_ac-data}
# Build e/s/ac datasets

## Build exposure / Sensitivity / Adaptive capacity dataset for total consumption (no gapfilling)

### Only run if file exists
if (!file.exists("../output/e_s_ac/e_s_ac_total_none_gapfill.parquet")) {
  e_s_ac_total_no_gapfill    <- build_e_s_ac_data(portfolio = "total",
                                                dir = file_path,
                                                gap_filling_method = "none")
} else {
  e_s_ac_total_no_gapfill <- read_parquet("../output/e_s_ac/e_s_ac_total_none_gapfill.parquet")
}

## Build Exposure / Sensitivity / Adaptive capacity dataset for total consumption (eez --> region gapfilling)

if (!file.exists("../output/e_s_ac/e_s_ac_total_region_gapfill.parquet")) {
  e_s_ac_total_region_gapfill    <- build_e_s_ac_data(portfolio = "total",
                                                dir = file_path,
                                                gap_filling_method = "region")
} else {
  e_s_ac_total_region_gapfill <- read_parquet("../output/e_s_ac/e_s_ac_total_region_gapfill.parquet")
}

## Build Exposure / Sensitivity / Adaptive capacity dataset for total consumption (eez --> realm gapfilling)
if (!file.exists("../output/e_s_ac/e_s_ac_total_realm_gapfill.parquet")) {
  e_s_ac_total_realm_gapfill    <- build_e_s_ac_data(portfolio = "total",
                                                dir = file_path,
                                                gap_filling_method = "realm")
} else {
  e_s_ac_total_realm_gapfill <- read_parquet("../output/e_s_ac/e_s_ac_total_realm_gapfill.parquet")
}
```
---

# Calculate climate vulnerability

This pulls from the *calculate_vulnerability.R* file.

```{r calculate-vulnerability}
source("../utils/calculate_vulnerability.R") ## For calculating vulnerability

# Calculate climate vulnerability of total consumption (no gapfilling)
vuln_total_no_gapfill <- calculate_vulnerability(e_s_ac_total_no_gapfill,   coverage = "entire portfolio")

# Calculate climate vulnerability of total consumption (eez --> region gapfilling)
vuln_total_realm_gapfill    <- calculate_vulnerability(e_s_ac_total_realm_gapfill,   coverage = "entire portfolio")

# Calculate climate vulnerability of total consumption (eez --> realm gapfilling)
vuln_total_region_gapfill    <- calculate_vulnerability(e_s_ac_total_region_gapfill,   coverage = "entire portfolio")

```

```{r FIXIT delete}
create_map(data = vuln_total_no_gapfill, fill = "vulnerability", country.col.name = "consumer_iso3c")

create_map(data = vuln_total_region_gapfill, fill = "vulnerability", country.col.name = "consumer_iso3c")

create_map(data = vuln_total_realm_gapfill, fill = "vulnerability", country.col.name = "consumer_iso3c")


# hist(round(vuln_total_no_gapfill$sensitivity,2))
```

# Sensitivity analyis

## Analyze the sensitivity of different exposure measures

```{r exposure measurement: sensitivity analysis}
# Create a list out of the different exposure calculations
exp_list <- list(vuln_total_no_gapfill$exposure, vuln_total_region_gapfill$exposure, vuln_total_realm_gapfill$vulnerability)

# Run sensitivity analysis
dir_probs_matrix <- exp_sensitivity_analysis(exp_list = exp_list)-

# Plot the correlation between methods of exposure.
(plot <- data.frame(dir_probs_matrix) %>%
  drop_na() %>%
    mutate(combination = factor(combination)) %>%
  mutate(combination = case_when( # Change code for how many exposure methods are being compared
    combination == 1 ~ "None vs region",
    combination == 2 ~ "None vs realm",
    combination == 3 ~ "Region vs realm"
    )
  ) %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = combination)) +
  geom_line() +
  facet_wrap(~ combination) +
  labs(y = "Kendall tau correlation", x = "Dirichlet-assigned weight for focal exposure method", fill = "Variable", title = "Sensitivity of exposure score") +
  theme_cowplot() +
  guides(color = "none") +
  theme(axis.text = element_text(size = 8),
        plot.title = element_text(hjust = 0.5)) +
    annotate("text", label = "Min") +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.8
  ))


ggsave("../images/exploratory/sensitiity_analysis_exposure.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```

## Sensitivity analysis of sensitivity measure

```{r sensitivity measurement: sensitivity analysis}
# Run sensitivity analysis
dir_probs_matrix_sensitivity <- sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                                                             analysis = "sensitivity")

plot <- dir_probs_matrix_sensitivity %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var),
         var = case_when(var == "1" ~ "Foreign reliance (%)",
                         var == "2" ~ "Marine capture reliance (%)")) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal sensitivity variable", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_sensitivity.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```

## Sensitivity analysis on adaptive capacity variables

### Variables within adaptive capacity

```{r adaptive capacity measurement: sensitivity analysis}
dir_probs_matrix_ac_variables <- sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                                                             analysis = "adaptive_capacity_variables")

plot <- dir_probs_matrix_ac_variables %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var),
         var = case_when(var == "1" ~ "GDP per capita",
                         var == "2" ~ "Trade (% of GDP)",
                         var == "3" ~ "Sanitation",
                         var == "4" ~ "Supermarkets",
                         var == "5" ~ "Life expectancy",
                         var == "6" ~ "Proportion labor force",
                         var == "7" ~ "Human capacity index",
                         var == "8" ~ "Government effectiveness",
                         var == "9" ~ "Food safety capacity",
                         var == "10" ~ "Rule of law")) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
    theme(legend.position = "none",
          axis.text = element_text(size = 8),
          strip.text = element_text(size = 7),
          plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal adaptive capacity variable", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_adaptive_capacity_variables.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```
### Components within adaptive capacity

```{r}
dir_probs_matrix_ac_components <- 
  sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                               analysis = "adaptive_capacity_components")

plot <- dir_probs_matrix_ac_components %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var)) %>%
  mutate(var = case_when(
         var == "1" ~ "Assets",
         var == "2" ~ "Flexibility",
         var == "3" ~ "Social Organization",
         var == "4" ~ "Learning"),
         dirichlet_prob = case_when(
           var != "Learning" ~ dirichlet_prob * 3,
           TRUE ~ dirichlet_prob
         )) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
    theme(legend.position = "none",
          axis.text = element_text(size = 8),
          strip.text = element_text(size = 7),
          plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal adaptive capacity component", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_adaptive_capacity_components.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```



# Analyze CV of different inerpolation by taxonomic level, by eez 

```{r}
consumption_mpc_realm <- read_parquet("../output/future_catch_joined_consumption/interpolated_future_data_realm_gapfill.parquet")

consumption_mpc_region <- read_parquet("../output/future_catch_joined_consumption/interpolated_future_data_region_gapfill.parquet")



consumption_mpc_region %>%
  mutate(
      gapfill_method_level = case_when(
        gapfill_method == "eez_gapfill" & gapfill_level == "subfamily" ~ "EEZ: Subfamily",
        gapfill_method == "eez_gapfill" & gapfill_level == "genus" ~ "EEZ: Genus",
        gapfill_method == "eez_gapfill" & gapfill_level == "family" ~ "EEZ: Family",
        gapfill_method == "eez_gapfill" & gapfill_level == "order" ~ "EEZ: Order",
        gapfill_method == "eez_gapfill" & gapfill_level == "class" ~ "EEZ: Class",
        gapfill_method == "eez_gapfill" & gapfill_level == "phylum" ~ "EEZ: Phylum",
        
        gapfill_method == "reigon_gapfill" & gapfill_level == "subfamily" ~ "Region: Subfamily",
        gapfill_method == "region_gapfill" & gapfill_level == "genus" ~ "Region: Genus",
        gapfill_method == "region_gapfill" & gapfill_level == "family" ~ "Region: Family",
        gapfill_method == "region_gapfill" & gapfill_level == "order" ~ "Region: Order",
        gapfill_method == "region_gapfill" & gapfill_level == "class" ~ "Region: Class",
        gapfill_method == "region_gapfill" & gapfill_level == "phylum" ~ "Region: Phylum",
        gapfill_method == "Gapfill not needed" ~ "No gapfill needed",
        gapfill_method == "Unsuccessful gapfill" ~ "Unsuccessful gapfill"
      ),
      gapfill_method_level = factor(
        gapfill_method_level,
        levels = c(
          "EEZ: Genus",
          "EEZ: Subfamily",
          "EEZ: Family",
          "EEZ: Order",
          "EEZ: Class",
          "EEZ: Phylum",
          "Region: Genus",
          "Region: Subfamily",
          "Region: Family",
          "Region: Order",
          "Region: Class",
          "Region: Phylum",
          "No gapfill needed",
          "Unsuccessful gapfill"
        )
      )) %>%
  drop_na(ssp126) %>%
  filter(!taxa_level == "species") %>%
  group_by(eez_iso3c, sciname)  %>%
  summarize(sd = sd(ssp126, na.rm = TRUE))
    asummarize(cv = mean(ssp126, na.rm = TRUE) / sd(ssp126, na.rm = TRUE))

consumption_mpc_region %>% distinct(taxa_level)
  
y %>%
  summarize(cv = mean(ssp126, na.rm = TRUE) / sd(ssp126, na.rm = TRUE))

consumption_mpc_region %>% count(gapfill_method)
```

  
---

# Diagnostics — correlations between risk components

```{r risk-correlations}
pairs(vuln_total_no_gapfill[, 2:5])
pairs(vuln_total_region_gapfill[, 2:5])
pairs(vuln_total_realm_gapfill[, 2:5])

cor(vuln_total_no_gapfill[, 2:5])[1,] # row selected: exposure
cor(vuln_total_region_gapfill[, 2:5])[1,] # row selected: exposure
cor(vuln_total_realm_gapfill[, 2:5])[1,] # row selected: exposure

hist(e_s_ac_total_realm_gapfill$pct_change)
```
