---
title: "01 — Climate Risk Analysis"
author: "Connor Quiroz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
```

# Objectives

This notebook:

1. Runs `00_Preprocess_Data.Rmd` to load all data-building functions into the
   environment.
2. Builds E/S/AC datasets for two portfolio types: a country's **foreign
   imports** and its **entire consumption** portfolio.
3. Calculates climate risk for each portfolio.
4. Runs a sensitivity analysis on adaptive capacity components *(work in
   progress)*.

---

# Setup

> **User input:** Set `file_path` to the root of your input data folder.
> Everything else runs automatically.

```{r user-input}
# Set your path here
file_path <- "C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/data/"

# Jessica's path:
# file_path <- ""
```

```{r load-functions}
# Knitting 00_Preprocess_Data.Rmd loads all helper functions into this session
knitr::knit("00_Preprocess_Data.Rmd", output = tempfile())

source("../R/create_map.R")
source("../R/calculate_risk.R")
# source("../R/sensitivity_analysis.R")  # work in progress
```

---

# Climate data preprocessing

Run once per session — reads DBEM species files and writes
`future_climate_joined.csv` to `../output/`. Subsequent calls to
`build_e_s_ac_data()` read from that file, so re-running this is only needed
if the underlying DBEM inputs change.

```{r preprocess-climate}
preprocess_climate_data(dir = file_path)
```

---

# Build E/S/AC datasets

```{r build-esac-data}
e_s_ac_foreign  <- build_e_s_ac_data("foreign", dir = file_path)
e_s_ac_total    <- build_e_s_ac_data("total",   dir = file_path)
```

---

# Calculate risk

```{r calculate-risk}
risk_foreign  <- calculate_risk(e_s_ac_foreign, coverage = "foreign imports")
risk_total    <- calculate_risk(e_s_ac_total,   coverage = "entire portfolio")
```

---

# Diagnostics — correlations between risk components

```{r risk-correlations}
pairs(risk_foreign[, 2:5])
cor(risk_foreign[, 2:5])
```

---

# Sensitivity analysis on adaptive capacity variables

> Note: still a work in progress — results may not be reliable yet.

```{r sensitivity-analysis, eval=FALSE}
# sensitivity_analysis(data = e_s_ac_foreign, risk_data = risk_foreign)
# sensitivity_analysis(data = e_s_ac_total,   risk_data = risk_total)
```

---

# Coverage diagnostics

Checks how much of the 2019 consumption data is covered by the interpolated
DBEM climate projections.

```{r coverage-diagnostics}
consumption_2019 <- read_parquet(paste0(file_path, "example_consumption_eez_2024_12_06.parquet")) %>%
  filter(year == 2019) %>%
  group_by(consumer_iso3c, sciname_hs_modified) %>%
  summarise(live_weight_t = sum(live_weight_t), .groups = "drop")

future_changes <- read_parquet("../output/interpolated_future_data.parquet")

coverage_check <- consumption_2019 %>%
  left_join(
    future_changes %>%
      select(consumer_iso3c, sciname_hs_modified, ssp126) %>%
      distinct() %>%
      filter(!is.na(ssp126)),
    by = c("consumer_iso3c", "sciname_hs_modified")
  )

# Proportion of rows with climate projection coverage
mean(!is.na(coverage_check$ssp126))
```
