---
title: "01_Analysis"
author: "Connor Quiroz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(rfishbase)
```

# Overview FIXIT: include documentation for function arguments

This script defines functions that build the exposure, sensitivity, and
adaptive capacity (E/S/AC) components to calculate each country's, $c$,
climate vulnerability using the International Panel on Climate Change's formula:

$\text{Vulnerability}_c = \text{Sensitvity}_c + \text{Exposure}_c + \text{Adaptive Capacity}_c$

Each function handles one job:

Fixit: Update documentation of function
<!-- | Function | Runs once? | Purpose | -->
<!-- |---|---|---| -->
<!-- | `preprocess_climate_data()` | ✅ Yes | Read DBEM files → interpolate missing taxa → write parquet | -->
<!-- | `build_exposure()` | Per portfolio | Join climate projections to consumption; compute % stock change | -->
<!-- | `build_sensitivity()` | Per portfolio | Compute aquatic animal reliance by portfolio type | -->
<!-- | `build_adaptive_capacity()` | Per portfolio | Load and join all AC indicators | -->

<!-- `build_e_s_ac_data()` orchestrates the three per-portfolio steps and writes the -->
<!-- final joined dataset to `../output/`. -->

---

# Step 1 — Disaggregate Aquatic Resource Trade In Species (ARTIS) consumption data by EEZ

An aquatic food's producing country is not the same as the EEZ that it
was caught in. Here, we disaggregate ARTIS consumption by EEZ of capture
to align with SeaAroundUs catch projection data, which enables us to calculate
exposure.

---

# Step 2 — Catch projection data preprocessing (run once)

Reads all Dynamic Bioclimate Envelope Model species files, joins species names,
converts EEZ strings to ISO3C codes, and writes a single joined CSV.

---

# Step 3 — Exposure (per portfolio)

Joins climate projections to 2019 consumption data, fills missing species
values up through the taxonomic hierarchy, then computes each country's
weighted % change in available stock under SSP1-2.6 and SSP5-8.5.

Line 253 is where catch projection gap filling happens.

**Portfolio differences:**

- `"foreign"` — keeps only rows where `consumer_iso3c != producer_iso3c` and
  applies a three-way case_when for the catch-change multiplier (handles
  positive, negative, and zero separately).
- `"total"` — retains the full consumption portfolio

```{r build-exposure}
build_exposure <- function(dir, portfolio, gap_filling_method, averaging, consumption, num_species, assumed_sd, mcp_sovereign = mcp_sovereign) {
  
  # Read in data
  
  ## Read in ARTIS sciname data
  sciname_taxa <- read_csv(paste0(dir, "/miscellaneous/sciname.csv"), show_col_types = FALSE) %>%
    build_taxa_levels() %>%
    bind_rows(
      
    # Correct one species in ARTIS that has but contains no sciname data
    data.frame(
      sciname = "crassostrea gigas", genus = "crassostrea",
      family = "ostreidae", order = "ostreida", class = "bivalvia",
      phylum = "mollusca", kingdom = "animalia", taxa_level = "species"
      )
      
    )

  
  ## Read in ecoregion data
  ### Load in ecoregion data (i.e., average mcp values across ecoregion data)
  
  ### Read in ecoregions for largest sovereign shape
  ecoregion_unique <-
    read.csv(
      "../data/exposure/ecoregion_data/spatially_joined_ecoregion_to_EEZ/ecoregion_clean.csv"
      ) %>%
    select(-FREQUENCY) %>%
    rename(
      iso3c = "ISO_SOV1",
      shape_area = "MAX_Shape_Area"
      )
  
  ### Read in ecoregion data (all shapes associated with soverign)
  ecoregion_full <-
    read.csv(
      "../data/exposure/ecoregion_data/spatially_joined_ecoregion_to_EEZ/ecoregion.csv"
      ) %>%
    rename(
      "iso3c" = "ISO_SOV1",
      ecoregion = "ECOREGION",
      province = "PROVINCE",
      realm = "REALM",
      shape_area = "Shape_Area"
      )
  
  ### Join largest shape for the respective soverign country to their ecoregion names
  ecoregion <- left_join(
    ecoregion_unique,
    ecoregion_full,
    by = c("iso3c", "shape_area")
    ) %>%
    filter(realm != "")
  
  ### Assign countries not associated with a marine ecoregion to "landlocked" code
  landlocked_countries <- consumption %>%
    distinct(consumer_iso3c) %>%
    drop_na() %>%
    add_region(col = "consumer_iso3c", region.col.name = "region") %>%
    filter(!consumer_iso3c %in% ecoregion$iso3c) %>%
    mutate(
      ecoregion = paste0("Landlocked, ", region),
      province = paste0("Landlocked, ", region),
      realm = paste0("Landlocked, ", region)
      ) %>%
    select(-region) %>%
    rename(iso3c = consumer_iso3c)
  
  ### Clean ecoregion data that will be used in analysis
  ecoregion <- bind_rows(ecoregion, landlocked_countries) %>%
    arrange(iso3c) %>%
    select(-shape_area)
  
  
  #   Correct Juliano's unaligned species names to Fishbase / SeaLifeBase naming
  
  ##  Problem: mcp data has (1) updated taxa names; and (2) out of date taxa names
  ##  We can't just correct all Juliano's data through Fishbase / SeaLifeBase
  ##  naming as they also have out of date items that are unaligned with ARTIS.
  
  ### Step 1: See which scientific names don't join to 2019 ARTIS consumption
  ### is not joined

  #### Get list of unique ARTIS scientific names
  consumption_species_list <- consumption %>% # Filter to only species names
  distinct(sciname_hs_modified)

  #### Which of Juliano's species are not matching to ARTIS?
  unjoined_species <- anti_join(
  mcp_sovereign %>%
    distinct(taxon_name),
  consumption,
  by = c("taxon_name" = "sciname_hs_modified")
  )
  
  # Obtain fishbase / sealifebase synonyms for Juliano's species not
  # aligned with ARTIS 
  if (!file.exists("../data/exposure/fb_slb_taxa_corrections/fishbase_corrections.rds")) {
    
    for (server in c("fishbase", "sealifebase")) {
    # Acquire full fishbase taxonomy
    taxa <- load_taxa(server = server, version = "25.04")
    
    # Get synonyms table
    syn <- synonyms(server = server, version = "25.04") %>%
      filter(Status == "accepted name" |
             Status == "synonym")
    
    # Join synonyms to corrected taxa names
    naming_corrections <- left_join(
      syn %>% filter(!(SpecCode == 0 | SpecCode == 1)) %>% select(synonym, SpecCode),
      taxa %>% select(SpecCode, Species)
    )
    
    # Save to RDS
    saveRDS(naming_corrections, paste0("../data/exposure/fb_slb_taxa_corrections/",server,"_corrections.rds"))
    
    }
  }
  
  # Load in Fishbase / Sealifebase taxa corrections
  fb_corrections <- readRDS(
    "../data/exposure/fb_slb_taxa_corrections/fishbase_corrections.rds"
    )
  slb_corrections <- readRDS(
    "../data/exposure/fb_slb_taxa_corrections/sealifebase_corrections.rds"
    )
  
  # Combine the rows of Fishbase / Sealifebase synonyms
  fb_slb_corrections <- bind_rows(fb_corrections, slb_corrections)
  
  # Join unaligned names to corrections
  corrections <- sapply(
    unjoined_species %>%
      pull(taxon_name) %>%
      str_to_sentence(),
    function (x)
      unique(fb_slb_corrections$Species[x == fb_slb_corrections$synonym])
    ) %>%
    unname() %>%
    str_to_lower()
  
  # Create lookup table, matching Juliano's unjoined species names to their
  # most common synonym found in Fisbhase or Sealifebase
  correction_lookups <- setNames(
    corrections,
    unjoined_species %>% 
      pull() %>%
      str_to_lower() # ARTIS consumption is in lower case; this enables joins
    )

  # Correct any unjoined species names in Juliano's catch projection data
  mcp_sovereign_corrected <- mcp_sovereign %>%
    mutate(
      taxon_name = case_when(
        taxon_name %in% names(correction_lookups) ~ correction_lookups[taxon_name],
        TRUE ~ taxon_name
      ))
  
  ### Join consumption data sciname and catch projections
    consumption_mcp <- consumption %>%
      # Left join onto sciname data, which has higher-level taxa names that
      # correspond to the `sciname_hs_modified`
    left_join(sciname_taxa, by = c("sciname_hs_modified" = "sciname")) %>%
      # Clean only artis sciname that didn't have data in sciname table
    left_join(
      mcp_sovereign_corrected,
      by = c("sciname_hs_modified" = "taxon_name", "eez_iso3c"))
    
  # Perform gap filling on catch projections if the
  # user input says to, otherwise avoid
  if (gap_filling_method == "none") {
    
    x <- consumption_mcp
    
  } else {
    
    x <- fill_with_higher_taxa(df = consumption_mcp,
                               gap_filling_method = gap_filling_method,
                               averaging = averaging,
                               ecoregion = ecoregion,
                               num_species = num_species,
                               assumed_sd = assumed_sd)
   }

  # Add taxa levels to data
  interpolated <- x %>%
    mutate(
      taxa_level = case_when(
        sciname_hs_modified == kingdom    ~ "kingdom",
        sciname_hs_modified == phylum     ~ "phylum",
        sciname_hs_modified == superclass ~ "superclass",
        sciname_hs_modified == class      ~ "class",
        sciname_hs_modified == order      ~ "order",
        sciname_hs_modified == family     ~ "family",
        sciname_hs_modified == subfamily  ~ "subfamily",
        sciname_hs_modified == genus      ~ "genus",
        str_detect(sciname_hs_modified, " ") ~ "species"
      ),
      species = if_else(
        str_count(sciname_hs_modified, " ") == 1,
        word(sciname_hs_modified, 2),
        NA_character_
      )
    )

# Save interpolated consumption data if gap-filling was performed
if (gap_filling_method != "none") {
  
  out_dir <- if (gap_filling_method == "bayesian") {
    paste0(
      "../output/consumption_catch_proj_gap_filled/bayesian/",
      assumed_sd,
      "_sd/"
    )
  } else {
    paste0(
      "../output/consumption_catch_proj_gap_filled/",
      num_species,
      "_species/"
    )
  }
  
  # Create output directory if it doesn't exist
  if (!dir.exists(out_dir)) {
    dir.create(out_dir, recursive = TRUE)
  }
  
  # Construct output file path
  out_path <- paste0(
    out_dir,
    "consumption_",
    portfolio, "_",
    averaging, "_",
    gap_filling_method,
    "_gapfill.parquet"
  )
  
  # Write only if the file doesn't already exist
  if (!file.exists(out_path)) {
    write_parquet(interpolated, out_path)
    cat("✅ Gap-filled consumption data saved to:", out_path, "\n")
  } else {
    cat("📄 Gap-filled consumption file already exists. Skipping write:", out_path, "\n")
  }
}
  
  # Pivot to longer
  future_long <- interpolated %>%
    pivot_longer(c(ssp126, ssp585), names_to = "scenario", values_to = "future_change_catch")

  # Portfolio-specific row filter + catch-change multiplier
  # Note: the two portfolios intentionally use different multiplier formulas —
  # the foreign portfolio handles positive/negative/zero separately to preserve
  # directionality; the total portfolio uses a single linear conversion.
  stock_pre <- if (portfolio == "foreign_portfolio") {
    future_long %>%
      filter(
        consumer_iso3c != producer_iso3c,
        !is.na(live_weight_t),
        !is.na(future_change_catch)
      ) %>%
      mutate(future_change_catch = 1 + (future_change_catch / 100))
  } else {
    future_long %>%
      filter(!is.na(live_weight_t), !is.na(future_change_catch)) %>%
      mutate(future_change_catch = 1 + (future_change_catch / 100))
  }

  # For each country, calculate the % change in aquatic food supply by calculating
  # a weighted average of each consumed taxa's projected percentage change in 
  # catch potential catch multiplied by by their respective proportional live
  # weight tonnage of their total consumed weight.
  
  # catch projection percentages by their proportional live weight tonnage
  # and converting to  
  exposure <- stock_pre %>%
    mutate(change_in_stock = live_weight_t * future_change_catch) %>%
    group_by(consumer_iso3c, scenario) %>% # FIXIT: changed from consumer_iso3c to producer_iso3c - switch back
    summarise(
      change_in_stock = sum(change_in_stock),
      live_weight_t   = sum(live_weight_t)
    ) %>%
    ungroup() %>%
    mutate(pct_change = 100 * ((change_in_stock - live_weight_t) / live_weight_t)) # FIXIT: Use a pipe here if I am testing the expsoure
    # On producer_iso3c
    # rename(consumer_iso3c = producer_iso3c) # FIXIT: Remove this as well when done testing
  
  # Return exposure dataframe
  return(exposure)
}
```

---

# Step 4 — Sensitivity (per portfolio)

Computes aquatic animal reliance as a share of total animal protein supply.

- `"foreign"` — sums across all consumption sources (domestic + foreign) to
  get total reliance.
- `"total"` — retains only the `foreign` consumption source row (no
  aggregation needed).

```{r build-sensitivity}
build_sensitivity <- function(dir) {

  supply_importance <- read_csv(
    paste0(dir, "sensitivity/fao_aquatic_reliance_source.csv"),
    show_col_types = FALSE
  )

  base <- supply_importance %>%
    filter(year == 2019, habitat == "marine", method == "capture", food_group == "aquatic")
  
  # Calculate reliance on foreign imported marine capture foods
  foreign <- base %>%
      filter(consumption_source == "foreign") %>%
      select(iso3c, foreign_reliance_pct = prop_animal_protein)
  
  # Calculate reiance on marine capture
  mc <- base %>%
      group_by(iso3c) %>%
      summarise(mc_reliance_pct = sum(prop_animal_protein)) %>%
      ungroup()
  
  # combine foreign / marine capture reliance data
  foreign_mc <- left_join(foreign, mc, by = "iso3c")
  
  return(foreign_mc)

}
```

---

# Step 5 — Adaptive capacity (portfolio-agnostic)

Loads all ten AC indicators across the four IPCC dimensions (Assets,
Flexibility, Learning, Social Organization), standardizes column names, and
returns a single joined data frame keyed on `consumer_iso3c`.

The `consuming_countries` argument (a vector of ISO3C codes) is passed in
explicitly so this function has no hidden dependency on objects created in
earlier steps.

```{r build-adaptive-capacity}
build_adaptive_capacity <- function(dir, consuming_countries) {

  cc_df <- tibble(consumer_iso3c = consuming_countries, artis = 1)

  # country_coverage() returns the ARTIS countries that are missing from a
  # given AC dataset — useful for diagnosing coverage gaps
  country_coverage <- function(data, join_by) {
    full_join(cc_df, data %>% mutate(covered = 1), by = c("consumer_iso3c" = join_by)) %>%
      filter(is.na(covered), artis == 1) %>%
      pull(consumer_iso3c)
  }

  # --- Assets ---
  sanitation <- read_csv(paste0(dir, "adaptive capacity/assets/sanitation.csv"), show_col_types = FALSE) %>%
    filter(Indicator_Code == "SH_STA_BASS_ZS", Time_Period == 2019) %>%
    rename(sanitation = Value, consumer_iso3c = Geography_Code) %>%
    select(consumer_iso3c, sanitation)

  gdp <- read_csv(paste0(dir, "adaptive capacity/assets/gdp.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, gdp = `2019`) %>%
    filter(!is.na(gdp))

  trade_gdp <- read_csv(paste0(dir, "adaptive capacity/assets/trade_gdp.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Code` == "NE.TRD.GNFS.ZS") %>%
    select(consumer_iso3c = `Country Code`, gdp_trade = `2019`) %>%
    filter(!is.na(gdp_trade))

  # --- Flexibility ---
  life_expectancy <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/life_expectancy_at_birth.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator Code` == "SP.DYN.LE00.IN") %>%
    select(consumer_iso3c = `Country Code`, life_expectancy = `2019`) %>%
    filter(!is.na(life_expectancy))

  supermarkets <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/supermarkets_per_100000.csv"),
    show_col_types = FALSE
  ) %>%
    mutate(consumer_iso3c = countrycode(Region, origin = "country.name", destination = "iso3c")) %>%
    select(consumer_iso3c, supermarkets = `2019`) %>%
    filter(!is.na(supermarkets))

  total_labor   <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_labor_force.csv"), show_col_types = FALSE)
  total_pop     <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_population.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, total_pop_2019 = `2019`)

  prop_labor <- total_labor %>%
    left_join(total_pop, by = c("Country Code" = "consumer_iso3c")) %>%
    select(consumer_iso3c = `Country Code`, labor_2019 = `2019`, total_pop_2019) %>%
    mutate(prop_labor = labor_2019 / total_pop_2019) %>%
    filter(!is.na(prop_labor)) %>%
    select(consumer_iso3c, prop_labor)

  # --- Learning ---
  hci <- read_csv(paste0(dir, "adaptive capacity/learning/HCIData.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Name` == "Human Capital Index (HCI) (scale 0-1)") %>%
    select(consumer_iso3c = `Country Code`, hci = `2018`) %>%
    filter(!is.na(hci))

  # --- Social organization ---
  gov_effectiveness <- read_csv(
    paste0(dir, "adaptive capacity/social organization/government_effectiveness_percentile.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator ID` == "WB.WWGI.GE.PER.RNK") %>%
    select(consumer_iso3c = `Economy ISO3`, gov_effectiveness = `2019`) %>%
    filter(!is.na(gov_effectiveness))

  fsc <- read_csv(
    paste0(dir, "adaptive capacity/social organization/food-systems-dashboard-2025-03-04.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Start Year` == 2019 | `End Year` == 2019) %>%
    rename(fsc = Value, consumer_iso3c = ISO3) %>%
    select(consumer_iso3c, fsc)

  rol <- read_csv(
    paste0(dir, "adaptive capacity/social organization/WB-WWGI.csv"),
    show_col_types = FALSE
  ) %>%
    filter(Indicator == "Rule of Law: Percentile Rank") %>%
    select(consumer_iso3c = `Economy ISO3`, rol = `2019`)

  # Join all AC indicators
  gdp %>%
    left_join(trade_gdp,        by = "consumer_iso3c") %>%
    left_join(sanitation,       by = "consumer_iso3c") %>%
    left_join(supermarkets,     by = "consumer_iso3c") %>%
    left_join(life_expectancy,  by = "consumer_iso3c") %>%
    left_join(prop_labor,       by = "consumer_iso3c") %>%
    left_join(hci,              by = "consumer_iso3c") %>%
    left_join(gov_effectiveness, by = "consumer_iso3c") %>%
    left_join(fsc,              by = "consumer_iso3c") %>%
    left_join(rol,              by = "consumer_iso3c")
}
```

---

# Step 6 — Combine exposure, sensitivity, and adaptive capacity into one dataset
`build_e_s_ac_data()`

Calls `build_exposure()`, `build_sensitivity()`, and
`build_adaptive_capacity()` for a given portfolio type, joins the results, and
writes the output parquet.

```{r build-e-s-ac-data}
build_e_s_ac_data <- function(portfolio = c("foreign_portfolio", "total_portfolio"),
                              dir,
                              gap_filling_method = c("none", "region", "realm", "thorough", "hierarchical", "diagonal", "bayesian"),
                              averaging = c("unweighted", "weighted", "none"),
                              num_species,
                              assumed_sd) {

  cat(sprintf("\n--- Building E/S/AC data: %s portfolio ---\n", portfolio))
  cat(sprintf("%s gap filling, %s averaging; %d minimum species\n", gap_filling_method, averaging, num_species))

  # Step 1 - Disaggregate ARTIS consumption by EEZ of catch
  consumption <- disagg_cons_by_eez()
  
  # Step 2 - Preprocess SeaAroundUs catch projection data
  mcp_sovereign <- preprocess_catch_projection_data(dir = dir)
  
  # Step 3 - Build exposure
  exposure     <- build_exposure(
    dir,
    portfolio,
    gap_filling_method = gap_filling_method,
    averaging = averaging,
    consumption = consumption,
    num_species = num_species,
    assumed_sd = assumed_sd,
    mcp_sovereign = mcp_sovereign
    )
  
  # Step 4 - Build sensitivity
  sensitivity  <- build_sensitivity(dir)

  # Step 5 - Build adaptive capacity
  adaptive_cap <- build_adaptive_capacity(
    dir,
    consuming_countries = exposure %>%
      distinct(consumer_iso3c) %>%
      pull()
    )

  # Step 6 - Combine exposure, sensitivity, and adaptive capacity into one
  # dataset
  e_s_ac <- adaptive_cap %>%
    right_join(sensitivity,  by = c("consumer_iso3c" = "iso3c")) %>%
    left_join(exposure,      by = "consumer_iso3c")
  
  # Determine output directory
  if (gap_filling_method == "bayesian") {
    
    output_dir <- file.path(
      "../output/e_s_ac",
      paste0(assumed_sd, "_sd")
    )
    
  } else {
    
    output_dir <- file.path(
      "../output/e_s_ac",
      paste0(num_species, "_species")
    )
  }
  
  # Create directory if it doesn't already exist
  dir.create(
    output_dir,
    recursive = TRUE,
    showWarnings = FALSE
  )
  
  # Construct output file path
  out_path <- file.path(
    output_dir,
    paste0(
      "e_s_ac_",
      portfolio, "_",
      if (gap_filling_method != "none") paste0(averaging, "_") else "",
      gap_filling_method,
      "_gapfill.csv"
    )
  )
  
  # Write E/S/AC dataset to csv
  write_csv(e_s_ac, out_path)
  
  # Return message to console
  cat(sprintf("✅ Saved %s\n", out_path))

  # Return E/S/AC dataset from function
  return(e_s_ac)
}
```

---

# Step 7 — Calculate climate vulnerability

```{r}
calculate_vulnerability <- function(
    portfolio = c("foreign_portfolio", "total_portfolio"),
    dir,
    gap_filling_method = c(
      "none", "region", "realm", "thorough",
      "hierarchical", "diagonal", "bayesian"
      ),
    averaging = c("unweighted", "weighted", "none"),
    num_species,
    assumed_sd, 
    drop_na = TRUE) {
  
  # Ensure user inputs match the arguments of function to limit user input
  portfolio <- match.arg(portfolio)
  gap_filling_method <- match.arg(gap_filling_method)
  averaging <- match.arg(averaging)
  
  # Name output directory given user input
  output_dir <- file.path(
      "../output/e_s_ac",
      paste0(num_species, "_species")
    )
  
  # Name file path given user input
  out_path <- file.path(
    output_dir,
    paste0(
      "e_s_ac_",
      portfolio, "_",
      if (gap_filling_method != "none") paste0(averaging, "_") else "",
      gap_filling_method,
      "_gapfill.csv"
    )
  )
  
  # Check if exposure, sensitivity, and adaptive capacity data already exists
  if(file.exists(out_path)) {
    
    # Read e_s_ac
    cat("📄 Exposure, sensitivity, and adaptive capacity (E/S/AC) data already exists for this parameterization. Skipping E/S/AC derivation and proceeding with calculating vulnerability.\n")
    e_s_ac <- read_csv(out_path)
    
  } else {
    
    # Run function to derive exposure, sensitivity, and adaptive capacity data
    e_s_ac <- build_e_s_ac_data(
      portfolio = portfolio,
      dir = dir,
      gap_filling_method = gap_filling_method,
      averaging = averaging,
      num_species = num_species,
      assumed_sd = assumed_sd)
    
  }
  
  
  # Adaptive capacity
  
  ## the `build_e_s_ac_data()` function returns a vector of all the adaptive
  ## capacity variables for a country – here, we want to combine them into one
  ## adaptive capacity measure that can be used in
  ## the IPCC's vulnerability formula
  
  ### Here, there are four components in adaptive capacity:
  ### learning, flexibility, assets, and social organization.
  ### We give them equal importance (i.e., 25%) contribution to the adaptive
  ### capacity. Each of the variables within a component have an equal
  ### contribution to their respective component.
  
  # Calculate adaptive capacity
  adaptive_capacity_calcs <- e_s_ac %>%
    # Scale adaptive capacity variables using min/max scaling
    mutate(
      across(
        c(gdp, gdp_trade, sanitation,
          supermarkets, life_expectancy, prop_labor,
          hci, gov_effectiveness,
          fsc, rol), ~ 
          ( . - min(., na.rm = TRUE) ) /
          ( max(., na.rm = TRUE) - min(., na.rm = TRUE) ),
        .names = "{.col}_scaled")) %>%
    # Weight each AC variable within AC component
    # (will add to 4 for country if each value is it's max value)
    mutate(gdp_scaled = gdp_scaled * (1/3),
           gdp_trade_scaled = gdp_trade_scaled * (1/3),
           sanitation_scaled = sanitation_scaled * (1/3),
           supermarkets_scaled = supermarkets_scaled * (1/3),
           life_expectancy_scaled = life_expectancy_scaled * (1/3),
           prop_labor_scaled = prop_labor_scaled * (1/3),
           gov_effectiveness_scaled = gov_effectiveness_scaled * (1/3),
           fsc_scaled = fsc_scaled * (1/3),
           rol_scaled = rol_scaled * (1/3),
           hci_scaled = hci * 1) %>%
    distinct(consumer_iso3c, .keep_all = TRUE) %>%
    group_by(consumer_iso3c) %>%
    mutate(adaptive_capacity = sum(gdp_scaled, gdp_trade_scaled,
                                   sanitation_scaled, supermarkets_scaled,
                                   life_expectancy_scaled,
                                   prop_labor_scaled,
                                   hci_scaled,
                                   gov_effectiveness_scaled,
                                   fsc_scaled,
                                   rol_scaled)) %>%
    mutate(adaptive_capacity = adaptive_capacity * 0.5) %>% # Divide by two, meaning max value would now = two, which would nullify max level expsorue + sensitivity
    { if (drop_na == TRUE) filter(., !is.na(adaptive_capacity)) else . } %>%
    select(consumer_iso3c, adaptive_capacity)
  
  # Calculate sensitivity (i.e., combine reliance on foreign sources
  #                        & reliance on marine capture)
  sensitivity_calcs <- e_s_ac %>%
    mutate(
      # Scale sensitivity capacity variables using min/max scaling
      across(
        c(
          mc_reliance_pct,
          foreign_reliance_pct
          ), ~
          ( . - min(., na.rm = TRUE) ) /
          ( max(., na.rm = TRUE) - min(., na.rm = TRUE) ),
        .names = "{.col}_scaled")) %>% # scale variables using min/max scaling
    # Weight each sensitivity variable
    # (will add to 1 for country if each value is it's max value)
    mutate(mc_reliance_pct_scaled = mc_reliance_pct_scaled * (1/2),
           foreign_reliance_pct_scaled = foreign_reliance_pct_scaled * (1/2)) %>%
    distinct(consumer_iso3c, .keep_all = TRUE) %>%
    group_by(consumer_iso3c) %>%
    mutate(
      sensitivity = sum(
        mc_reliance_pct_scaled,
        foreign_reliance_pct_scaled
        )) %>%
    filter(!is.na(sensitivity)) %>%
    select(consumer_iso3c, sensitivity)
  
  
  # Calculate vulnerability following the IPCC formula
  # (vulnerability = Exposure + vulnerability - Adaptive Capacity)
  vulnerability <- left_join(
    e_s_ac %>%
      filter(scenario == "ssp126"),
    adaptive_capacity_calcs,
    by = "consumer_iso3c"
    ) %>%
    left_join(sensitivity_calcs, by = "consumer_iso3c") %>%
    # If use input is to remove NA's, this removes countries with missing
    # adaptive capacity data, otherwise keep
    { if (drop_na == TRUE) filter(., !is.na(adaptive_capacity)) else . } %>%
    # Scale exposure data using min/max scaling so values are between 0 and 1
    mutate(
      exposure =
        (pct_change - min(pct_change, na.rm = TRUE) ) /
        (max(pct_change, na.rm = TRUE) - min(pct_change,
                                              na.rm = TRUE)),
      exposure = -exposure + 1
      ) %>%
    group_by(consumer_iso3c) %>%
    select(consumer_iso3c, exposure, sensitivity, adaptive_capacity) %>%
    # Calculate vulnerability
    mutate(vulnerability = (exposure + sensitivity) - adaptive_capacity)
  
  # Final cleaning
  vulnerability <- vulnerability %>%
    add_region(col = "consumer_iso3c", region.col.name = "region") %>%
    arrange(region, .after = "consumer_iso3c")
  
  # Save vulnerability to file
  # Create directory if it doesn't already exist
  if (gap_filling_method == "bayesian") {
    
    output_dir <- file.path(
      "../output/vulnerability",
      paste0(assumed_sd, "_sd")
    )
    
  } else {
    
    output_dir <- file.path(
      "../output/vulnerability",
      paste0(num_species, "_species")
    )
  }
  
  dir.create(
    output_dir,
    recursive = TRUE,
    showWarnings = FALSE
  )
  
  # Construct output file path
  out_path <- file.path(
    output_dir,
    paste0(
      "vuln_",
      portfolio, "_",
      if (gap_filling_method != "none") paste0(averaging, "_") else "",
      gap_filling_method,
      "_gapfill.csv"
    )
  )
  
  # Write only if the file doesn't already exist
  if (!file.exists(out_path)) {
    write_csv(vulnerability, out_path)
    cat("✅ Vulnerability data saved to:", out_path, "\n")
  } else {
    cat("📄 Vulnerability file already exists. Skipping write:", out_path, "\n")
  }

  cat("Returning vulnerability\n")
  
  
  # Return climate vulnerability dataset
  return(vulnerability)
}
```