---
title: "01 — Climate Risk Analysis"
author: "Connor Quiroz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(gtools)
```

# Objectives

This notebook:

1. Runs `00_Preprocess_Data.Rmd` to load all data-building functions into the
   environment.
2. Builds E/S/AC datasets for two portfolio types: a country's **foreign
   imports** and its **entire consumption** portfolio.
3. Calculates climate risk for each portfolio.
4. Runs a sensitivity analysis on adaptive capacity components *(work in
   progress)*.

---

# Setup

> **User input:** Set `file_path` to the root of your input data folder.
> Everything else runs automatically.

## Calculate the risk index based on a country’s exposure, sensitivity, and adaptive capacity components.

```{r load-working-directory-and-functions}
# Set working directory for pulling in data easily path here

## Connor's laptop path
file_path <- "C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/data/"

dir <- file_path

## Connor's desktop path
# file_path <- "C:/Users/cjqui/OneDrive/Desktop/SGL/Chapter 2/trade-climate-risk/data/"

## Jessica's path:
# file_path <- ""

# Knitting 00_Preprocess_Data.Rmd loads all helper functions for generating the exposure, sensitiviyty, adaptive capacity.
knitr::knit("00_Preprocess_Data.Rmd", output = tempfile())

source("../utils/create_map.R") ## For creating maps
source("../utils/calculate_vulnerability.R") ## For calculating vulnerability
source("../utils/calculate_simulated_vulnerability.R") ## For calculating simulated vulnerability (sensitivity analysis)
# source("../utils/sens_sensitivity_analysis.R")  ## Sensitivity analysis on the sensitivity measure (FIXIT: work in progress)
source("../utils/exp_sensitivity_analysis.R") ## Sensitivity analysis on the exposure measure
source("../utils/sens_ac_sensitivity_analysis.R") ## Sensitivity analysis on the exposure measure
source("../utils/number_unique_combs.R")  # Calculate the different unique combinations given the number of features (Used in sensitivity analysis)
```

---

# Climate data preprocessing

Run once per session — reads DBEM species files and writes
`future_climate_joined.csv` to `../output/`. Subsequent calls to
`build_e_s_ac_data()` read from that file, so re-running this is only needed
if the underlying DBEM inputs change.

```{r preprocess-climate}
# preprocess_climate_data(dir = file_path)
```

---

# Build Exposure / Sensitivity / Adaptive capacity datasets

## Read in consumption data

```{r}
## Read in conumption data
  consumption <- read_parquet(paste0(dir, "example_consumption_eez_2024_12_06.parquet")) %>%
    filter(year == 2019, habitat == "marine", method == "capture")
```

This pulls from the *00_Preprocess_Data.Rmd* file

```{r build-e_s_ac-data}
# Build e/s/ac datasets

# ---------------------------------------------------------------------------
# No gap filling
# ---------------------------------------------------------------------------
## Total consumption portfolio: No gapfilling
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_none_gapfill.csv")) {
  e_s_ac_total_no_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "none",
                                                averaging = "none",
                                                consumption = consumption)
} else {
  e_s_ac_total_no_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_none_gapfill.csv")
}

# ---------------------------------------------------------------------------
# Region-level gap filling
# ---------------------------------------------------------------------------
### Unweighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_region_gapfill.csv")) {
  e_s_ac_total_unwgt_region_gapfill    <- build_e_s_ac_data(portfolio = "total",
                                                dir = file_path,
                                                gap_filling_method = "region",
                                                averaging = "unweighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_unwgt_region_gapfill <-
    read_csv("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_region_gapfill.csv")
}

### Weighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_weighted_region_gapfill.csv")) {
  e_s_ac_total_wgt_region_gapfill    <- build_e_s_ac_data(portfolio = "total",
                                                dir = file_path,
                                                gap_filling_method = "region",
                                                averaging = "weighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_wgt_region_gapfill <-
    read_csv("../output/e_s_ac/e_s_ac_total_portfolio_weighted_region_gapfill.csv")
}

# ---------------------------------------------------------------------------
# Realm-level gap filling
# ---------------------------------------------------------------------------
### Unweighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_realm_gapfill.csv")) {
  e_s_ac_total_unwgt_realm_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "realm",
                                                averaging = "unweighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_unwgt_realm_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_realm_gapfill.csv")
}

### Weighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_weighted_realm_gapfill.csv")) {
  e_s_ac_total_wgt_realm_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "realm",
                                                averaging = "weighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_wgt_realm_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_weighted_realm_gapfill.csv")
}

# ---------------------------------------------------------------------------
# Thorough-level gap filling (i.e., EEZ --> )
# ---------------------------------------------------------------------------
### Unweighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_thorough_gapfill.csv")) {
  e_s_ac_total_unwgt_thorough_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "thorough",
                                                averaging = "unweighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_unwgt_thorough_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_thorough_gapfill.csv")
}

### Weighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_weighted_thorough_gapfill.csv")) {
  e_s_ac_total_wgt_thorough_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "thorough",
                                                averaging = "weighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_wgt_thorough_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_weighted_thorough_gapfill.csv")
}


# ---------------------------------------------------------------------------
# Hierarchical-level gap filling
# ---------------------------------------------------------------------------
### Unweighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_hierarchical_gapfill.csv")) {
  e_s_ac_total_unwgt_hierarchical_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "hierarchical",
                                                averaging = "unweighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_unwgt_hierarchical_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_hierarchical_gapfill.csv")
}

### Weighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_weighted_hierarchical_gapfill.csv")) {
  e_s_ac_total_wgt_hierarchical_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "hierarchical",
                                                averaging = "weighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_wgt_hierarchical_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_weighted_hierarchical_gapfill.csv")
}


# ---------------------------------------------------------------------------
# Diagonal-level gap filling
# ---------------------------------------------------------------------------
### Unweighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_diagonal_gapfill.csv")) {
  e_s_ac_total_unwgt_diagonal_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "diagonal",
                                                averaging = "unweighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_unwgt_diagonal_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_unweighted_diagonal_gapfill.csv")
}

### Weighted averaging
if (!file.exists("../output/e_s_ac/e_s_ac_total_portfolio_weighted_diagonal_gapfill.csv")) {
  e_s_ac_total_wgt_diagonal_gapfill    <- build_e_s_ac_data(portfolio = "total_portfolio",
                                                dir = file_path,
                                                gap_filling_method = "diagonal",
                                                averaging = "weighted",
                                                consumption = consumption)
} else {
  e_s_ac_total_wgt_diagonal_gapfill <- read_csv("../output/e_s_ac/e_s_ac_total_portfolio_weighted_diagonal_gapfill.csv")
}
```

---

# Calculate climate vulnerability

This pulls from the *calculate_vulnerability.R* file.

```{r calculate-vulnerability}
# ---------------------------------------------------------------------------
# No gap filling
# ---------------------------------------------------------------------------
vuln_total_no_gapfill <- calculate_vulnerability(e_s_ac_total_no_gapfill, coverage = "entire portfolio")

# ---------------------------------------------------------------------------
# Realm gap filling
# ---------------------------------------------------------------------------
## Unweighted
vuln_total_unwgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_realm_gapfill,   coverage = "entire portfolio")

# Weighted
vuln_total_wgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_wgt_realm_gapfill,   coverage = "entire portfolio")

# ---------------------------------------------------------------------------
# Region-level gap filling
# ---------------------------------------------------------------------------
## Unweighted
vuln_total_unwgt_region_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_region_gapfill,   coverage = "entire portfolio")

## Weighted
vuln_total_wgt_region_gapfill    <- calculate_vulnerability(e_s_ac_total_wgt_region_gapfill,   coverage = "entire portfolio")

# ---------------------------------------------------------------------------
# Thorough gap filling
# ---------------------------------------------------------------------------
## Unweighted
vuln_total_unwgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_realm_gapfill,   coverage = "entire portfolio")

# Weighted
vuln_total_wgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_wgt_realm_gapfill,   coverage = "entire portfolio")
## Unweighted
vuln_total_unwgt_thorough_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_thorough_gapfill,   coverage = "entire portfolio")

# Weighted
vuln_total_wgt_thorough_gapfill <- calculate_vulnerability(e_s_ac_total_wgt_thorough_gapfill,   coverage = "entire portfolio")

# ---------------------------------------------------------------------------
# Hierarchical gap filling
# ---------------------------------------------------------------------------
## Unweighted
vuln_total_unwgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_realm_gapfill,   coverage = "entire portfolio")

# Weighted
vuln_total_wgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_wgt_realm_gapfill,   coverage = "entire portfolio")

## Unweighted
vuln_total_unwgt_hierarchical_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_hierarchical_gapfill,   coverage = "entire portfolio")

# Weighted
vuln_total_wgt_hierarchical_gapfill <- calculate_vulnerability(e_s_ac_total_wgt_hierarchical_gapfill,   coverage = "entire portfolio")

# ---------------------------------------------------------------------------
# Diagonal gap filling
# ---------------------------------------------------------------------------
## Unweighted
vuln_total_unwgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_realm_gapfill,   coverage = "entire portfolio")

# Weighted
vuln_total_wgt_realm_gapfill <- calculate_vulnerability(e_s_ac_total_wgt_realm_gapfill,   coverage = "entire portfolio")
## Unweighted
vuln_total_unwgt_diagonal_gapfill <- calculate_vulnerability(e_s_ac_total_unwgt_diagonal_gapfill,   coverage = "entire portfolio")

# Weighted
vuln_total_wgt_diagonal_gapfill <- calculate_vulnerability(e_s_ac_total_wgt_diagonal_gapfill,   coverage = "entire portfolio")
```

# Create maps of climate vulnerability

```{r create maps}
create_map(data = vuln_total_no_gapfill, fill = "exposure", country.col.name = "consumer_iso3c") +
  labs(fill = "Climate exposure")

create_map(data = e_s_ac_total_unwgt_hierarchical_gapfill %>%
             mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) / ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1), fill = "exposure", country.col.name = "consumer_iso3c") +
  labs(fill = "Exposure")

create_map(data = e_s_ac_total_wgt_hierarchical_gapfill %>%
             mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) / ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1), fill = "exposure", country.col.name = "consumer_iso3c") +
  labs(fill = "Exposure")

cor(vuln_total_no_gapfill$exposure, vuln_total_wgt_diagonal_gapfill$exposure, method = "kendall")

cor(vuln_total_no_gapfill$exposure, vuln_total_wgt_diagonal_gapfill$exposure, method = "kendall")

cor(vuln_total_unwgt_diagonal_gapfill$exposure, vuln_total_wgt_diagonal_gapfill$exposure, method = "kendall")
# What's contributing greater to the weighted score - Jessica
# pick a couple countries where there's data - look at the distribution for which things were filled at. Pick some taxa names, what values form which  were contributing the most to the exposure
read_parquet("../output/future_catch_joined_consumption/interpolated_future_dats_total_portfolio_unweighted_diagonal_gapfill.parquet")



a <- e_s_ac_total_unwgt_hierarchical_gapfill %>%
             mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) / ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1)

b <- e_s_ac_total_wgt_hierarchical_gapfill %>%
             mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) / ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1)

cor(a$exposure, b$exposure, method = "kendall")
```

# Sensitivity analyis

Here, we compare use Kendall tau correlation to compare differences in rank across diffrerent methods for (1) different methdods for calculating exposure; (3) different contributing meausres to sensitivity; (3) the individual variables contributing measures within adaptive capacity; and (4) the grouped measures that contributing to adaptive capacity (i.e., learning, flexibility, assets, social organization), where each grouped measure can contain multiple variables of adaptive capacity. We first run sensitivity analysis on each of the measures

## Analyze the sensitivity of different exposure measures

```{r exposure measurement: sensitivity analysis}
# Helper function to dynamically select either exposure or vulnerability columns
# to run sensitivity analysis
run_sensitivity_analysis <- function(dfs, comparison = c("exposure", "vulnerability")) {
  comparison <- match.arg(comparison)

  list <- lapply(dfs, function(df) df[[comparison]])   # or `pull()` if you want a vector
  
  # Run sensitivity analysis
  dir_probs_matrix <- exp_sensitivity_analysis(exp_list = list)
  
  return(dir_probs_matrix)
}

# Create a list out of the different exposure calculations
dfs <- list(vuln_total_no_gapfill %>%,
             vuln_total_unwgt_region_gapfill,
             vuln_total_wgt_region_gapfill,
             vuln_total_unwgt_thorough_gapfill,
             vuln_total_wgt_thorough_gapfill,
             vuln_total_unwgt_diagonal_gapfill,
             vuln_total_wgt_diagonal_gapfill)

# Run sensitvity analysis
dir_probs_matrix <- run_sensitivity_analysis(dfs, comparison = "exposure")


# compute min r_kendall per combination
mins <- dir_probs_matrix %>%
  drop_na() %>%
  group_by(combination) %>%
  summarise(
    min_y = min(r_kendall, na.rm = TRUE) # x‑position of the min
  )

# Step 2: main plot
plot <- data.frame(dir_probs_matrix) %>%
  drop_na() %>%
  mutate(combination = factor(combination)) %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = combination)) +
  geom_line() +
  facet_wrap(~ combination, nrow = 6) +
  labs(
    y = "Kendall tau correlation",
    x = "Dirichlet-assigned weight for focal exposure method",
    title = "Sensitivity of exposure score"
  ) +
  theme_cowplot() +
  guides(color = "none") +
  theme(
    axis.text = element_text(size = 8),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +

  # Step 3: add the min labels
  geom_text(
    data = mins,
    aes(x = 0.6, y = 0.6, label = sprintf("Min: %.3f", min_y)),
    color = "black",
    size = 2.5,
    vjust = -0.5,
    inherit.aes = FALSE
  )


# Save plot
ggsave("../images/exploratory/sensitiity_analysis_exposure.jpg", plot = plot, device = "jpg", width = 10, height = 6.5, dpi = 300)
```

## Sensitivity analysis of sensitivity measure

```{r sensitivity measurement: sensitivity analysis}
# Run sensitivity analysis
dir_probs_matrix_sensitivity <- sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                                                             analysis = "sensitivity")

plot <- dir_probs_matrix_sensitivity %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var),
         var = case_when(var == "1" ~ "Foreign reliance (%)",
                         var == "2" ~ "Marine capture reliance (%)")) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal sensitivity variable", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_sensitivity.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```

## Sensitivity analysis on adaptive capacity variables

### Variables within adaptive capacity

```{r adaptive capacity measurement: sensitivity analysis}
dir_probs_matrix_ac_variables <- sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                                                             analysis = "adaptive_capacity_variables")

plot <- dir_probs_matrix_ac_variables %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var),
         var = case_when(var == "1" ~ "GDP per capita",
                         var == "2" ~ "Trade (% of GDP)",
                         var == "3" ~ "Sanitation",
                         var == "4" ~ "Supermarkets",
                         var == "5" ~ "Life expectancy",
                         var == "6" ~ "Proportion labor force",
                         var == "7" ~ "Human capacity index",
                         var == "8" ~ "Government effectiveness",
                         var == "9" ~ "Food safety capacity",
                         var == "10" ~ "Rule of law")) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
    theme(legend.position = "none",
          axis.text = element_text(size = 8),
          strip.text = element_text(size = 7),
          plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal adaptive capacity variable", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_adaptive_capacity_variables.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```

### Components within adaptive capacity

```{r}
dir_probs_matrix_ac_components <- 
  sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                               analysis = "adaptive_capacity_components")

plot <- dir_probs_matrix_ac_components %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var)) %>%
  mutate(var = case_when(
         var == "1" ~ "Assets",
         var == "2" ~ "Flexibility",
         var == "3" ~ "Social Organization",
         var == "4" ~ "Learning"),
         dirichlet_prob = case_when(
           var != "Learning" ~ dirichlet_prob * 3,
           TRUE ~ dirichlet_prob
         )) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
    theme(legend.position = "none",
          axis.text = element_text(size = 8),
          strip.text = element_text(size = 7),
          plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal adaptive capacity component", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_adaptive_capacity_components.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```

---

# Diagnostics — correlations between risk components

```{r risk-correlations}
pairs(vuln_total_no_gapfill[, 2:5])
pairs(vuln_total_region_gapfill[, 2:5])
pairs(vuln_total_realm_gapfill[, 2:5])

cor(vuln_total_no_gapfill[, 2:5])[1,] # row selected: exposure
cor(vuln_total_region_gapfill[, 2:5])[1,] # row selected: exposure
cor(vuln_total_realm_gapfill[, 2:5])[1,] # row selected: exposure

hist(e_s_ac_total_realm_gapfill$pct_change)
```

# Compare weighted vs unweighted interpolated data

```{r}
unweighted <- read_parquet("../output/future_catch_joined_consumption/interpolated_future_data_total_portfolio_unweighted_diagonal_gapfill.parquet")

weighted <- read_parquet("../output/future_catch_joined_consumption/interpolated_future_data_total_portfolio_weighted_diagonal_gapfill.parquet")

weighted %>% 
  group_by(sciname, eez_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(eez_iso3c) %>%
  mutate(prop_of_total = live_weight_t / sum(live_weight_t)) %>%
  View()

weighted %>% filter(class == "actinopterygii") %>% distinct(gapfill_level)

cor(vuln_total_wgt_diagonal_gapfill$exposure, vuln_total_unwgt_diagonal_gapfill$exposure, method = "kendall")


unweighted %>%
  filter(!gapfill_method == "Gapfill not needed") %>%
  distinct(eez_iso3c, sciname, live_weight_t, ssp126) %>%
  arrange(-live_weight_t)

weighted %>%
  filter(!gapfill_method == "Gapfill not needed") %>%
  distinct(eez_iso3c, sciname, live_weight_t, ssp126) %>%
  arrange(-live_weight_t)


sciname <- read.csv("../data/sciname.csv")

weighted %>%
  filter(sciname == "gadus chalcogrammus",
         eez_iso3c == "RUS") %>%
  distinct(gapfill_method, gapfill_level)
```

```{r}
russian_foods <- consumption_mpc %>%
  filter(genus == "gadus",
         eez_iso3c %in% "RUS")

df <- consumption_mpc %>%
  # add_region("eez_iso3c", "region") %>%
  # left_join(ecoregion %>% select(iso3c, realm), by = c("eez_iso3c" = "iso3c")) %>%
  filter(family == "synodontidae",
         eez_iso3c == "VNM") %>%
  select(-realm)

test %>% distinct(eez_iso3c)

# See which distinct species have non NAs (looking for more than 1 to test weighted averages)

japanese_foods %>%
  filter(!is.na(ssp126)) %>%
  distinct(sciname)



# Check which gapfilling level / method was used on target taxa
weighted %>%
  filter(sciname == "scomber",
         eez_iso3c == "JPN") %>%
  distinct(gapfill_method, gapfill_level)

read.csv("../output/future_catch_data_joined/future_catch_joined.csv") %>%
  filter(sciname == "scomber japonicus") %>% View()
```

