---
title: "00 — Preprocess Data"
author: "Connor Quiroz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(rfishbase)
```

FIXIT: Warning: There were 250 warnings in `mutate()` check out where those 250 warnings are coming from and see whether that is okay or not.

# Overview

This script defines four functions that build the exposure, sensitivity, and
adaptive capacity (E/S/AC) components used to calculate climate risk. Each
function handles one job:

| Function | Runs once? | Purpose |
|---|---|---|
| `preprocess_climate_data()` | ✅ Yes | Read DBEM files → interpolate missing taxa → write parquet |
| `build_exposure()` | Per portfolio | Join climate projections to consumption; compute % stock change |
| `build_sensitivity()` | Per portfolio | Compute aquatic animal reliance by portfolio type |
| `build_adaptive_capacity()` | Per portfolio | Load and join all AC indicators |

`build_e_s_ac_data()` orchestrates the three per-portfolio steps and writes the
final joined dataset to `../output/`.

---

# Step 1 — Catch proejection data preprocessing (run once)

Reads all DBEM species files, joins species names, converts EEZ strings to
ISO3C codes, and writes a single joined CSV. Because this step is
portfolio-agnostic it only needs to run once per session.

```{r preprocess-climate-data}
preprocess_climate_data <- function(dir) {

  file_names <- list.files(
    paste0(dir, "exposure/dbem_delta_mcp/"),
    pattern = "\\.csv",
    full.names = TRUE
  )

  species_names <- fread(paste0(dir, "exposure/dbem_spp_list.csv"))

  # Extract SeaAroundUS (Juliano's; SAU) data across folders and combine into
  # one file
  mcp_combined <- map_dfr(file_names, function(f) {
    df_i <- fread(f)
    if (nrow(df_i) == 0) return(NULL)

    df_i %>%
      left_join(species_names, by = "taxon_key") %>%
      mutate(taxon_name = str_to_lower(worms_name)) %>% 
      select(taxon_name, eez_name, ssp, mean_mcp_delta_2030) %>%
      pivot_wider(names_from = ssp, values_from = mean_mcp_delta_2030, values_fn = mean)
  })
  
  # Problem: SAUC catch projections contain information at
  # sub-sovereign EEZ level (for example, Russia (Kara Sea)) — we want
  # to aggregate EEZ projections across sub-eez names. To aggregate 
  # EEZ projections, we will take a weighted average of the projections
  # multiplied byt each sub-EEZ's production weight, another SAU data source
    
  # setDTthreads(1) # this was for faster file reading initially when using 
  # fread to load in the .csv, but we converted to .parquet
  
  # Read in production data by EEZ
  sau_prod <-read_parquet("../data/miscellaneous/sau_prod.parquet")
  
  # Calculate production totals by EEZ
  prod_totals <- sau_prod %>%
    group_by(eez) %>%
    summarize(total_production = sum(sum)) %>%
    rename(eez_name = eez) %>%
    # Correct specific cases to enable joins on catch projection data
    # Ask Juliano about Greece - his data names are "Greece" and "Crete (Greece)"
    # The SAU data only recognizes "Crete (Greece)" and "Greece (without Crete)"
    # Does your "Greece" not include Crete?
    mutate(eez_name = case_when(
      
      eez_name == "Greece (without Crete)" ~ "Greece",
      TRUE ~ eez_name
      
    ))
  
  # Clean up EEZ names
  mcp_clean <- mcp_combined %>%
    mutate(
      eez_iso3c = case_when(
        eez_name == "Chagos Archipelago (UK)"            ~ "GBR",
        eez_name == "Christmas Isl. (Australia)"         ~ "AUS",
        eez_name == "High seas"                          ~ "NEI",
        eez_name == "Iran (Sea of Oman)"                 ~ "IRN",
        eez_name == "Mozambique Channel Isl. (France)"   ~ "FRA",
        eez_name == "Mayotte (France)"                   ~ "FRA",
        eez_name == "New Caledonia (France)"             ~ "FRA",
        eez_name == "Norfolk Isl. (Australia)"           ~ "AUS",
        eez_name == "Réunion (France)"                   ~ "FRA",
        eez_name == "Saudi Arabia (Persian Gulf)"        ~ "SAU",
        eez_name == "Svalbard Isl. (Norway)"             ~ "NOR",
        eez_name == "Tokelau (New Zealand)"              ~ "NZL",
        eez_name == "Wallis & Futuna Isl. (France)"      ~ "FRA",
        eez_name == "Yemen (Arabian Sea)"                ~ "YEM",
        TRUE ~ countrycode(
          eez_name, origin = "country.name", destination = "iso3c"
          ))) %>%
    # Get rid of NEI (not elsewhere included EEZ's) - won't be used in analysis
    filter(eez_iso3c != "NEI")
  
  
  # Calculate the proportion of production for each sub-eez name within each
  # sovereign EEZ (used for weighted averaging)
  prod_props_eez <- mcp_clean %>%
    distinct(eez_name, eez_iso3c) %>%
    left_join(prod_totals, by = "eez_name") %>%
    # The join will result in NA catch projection values for Ecuador.
    # This doesn't matter though because Ecuador has no sub-eez names in the
    # catch projection data, so we do not need to weighted average any of its
    # values and thus will impute 1, otherwise leave the values unchanged
    mutate(total_production = case_when(
      eez_name == "Ecuador" ~ 1,
      TRUE ~ total_production
    )) %>%
    # Workflow for calculating weighted averages:
    # Calculate the proportion of production for each sub-eez within a
    # sovereign EEZ
    group_by(eez_iso3c) %>%
    mutate(prop_prod = total_production / sum(total_production)) %>%
    ungroup() %>%
    select(eez_name, prop_prod)
  
  # Join production proportions at sub-eez level onto catch projection data,
  # calculating the weighted average at the sovereign level - this will be the
  # data we use for exposure
  mcp_sovereign <- mcp_clean %>%
    left_join(prod_props_eez, by = "eez_name") %>%
    # Calculate weighted averages for catch projections for each sovereign EEZ
    mutate(ssp126 = ssp126 * prop_prod,
           ssp585 = ssp585 * prop_prod) %>%
    group_by(taxon_name, eez_iso3c) %>%
    summarize(ssp126 = sum(ssp126),
             ssp585 = sum(ssp585))
  
  write_csv(df, "../data/exposure/mcp_sovereign.csv")
  message("✅ mcp_sovereign.csv written to ../data/exposure")

  invisible(df)
}
```

---

# Step 2 — Exposure (per portfolio)

Joins climate projections to 2019 consumption data, fills missing species
values up through the taxonomic hierarchy, then computes each country's
weighted % change in available stock under SSP1-2.6 and SSP5-8.5.

**Portfolio differences:**

- `"foreign"` — keeps only rows where `consumer_iso3c != producer_iso3c` and
  applies a three-way case_when for the catch-change multiplier (handles
  positive, negative, and zero separately).
- `"total"` — retains the full consumption portfolio and uses a simpler
  `1 + (x / 100)` multiplier.

## Bayesian gap filling

### Perform Bayesian estimation on mcp measures to generate posteriors

Here we collect the data, format the data, and set starting values to be passsed
into Stan model for model fit. This function returns the Stan fit, which
posteriors will be extracted from in the following function.

```{r build-bayesian-lookup-table}
taxonomic_bayesian_gapfill <- function(df = df, 
                                  assumed_sd,
                                  method = c("eez_iso3c", "realm", "region"),
                                  ssp = c("ssp126", "ssp585")) {
 
  # Get distinct values of sciname / EEZ iso3c pairs with their respective ssp value
  df <- df %>%
    filter(!is.na(ssp126) | !is.na(ssp585)) %>%
    distinct(sciname, eez_iso3c, all_of(across(ssp)))
  
  # Average ssp values by sciname / realm pair
  if (method == "realm") {
    df <- df %>%
    left_join(
        ecoregion %>% select(iso3c, realm),
        by = c("eez_iso3c" = "iso3c")
      ) %>%
      group_by(sciname, realm) %>%
      summarize(!!sym(ssp) := mean(!!sym(ssp)))
  }
  
  # Average ssp values by sciname / region pair
  if (method == "region") {
    df <- df %>%
    add_region("eez_iso3c", "region") %>%
      group_by(sciname, region) %>%
      summarize(!!sym(ssp) := mean(!!sym(ssp)))
  }
  
  # Add taxa levels back to data
  df <- df %>%
    left_join(sciname_taxa, by = "sciname") %>%
    select(-species, -superclass, -subfamily, -taxa_level) %>%
    mutate(
        genus = case_when(
          sciname == "crassostrea gigas" ~ "crassostrea", TRUE ~ genus
        ),
        family = case_when(
          sciname == "crassostrea gigas" ~ "ostreidae", TRUE ~ family
        ),
        order = case_when(
          sciname == "crassostrea gigas" ~ "ostreidae", TRUE ~ order
        ),
        class = case_when(
          sciname == "crassostrea gigas" ~ "bivalvia", TRUE ~ class
        ),
        phylum = case_when(
          sciname == "crassostrea gigas" ~ "mollusca", 
          sciname == "petromyzon marinus" ~ "chordata",
          TRUE ~ phylum,
        ),
        kingdom = case_when(
          sciname == "crassostrea gigas" ~ "animalia", TRUE ~ kingdom
        )
      )
  
  ##################### KINGDOM #####################
  # Get the unique EEZ's used in each kingdom
  kingdom_eezs <- df %>%
    distinct(!!sym(method)) %>%
    pull(!!sym(method))
  
  # Get the unique kingdoms used
  kingdom <- df %>%
    distinct(kingdom) %>%
    pull(kingdom)
  
  # Get the all the kingdoms that correspond to the geographic grouping
  kingdoms <- df %>%
    distinct(!!sym(method), kingdom) %>%
    pull(kingdom)
  
  num_eezs <- length(kingdom_eezs) # Calculate the number of EEZs
  num_kingdom <- length(kingdom) 
  
  kingdom_combination_nrow <- num_eezs * num_kingdom
  
  eez_kingdom_lookup <- rep(NA_character_, kingdom_combination_nrow)
  kingdom_name_lookup <- rep(NA_character_, kingdom_combination_nrow)
  
  loop_iteration_nrow_kingdom <- 1
  
  for (eez in kingdom_eezs) {
    for (k in kingdom) {
      eez_kingdom_lookup[loop_iteration_nrow_kingdom] <- eez
      kingdom_name_lookup[loop_iteration_nrow_kingdom] <- k
      loop_iteration_nrow_kingdom <- loop_iteration_nrow_kingdom + 1
    }
  }
  
  ##################### PHYLUM #####################
  phylum_eezs <- df %>%
    
    distinct(!!sym(method), kingdom, phylum) %>%
    pull(!!sym(method))
  
  phylum_kingdom <- df %>%
    
    distinct(!!sym(method), kingdom, phylum) %>%
    pull(kingdom)
  
  phylum <- df %>%
    
    distinct(!!sym(method), kingdom, phylum) %>%
    pull(phylum)
  
  phylum_combination_nrow <- length(phylum)
  
  phylum_to_kingdom_lookup <- rep(NA_integer_, phylum_combination_nrow)
  
  for(loop in 1:phylum_combination_nrow){
    idx <- which(
      phylum_eezs[loop] == eez_kingdom_lookup &
      phylum_kingdom[loop] == kingdom_name_lookup
    )
    phylum_to_kingdom_lookup[loop] <- idx[1]
  }
  
  ##################### CLASS #####################
  class_eezs <- df %>%
    
    distinct(!!sym(method), phylum, class) %>%
    pull(!!sym(method))
  
  class_phylum <- df %>%
    
    distinct(!!sym(method), phylum, class) %>%
    pull(phylum)
  
  class <- df %>%
    
    distinct(!!sym(method), phylum, class) %>%
    pull(class)
  
  class_combination_nrow <- length(class)
  
  class_to_phylum_lookup <- rep(NA_integer_, class_combination_nrow)
  
  for(loop in 1:class_combination_nrow){
    idx <- which(
      class_eezs[loop] == phylum_eezs &
      class_phylum[loop] == phylum
    )
    class_to_phylum_lookup[loop] <- idx[1]
  }
  
  ##################### ORDER #####################
  order_eezs <- df %>%
    
    distinct(!!sym(method), class, order) %>%
    pull(!!sym(method))
  
  order_class <- df %>%
    
    distinct(!!sym(method), class, order) %>%
    pull(class)
  
  order <- df %>%
    
    distinct(!!sym(method), class, order) %>%
    pull(order)
  
  order_combination_nrow <- length(order)
  
  order_to_class_lookup <- rep(NA_integer_, order_combination_nrow)
  
  for(loop in 1:order_combination_nrow){
    idx <- which(
      order_eezs[loop] == class_eezs &
      order_class[loop] == class
    )
    order_to_class_lookup[loop] <- idx[1]
  }
  
  ##################### FAMILY #####################
  family_eezs <- df %>%
    
    distinct(!!sym(method), order, family) %>%
    pull(!!sym(method))
  
  family_order <- df %>%
    
    distinct(!!sym(method), order, family) %>%
    pull(order)
  
  family <- df %>%
    
    distinct(!!sym(method), order, family) %>%
    pull(family)
  
  family_combination_nrow <- length(family)
  
  family_to_order_lookup <- rep(NA_integer_, family_combination_nrow)
  
  for(loop in 1:family_combination_nrow){
    idx <- which(
      family_eezs[loop] == order_eezs &
      family_order[loop] == order
    )
    family_to_order_lookup[loop] <- idx[1]
  }
  
  ##################### GENUS #####################
  genus_eezs <- df %>%
    
    distinct(!!sym(method), family, genus) %>%
    pull(!!sym(method))
  
  genus_family <- df %>%
    distinct(!!sym(method), family, genus) %>%
    pull(family)
  
  genus <- df %>%
    distinct(!!sym(method), family, genus) %>%
    pull(genus)
  
  genus_combination_nrow <- length(genus)
  
  genus_to_family_lookup <- rep(NA_integer_, genus_combination_nrow)
  
  for(loop in 1:genus_combination_nrow){
    idx <- which(
      genus_eezs[loop] == family_eezs &
      genus_family[loop] == family
    )
    genus_to_family_lookup[loop] <- idx[1]
  }
  
  ##################### SPECIES #####################
  species_eezs <- df %>%
    distinct(!!sym(method), genus, sciname) %>%
    pull(!!sym(method))
  
  species_genus <- df %>%
    distinct(!!sym(method), genus, sciname) %>%
    pull(genus)
  
  species <- df %>%
    distinct(!!sym(method), genus, sciname) %>%
    pull(sciname)
  
  species_combination_nrow <- length(species)
  
  species_to_genus_lookup <- rep(NA_integer_, species_combination_nrow)
  
  for(loop in 1:species_combination_nrow){
    idx <- which(
      species_eezs[loop] == genus_eezs &
      species_genus[loop] == genus
    )
    species_to_genus_lookup[loop] <- idx[1]
  }
  
############################## Set input data for Stan model
  ssp_obs <- df %>%
    filter(!is.na(!!sym(ssp))) %>%
    distinct(
      across(all_of(method)),
      across(all_of(ssp)),
      genus,
      species,
      ) %>%
  pull(!!sym(ssp))

  
# Test to ensure number of observations matches the number of predicted columns of the lookup table
stopifnot(
  length(ssp_obs) ==
    species_combination_nrow
)

# Create list of data to be passed into Stan
stan_data <- list(

  kingdom_combination_nrow =
    kingdom_combination_nrow,

  phylum_combination_nrow =
    phylum_combination_nrow,

  class_combination_nrow =
    class_combination_nrow,

  order_combination_nrow =
    order_combination_nrow,

  family_combination_nrow =
    family_combination_nrow,

  genus_combination_nrow =
    genus_combination_nrow,

  species_combination_nrow =
    species_combination_nrow,

  phylum_to_kingdom_lookup =
    as.integer(phylum_to_kingdom_lookup),

  class_to_phylum_lookup =
    as.integer(class_to_phylum_lookup),

  order_to_class_lookup =
    as.integer(order_to_class_lookup),

  family_to_order_lookup =
    as.integer(family_to_order_lookup),

  genus_to_family_lookup =
    as.integer(genus_to_family_lookup),

  species_to_genus_lookup =
    as.integer(species_to_genus_lookup),

  ssp_obs = as.vector(ssp_obs),
  
  sigma_obs = assumed_sd

)
  
############################## Set inits for stan model
  init_fun <- function() {

  species_init <- ssp_obs

  genus_init <- rnorm(
    genus_combination_nrow,
    mean(species_init),
    5
  )

  family_init <- rnorm(
    family_combination_nrow,
    mean(genus_init),
    5
  )

  order_init <- rnorm(
    order_combination_nrow,
    mean(family_init),
    5
  )

  class_init <- rnorm(
    class_combination_nrow,
    mean(order_init),
    5
  )

  phylum_init <- rnorm(
    phylum_combination_nrow,
    mean(class_init),
    5
  )

  kingdom_init <- runif(
    kingdom_combination_nrow,
    -25,
    25  
  )

  list(

    ssp_pred_kingdom_lookup =
      kingdom_init,

    ssp_pred_phylum_lookup =
      phylum_init,

    ssp_pred_class_lookup =
      class_init,

    ssp_pred_order_lookup =
      order_init,

    ssp_pred_family_lookup =
      family_init,

    ssp_pred_genus_lookup =
      genus_init,

    ssp_pred_species_lookup =
      species_init,
    
    sigma_phylum = 5,
    
    sigma_class = 7,
    
    sigma_order = 8,
    
    sigma_family = 10,
    
    sigma_genus = 12,
    
    sigma_species = 15

  )

  }
  
  
  

  fit <- rstan::stan(
    file = "hierarchical_ssp.stan",
    data = stan_data,
    init = init_fun,
    chains = 3,
    iter = 2000, 
    warmup = 1000,
    cores = 3,
    refresh = 100)
  
  return(list(
  fit           = fit,
  genus_eezs    = genus_eezs,
  family_eezs   = family_eezs,
  order_eezs    = order_eezs,
  class_eezs    = class_eezs,
  phylum_eezs   = phylum_eezs,
  kingdom_eezs  = kingdom_eezs,
  genus         = genus,
  family        = family,
  order         = order,
  class         = class,
  phylum        = phylum,
  kingdom       = kingdoms
))
   
}
```

### Create Bayesian lookup table

```{r}
build_bayesian_lookup <-function(df, ecoregion, assumed_sd) {

  cat(sprintf("Building Bayesian lookup table...\n"))

# ------------------------------- EEZ

## SSP 126

### Run bayesian model DO NOT OPEN OUTPUT FIT OR R WILL CRASH
output <- taxonomic_bayesian_gapfill(df = df, assumed_sd = assumed_sd,
                                     method = "eez_iso3c", ssp = "ssp126")

fit_eez_iso3c_126 <- output$fit

### Extract mcmc samples from EEZ fit
mcmc_eez_126 <- rstan::extract(fit_eez_iso3c_126, permuted = TRUE)

saveRDS(mcmc_eez_126, paste0("../output/bayesian/",as.character(assumed_sd),"_sd","/mcmc_ssp126.rds"))

### calculate median posteriors for each taxa
genus_eez_med_126 <- apply(mcmc_eez_126$ssp_pred_genus_lookup, MARGIN = 2, FUN = median)
family_eez_med_126 <- apply(mcmc_eez_126$ssp_pred_family_lookup, MARGIN = 2, FUN = median) 
order_eez_med_126 <- apply(mcmc_eez_126$ssp_pred_order_lookup, MARGIN = 2, FUN = median) 
class_eez_med_126 <- apply(mcmc_eez_126$ssp_pred_class_lookup, MARGIN = 2, FUN = median) 
phylum_eez_med_126 <- apply(mcmc_eez_126$ssp_pred_phylum_lookup, MARGIN = 2, FUN = median) 
kingdom_eez_med_126 <- apply(mcmc_eez_126$ssp_pred_kingdom_lookup, MARGIN = 2, FUN = median)

# Refersh environment
rm(output)

## SSP 585

### Run bayesian modelDO NOT OPEN OUTPUT FIT OR R WILL CRASH
output <- taxonomic_bayesian_gapfill(df = df, assumed_sd = assumed_sd,
                                                method = "eez_iso3c", ssp = "ssp585")

fit_eez_iso3c_585 <- output$fit

### Extract mcmc samples from EEZ fit
mcmc_eez_585 <- rstan::extract(fit_eez_iso3c_585, permuted = TRUE)

saveRDS(mcmc_eez_126, paste0("../output/bayesian/",as.character(assumed_sd),"_sd","/mcmc_ssp585.rds"))


### calculate median posteriors for each taxa
genus_eez_med_585 <- apply(mcmc_eez_585$ssp_pred_genus_lookup, MARGIN = 2, FUN = median)
family_eez_med_585 <- apply(mcmc_eez_585$ssp_pred_family_lookup, MARGIN = 2, FUN = median) 
order_eez_med_585 <- apply(mcmc_eez_585$ssp_pred_order_lookup, MARGIN = 2, FUN = median) 
class_eez_med_585 <- apply(mcmc_eez_585$ssp_pred_class_lookup, MARGIN = 2, FUN = median) 
phylum_eez_med_585 <- apply(mcmc_eez_585$ssp_pred_phylum_lookup, MARGIN = 2, FUN = median) 
kingdom_eez_med_585 <- apply(mcmc_eez_585$ssp_pred_kingdom_lookup, MARGIN = 2, FUN = median) 



## Create EEZ lookup table
bayesian_lookup_eez_unwgt <- data.frame(eez_iso3c = c(output$genus_eezs, output$family_eezs, output$order_eezs, output$class_eezs, output$phylum_eezs, output$kingdom_eezs),
           taxa_name = c(output$genus, output$family, output$order, 
                         output$class, output$phylum, output$kingdom),
           eez_avg_ssp126 = c(genus_eez_med_126, family_eez_med_126,
                              order_eez_med_126, class_eez_med_126,
                              phylum_eez_med_126, kingdom_eez_med_126),
           eez_avg_ssp585 = c(genus_eez_med_585, family_eez_med_585,
                              order_eez_med_585, class_eez_med_585,
                              phylum_eez_med_585, kingdom_eez_med_585),
           taxa_level = c(rep("genus", length(output$genus_eezs)),
                          rep("family", length(output$family_eezs)),
                          rep("order", length(output$order_eezs)),
                          rep("class", length(output$class_eezs)),
                          rep("phylum", length(output$phylum_eezs)),
                          rep("kingdom", length(output$kingdom_eezs)))) %>%
  # FIXIT: quick fix, figure out why ostreidae is having duplicates
  mutate(taxa_level = case_when(
    taxa_name == "ostreidae" ~ "family",
    TRUE ~ taxa_level
  )) %>%
  group_by(eez_iso3c, taxa_name, taxa_level) %>%
  summarize(eez_avg_ssp126 = mean(eez_avg_ssp126),
            eez_avg_ssp585 = mean(eez_avg_ssp585)) %>%
  mutate(eez_cv_ssp126 = NA,
         eez_cv_ssp585 = NA) %>%
  relocate(taxa_level, .after = eez_cv_ssp585) %>%
  ungroup()

# ------------------------------- Realm

bayesian_lookup_realm_unwgt <- bayesian_lookup_eez_unwgt %>%
  left_join(
    ecoregion %>%
      select(iso3c, realm), 
    by = c("eez_iso3c" = "iso3c")) %>%
  select(-eez_iso3c) %>%
  group_by(realm, taxa_name) %>%
  summarize(realm_avg_ssp126 = mean(eez_avg_ssp126),
         realm_avg_ssp585 = mean(eez_avg_ssp585),
         realm_cv_ssp126 = NA,
         realm_cv_ssp585 = NA) %>%
  ungroup() %>%
  distinct() %>%
  relocate(realm, .before = "taxa_name") %>%
  # Add on taxa level
  left_join(
    bayesian_lookup_eez_unwgt %>% select(taxa_name, taxa_level) %>% distinct(),
    by = "taxa_name")



# ------------------------------- Region
bayesian_lookup_region_unwgt <- bayesian_lookup_eez_unwgt %>%
  exploreARTIS::add_region(col = "eez_iso3c", region.col.name = "region") %>%
  group_by(region, taxa_name) %>%
  summarize(region_avg_ssp126 = mean(eez_avg_ssp126),
         region_avg_ssp585 = mean(eez_avg_ssp585),
         region_cv_ssp126 = NA,
         region_cv_ssp585 = NA) %>%
  ungroup() %>%
  distinct() %>%
  # Add on taxa level
  left_join(
    bayesian_lookup_eez_unwgt %>% select(taxa_name, taxa_level)  %>% distinct(),
    by = "taxa_name")

# ------------------------------- Global
bayesian_lookup_global_unwgt <- bayesian_lookup_eez_unwgt %>%
  exploreARTIS::add_region(col = "eez_iso3c", region.col.name = "region") %>%
  group_by(taxa_name) %>%
  summarize(global_avg_ssp126 = mean(eez_avg_ssp126),
         global_avg_ssp585 = mean(eez_avg_ssp585),
         global_cv_ssp126 = NA,
         global_cv_ssp585 = NA) %>%
  ungroup() %>%
  distinct() %>%
  # Add on taxa level
  left_join(
    bayesian_lookup_eez_unwgt %>% select(taxa_name, taxa_level)  %>% distinct(),
    by = "taxa_name")

return(list(
  eez_gapfills_unwgt      = bayesian_lookup_eez_unwgt,
  realm_gapfills_unwgt    = bayesian_lookup_realm_unwgt,
  region_gapfills_unwgt   = bayesian_lookup_region_unwgt,
  global_gapfills_unwgt   = bayesian_lookup_global_unwgt
))

}
```


## Build taxa levels

```{r build-taxa-levels}
# Helper: classify each sciname to its taxonomic level and extract species epithet
build_taxa_levels <- function(sciname_df) {
  sciname_df %>%
    mutate(
      taxa_level = case_when(
        sciname == kingdom   ~ "kingdom",
        sciname == phylum    ~ "phylum",
        sciname == superclass ~ "superclass",
        sciname == class     ~ "class",
        sciname == order     ~ "order",
        sciname == family    ~ "family",
        sciname == subfamily ~ "subfamily",
        sciname == genus     ~ "genus",
        str_detect(sciname, " ") ~ "species"
      ),
      species = if_else(str_count(sciname, " ") == 1, word(sciname, 2), NA_character_)
    ) %>%
    relocate(species, .before = genus) %>%
    select(-common_name, -isscaap)
}
```

## Build MCP averages (for frequentist gap filling)

```{r build-MCP-averages}
# Helper: for each higher taxonomic level, compute mean SSP values per EEZ
# so that species with missing projections can borrow from relatives
generate_mcp_averages <- function(df,
                                  method = c("eez", "realm", "region", "global"),
                                  averaging = c("unweighted", "weighted"),
                                  ecoregion,
                                  num_species) {
  
  method <- match.arg(method)
  averaging <- match.arg(averaging)
  tax_levels <- c("genus", "family", "order", "class", "phylum", "kingdom")
  
  x <- map_dfr(tax_levels, function(level) {
    
    # Based on function input, group by function method input
    group_cols <- switch(
      method,
      eez   = c("eez_iso3c", level),
      realm = c("realm", level),
      region = c("region", level),
      global = c(level)
    )
    
    if (averaging == "weighted") { # Compute weighted avearages
      
      df2 <- df %>%
        filter(!is.na(ssp126) | !is.na(ssp585)) %>%
        group_by(sciname, eez_iso3c) %>%
        summarize(live_weight_t = sum(live_weight_t)) %>%
        ungroup() %>%
        left_join(
          ecoregion %>% select(iso3c, realm),
          by = c("eez_iso3c" = "iso3c")
        ) %>%
        add_region("eez_iso3c", "region")
      
      # Join ssp estimates back onto df dataset
      df2 <- df2 %>%
        left_join(
          df %>%
            add_region("eez_iso3c", "region") %>%
            left_join(ecoregion, by = c("eez_iso3c" = "iso3c")) %>%
            filter(!is.na(ssp126)) %>%
            distinct(sciname, across(all_of(c("eez_iso3c", group_cols[length(group_cols)]))), ssp126, ssp585),
          by = c("sciname", "eez_iso3c")
        )
      
      # Allow averages to only include at least X species (so that one species doesn't dominate).
      # Number of minimum species used for gapfilling depends on user input
      df2 <- df2 %>%
        group_by(across(all_of(c(group_cols[1], level)))) %>%
        mutate(n = n_distinct(sciname)) %>%   # Fixit: Currently set to minimum number of species. Do we want this to be minimum number of eez / iso3c pairs? That would be much easier to satisfy.
        ungroup() %>%
        filter(n >= num_species)

      
      out <- df2 %>%
        group_by(across(all_of(group_cols))) %>%
        mutate(total_live_weight_group = sum(live_weight_t)) %>%
        
        summarize(
          # Weighted averages
          
          !!paste0(method, "_avg_ssp126") := weighted.mean(ssp126, live_weight_t, na.rm = TRUE),
          !!paste0(method, "_avg_ssp585") := sum(ssp585 * (live_weight_t / total_live_weight_group)),

          # Weighted Coefficients of variation
          !!paste0(method, "_cv_ssp126") := {
            
            wm <- weighted.mean(ssp126, live_weight_t, na.rm = TRUE)

            wvar <- sum(live_weight_t * (ssp126 - wm)^2, na.rm = TRUE) /
              sum(live_weight_t, na.rm = TRUE)

            sqrt(wvar) / wm
            
          },
          
          !!paste0(method, "_cv_ssp585") := {
            
            wm <- weighted.mean(ssp585, live_weight_t, na.rm = TRUE)

            wvar <- sum(live_weight_t * (ssp585 - wm)^2, na.rm = TRUE) /
              sum(live_weight_t, na.rm = TRUE)

            sqrt(wvar) / wm
            
          }
        ) %>%
        ungroup()
      
    } else { # Compute unweighted averages
      
      df2 <- df %>%
      distinct() %>%
      left_join(
        ecoregion %>% select(iso3c, realm),
        by = c("eez_iso3c" = "iso3c")
      ) %>%
      filter(!is.na(ssp126) | !is.na(ssp585)) %>%
      add_region("eez_iso3c", "region") %>%
      distinct(eez_iso3c, sciname, .keep_all = TRUE)
      
      
      # Allow averages to only include at least 2 species (so that one species doesn't dominate)
      df2 <- df2 %>%
        group_by(across(all_of(c(group_cols[1], level)))) %>%
        mutate(n = n()) %>%   # number of rows in each pair
        ungroup() %>%
        filter(n >= num_species)
      
      out <- df2 %>%
        group_by(across(all_of(group_cols))) %>%
        summarise(
          !!paste0(method, "_avg_ssp126") := mean(ssp126, na.rm = TRUE),
          !!paste0(method, "_avg_ssp585") := mean(ssp585, na.rm = TRUE),
          
          !!paste0(method, "_cv_ssp126") :=
            sd(ssp126, na.rm = TRUE) / mean(ssp126, na.rm = TRUE),
          
          !!paste0(method, "_cv_ssp585") :=
            sd(ssp585, na.rm = TRUE) / mean(ssp585, na.rm = TRUE)
        ) %>%
        ungroup()
      
    }
    
    out %>%
      rename(taxa_name = all_of(level)) %>%
      mutate(taxa_level = level)
    
  })
  
  return(x)
  
}
```

## Build taxa lookup (for frequentist gap filling)

```{r build-taxa-lookup-frequentist}
# Helper: compile averages across all gap filling combinations into a list
# that will be used to gap fill NA MCP projections in consumption data
build_taxa_lookup <- function(df, ecoregion, num_species) {
  tax_levels <- c("genus", "family", "order", "class", "phylum", "kingdom")

  # For each taxa in an EEZ, compute the average ssp126 / 585 value for all species within in the taxa
  
  # ---------------------------------------------------------------------------
  # EEZ-level averages
  # ---------------------------------------------------------------------------
  
  eez_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "eez",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  eez_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "eez",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)

  # ---------------------------------------------------------------------------
  # Realm-level averages
  # ---------------------------------------------------------------------------
  
  realm_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "realm",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  realm_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "realm",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  # ---------------------------------------------------------------------------
  # Region-level averages
  # ---------------------------------------------------------------------------
  region_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "region",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  region_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "region",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  # ---------------------------------------------------------------------------
  # Global-level averages
  # ---------------------------------------------------------------------------
  
  global_gapfills_unwgt <- generate_mcp_averages(df = df,
                                              method = "global",
                                              averaging = "unweighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  global_gapfills_wgt <- generate_mcp_averages(df = df,
                                              method = "global",
                                              averaging = "weighted",
                                              ecoregion = ecoregion,
                                              num_species = num_species)
  
  return(list(
    eez_gapfills_unwgt = eez_gapfills_unwgt,
    eez_gapfills_wgt = eez_gapfills_wgt,
    region_gapfills_unwgt = region_gapfills_unwgt,
    region_gapfills_wgt = region_gapfills_wgt,
    realm_gapfills_unwgt = realm_gapfills_unwgt,
    realm_gapfills_wgt = realm_gapfills_wgt,
    global_gapfills_unwgt = global_gapfills_unwgt,
    global_gapfills_wgt = global_gapfills_wgt
  ))
      
}
```

## Helper for gap filling values from lookup table

```{r gap-filling-helper-from-lookup-table}
# Helper: walk up the taxonomic ladder and fill NA SSP values from the nearest
# available ancestor within the same EEZ
fill_higher_taxa_helper <- function(df,
                                    gap_filling_method,
                                    averaging,
                                    ecoregion,
                                    lookup_table,
                                    level,
                                    join_by_map) {
  
  suffix <- if (averaging == "weighted") "wgt" else "unwgt"
  selector <- paste0(gap_filling_method,"_gapfills_",suffix)
  
  if (level == "genus") {
    
    if (gap_filling_method == "realm") {
      df <- df %>%
        left_join(ecoregion %>% select(iso3c, realm), by = c("eez_iso3c" = "iso3c"))
    } else if (gap_filling_method == "region") {
      df <- df %>%
        add_region("eez_iso3c", "region")
    }
    
  }
  
  # Based on function input, group by function method input
  join_cols <- switch(
    gap_filling_method,
    eez   = c("eez_iso3c", join_by_map),
    realm = c("realm", join_by_map),
    region = c("region", join_by_map),
    global = c(join_by_map)
  )
  
   df <- df %>%
    select(-taxa_level) %>%
    ## Join data to EEZ interpolated data
    left_join(
      lookup_table[[selector]] %>% filter(taxa_level == level),
      by = join_cols) %>%
    ## Conduct EEZ gapfills
    ### Only assign the gapfill method if there was actually a gapfill occurring
    ### Only assign the taxonomic level for which the gapfill was performed at
    ### if a gapfill occurred
    mutate(
      gapfill_level = case_when( ## Define at which taxa level gap filled data was gotten from
        is.na(ssp126) & !is.na(!!sym(paste0(gap_filling_method,"_avg_ssp126"))) ~ level, # Data coverage between ssp126/585 are the same, so we only need to check one variable
        TRUE                                              ~ gapfill_level
      ),
      gapfill_method = case_when(
        is.na(ssp126) & !is.na(!!sym(paste0(gap_filling_method,"_avg_ssp126"))) ~ paste0(gap_filling_method,"_gapfill"),
        TRUE                                              ~ gapfill_method
      ),
      ssp126 = coalesce(ssp126, !!sym(paste0(gap_filling_method,"_avg_ssp126"))),
      ssp585 = coalesce(ssp585, !!sym(paste0(gap_filling_method,"_avg_ssp585"))),
      cv_ssp126 = coalesce(cv_ssp126, !!sym(paste0(gap_filling_method,"_cv_ssp126"))),
      cv_ssp585 = coalesce(cv_ssp585, !!sym(paste0(gap_filling_method,"_cv_ssp585")))
    ) %>%
    select(-!!sym(paste0(gap_filling_method,"_avg_ssp126")),
           -!!sym(paste0(gap_filling_method,"_avg_ssp585")),
           -!!sym(paste0(gap_filling_method,"_cv_ssp126")),
           -!!sym(paste0(gap_filling_method,"_cv_ssp585")))
  
  return(df)
  
}
```

## Perform gap filling

```{r perform-gap-filling}
# Helper: conduct gap filling based on selected method
fill_with_higher_taxa <- function(df,
                                  gap_filling_method,
                                  averaging,
                                  ecoregion,
                                  num_species,
                                  assumed_sd) {
  
  if (gap_filling_method != "bayesian") {
    if(!file.exists(paste0("../data/exposure/lookup_tables/",num_species,"_species/lookup_table.rds"))) {
    # Create directory if doesn't already exist
    if (!dir.exists(paste0("../data/exposure/lookup_tables/",num_species,"_species/"))) {
      dir.create(paste0("../data/exposure/lookup_tables/",num_species,"_species/"))
    }
    # Build lookup_table
    lookup_table <- build_taxa_lookup(df = df, ecoregion = ecoregion, num_species = num_species)
    saveRDS(lookup_table, paste0("../data/exposure/lookup_tables/",num_species,"_species/lookup_table.rds")) # Save lookup table
  } else {
    lookup_table <- readRDS(paste0("../data/exposure/lookup_tables/",num_species,"_species/lookup_table.rds")) 
  }
  } else {
    if (!file.exists("../data/exposure/lookup_tables/bayesian/lookup_table.rds")) {
      # FIXIT: Add an automatically creating directory, instead of me putting it in manually
      lookup_table <- build_bayesian_lookup(df = df, ecoregion, assumed_sd = assumed_sd)
      saveRDS(lookup_table, paste0("../data/exposure/lookup_tables/bayesian/",as.character(assumed_sd),"_sd/","lookup_table.rds"))
    } else { 
      lookup_table <- readRDS(paste0("../data/exposure/lookup_tables/bayesian/",as.character(assumed_sd),"_sd/","lookup_table.rds"))
    }
  }

  cat(sprintf("✔️ Lookup table loaded in\n"))
  
  # Set taxonomic levels to iterate through
  tax_levels <- c("genus", "family", "order", "class", "phylum", "kingdom")

  # Create empty variables that will be filled as gap filling happens
  out_df <- df %>% mutate(gapfill_level = NA_character_,
                      gapfill_method = NA_character_,
                      cv_ssp126 = NA,
                      cv_ssp585 = NA)

  # ---------------------------------------------------------------------------
  # EEZ-level gap filling
  # ---------------------------------------------------------------------------
  if ((gap_filling_method != "hierarchical" & gap_filling_method != "diagonal" & gap_filling_method != "bayesian")) {
   
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level)
      
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "eez",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    
    }
  
  # ---------------------------------------------------------------------------
  # Realm-level gap filling
  # ---------------------------------------------------------------------------
  
  if (gap_filling_method == "realm") {
    
    for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level) 
    
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "realm",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    }
    
  }
    
    
  # ---------------------------------------------------------------------------
  # Region-level gap filling
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "region") {
    
    for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level) 
    
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "region",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    }
    
  }
    
  # ---------------------------------------------------------------------------
  # Realm --> Region --> Global gapfilling 
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "thorough") {
    
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level) 
      
      # 1. Attempt realm-level gap fills
      out_df <- fill_higher_taxa_helper(out_df,
                                    gap_filling_method = "realm",
                                    averaging = averaging,
                                    lookup_table = lookup_table,
                                    ecoregion = ecoregion,
                                    level = level,
                                    join_by_map = join_by_map)
      
    }
    
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level) 
      
      # 2. Attempt region-level gap fills
      out_df <- fill_higher_taxa_helper(out_df,
                                    gap_filling_method = "region",
                                    averaging = averaging,
                                    lookup_table = lookup_table,
                                    ecoregion = ecoregion,
                                    level = level,
                                    join_by_map = join_by_map)
      
    }
    
    for (level in tax_levels) {
      
      join_by_map <- setNames("taxa_name", level) 
      
      # 2. Attempt region-level gap fills
      out_df <- fill_higher_taxa_helper(out_df,
                                    gap_filling_method = "global",
                                    averaging = averaging,
                                    lookup_table = lookup_table,
                                    ecoregion = ecoregion,
                                    level = level,
                                    join_by_map = join_by_map)
      
    }
  
  }
    
  }

  # ---------------------------------------------------------------------------
  # Hierarchical gapfilling 
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "hierarchical") {
    
    for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level)

    # 1. Attempt EEZ-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "eez",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)

    # 2. Attempt realm-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "realm",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)

    # 3. Attempt region-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "region",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)

    # 4. Attempt global-level gap fills
    out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = "global",
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = level,
                                  join_by_map = join_by_map)
    }
    
  }
  

  
  # ---------------------------------------------------------------------------
  # Diagonal gapfilling (frequentist & Bayesian)
  # ---------------------------------------------------------------------------
  if (gap_filling_method == "diagonal" | gap_filling_method == "bayesian") {
    
    # 1. Initialize the 7x4 matrix
    pattern <- matrix(data = c(1,3,6,10,2,5,9,14,4,8,13,18,7,12,
                               17,21,11,16,20,23,15,19,22,24), nrow = 6, ncol = 4,
                      byrow = TRUE)
    
    colnames(pattern) <- c("eez", "realm", "region", "global")
    rownames(pattern) <- c("genus", "family", "order", "class", "phylum", "kingdom")
    
    
    lookup <- expand.grid(row = rownames(pattern), col = colnames(pattern))
    lookup$value <- as.vector(pattern)
    
    lookup <- lookup[order(lookup$value), ]
    
    lookup$col <- as.character(lookup$col)
    lookup$row <- as.character(lookup$row)
    
    for (i in 1:nrow(lookup)) {
      
      # Set join by map key (e.g., join by EEZ & genus, EEZ & family, realm & phylum, etc)
      join_by_map <- setNames("taxa_name", lookup[,1][i])
      
      out_df <- fill_higher_taxa_helper(out_df,
                                  gap_filling_method = lookup$col[lookup$value == i],
                                  averaging = averaging,
                                  lookup_table = lookup_table,
                                  ecoregion = ecoregion,
                                  level = lookup$row[lookup$value == i],
                                  join_by_map = join_by_map)
      
    }
    
  }

  # Post gapfilling, delineate whether the gapfilling
  # was not needed or was unable to be performed.
  # Gapfill levels / methods are otherwise left unchanged.
  
  out_df <- out_df %>%
    mutate(gapfill_method = case_when(
      !is.na(ssp126) & is.na(gapfill_method) ~ "Gapfill not needed",
      is.na(ssp126) ~ "Unsuccessful gapfill",
      TRUE ~ gapfill_method 
    ),
    gapfill_level = case_when(
      !is.na(ssp126) & is.na(gapfill_level) ~ "Gapfill not needed",
      is.na(ssp126) ~ "Unsuccessful gapfill",
      TRUE ~ gapfill_level 
    ),
    averaging = case_when(!(gapfill_method == "Gapfill not needed" |
                          gapfill_method == "Unsuccessful gapfill") ~ averaging,
                          TRUE ~ NA_character_) # define whether weighted or unweighted averaging was used.
    ) %>% 
    select(-taxa_level) # not needed in final output
  
  return(out_df)
}
```

## Calculate exposure

```{r build-exposure}
build_exposure <- function(dir, portfolio, gap_filling_method, averaging, consumption, num_species, assumed_sd) {
  # Load in file names and species names acquired from the sau mcp data
  file_names <- readRDS("../data/exposure/utils/file_names.rds")
  species_names <- readRDS("../data/exposure/utils/species_names.rds")
  
  ## Read in Juliano's mcp data
  climate_by_spp_eez <- read_csv("../output/future_catch_data_joined/future_catch_joined.csv", show_col_types = FALSE)
  
  # File integrity diagnostics of Juliano's climate data
  
  filename_keys <- as.integer(str_extract(basename(file_names), "\\d+"))
  
  # Scan all files once, collecting all needed info in a single pass
  file_scan <- map_dfr(file_names, function(f) {
    df_i <- fread(f, colClasses = list(character = c("eez_name", "ssp")))
    filename_key <- as.integer(str_extract(basename(f), "\\d+"))
    
    if (nrow(df_i) == 0) {
      return(tibble(filename_key = filename_key, actual_key = NA_integer_, n_rows = 0L, found_in_file = basename(f)))
    }
    
    tibble(filename_key = filename_key, actual_key = unique(df_i$taxon_key), n_rows = nrow(df_i), found_in_file = basename(f))
  })
  
  # File has data, but no accompanying species key
  data_but_no_key <- climate_by_spp_eez %>%
    filter(is.na(sciname)) %>%
    group_by(sciname) %>%
    distinct(file_found_in)
  
  # No file at all
  no_file <- species_names %>% 
    filter(!taxon_key %in% filename_keys) %>%
    as_tibble() %>%
    mutate(issue = "no file", found_in_file = NA_character_)
  
  # Empty files
  empty_files <- file_scan %>%
    filter(n_rows == 0) %>%
    left_join(as_tibble(species_names), by = c("filename_key" = "taxon_key")) %>%
    mutate(issue = "empty file") %>%
    select(taxon_key = filename_key, taxon_name, issue, found_in_file)
  
  # Mismatched keys — also check if no_file species are hiding in these files
  mismatched <- file_scan %>%
    filter(n_rows > 0, filename_key != actual_key) %>%
    left_join(as_tibble(species_names), by = c("actual_key" = "taxon_key")) %>%
    mutate(issue = "mismatched key") %>%
    select(taxon_key = actual_key, taxon_name, issue, found_in_file)
  
  # Check if any no_file species are hiding inside mismatched files
  no_file_found <- no_file %>%
    filter(taxon_key %in% mismatched$taxon_key) %>%
    left_join(mismatched %>% select(taxon_key, found_in_file), by = "taxon_key") %>%
    mutate(issue = "no file — found in mismatched file")
  
  no_file_not_found <- no_file %>%
    filter(!taxon_key %in% mismatched$taxon_key)
  
  # Get the files which have no data themselves and contain files from another species id
  mismatched_origin_file <- gsub(c("percentage_change_"), "", mismatched$found_in_file) 
  mismatched_origin_file <- as.numeric (gsub(".csv", "", mismatched_origin_file))
  
  mismatched_origin_file <- species_names %>%
    filter(
      taxon_key %in% 
             species_names$taxon_key[species_names$taxon_key %in% mismatched_origin_file]
    )
  
  # Correct files
  correct <- species_names %>%
    as_tibble() %>%
    filter(
      taxon_key %in% filename_keys,
      !taxon_key %in% empty_files$taxon_key,
      !taxon_key %in% mismatched$taxon_key,
      !taxon_key %in% mismatched_origin_file$taxon_key
    ) %>%
    mutate(issue = "correct", found_in_file = NA_character_)
  
  # Summary
  cat("Exposure data quality__________________________________________________________________", "\n")
  cat("File has data, but no accompanying species key                                     ", nrow(data_but_no_key), "\n")
  cat("No file (truly missing):                                                           ", nrow(no_file_not_found), "\n")
  cat("File unavailable, but data found in other file:                                    ", nrow(mismatched), "\n")
  cat("File available, but contains another species data (original data is nowhere else): ", nrow(mismatched_origin_file), "\n")
  cat("Empty file:                                                                        ", nrow(empty_files), "\n")
  cat("Incorrect:                                                                         ", nrow(no_file_not_found) + nrow(data_but_no_key) + nrow(no_file_found) + nrow(empty_files) + nrow(mismatched) + nrow(mismatched_origin_file), "\n")
  cat("Correct:                                                                           ", nrow(correct) - nrow(data_but_no_key), "\n")
  cat("Total accounted for:                                                               ", nrow(no_file_not_found) + nrow(no_file_found) + nrow(empty_files) + nrow(mismatched) + nrow(mismatched_origin_file) + nrow(correct), "\n")
  cat("Total in species_names:                                                            ", nrow(species_names), "\n")
  
  ## Remove file_found_in variable (only needed for data coverage diagnostics)
  ## Also drop NAs so that no na EEZs are joined to potential NAs in consumption
  ## data with EEZs
  
  ## Read in Juliano's mcp data
  climate_by_spp_eez <- climate_by_spp_eez %>%
    select(-file_found_in) %>%
    drop_na()
  
  # Read data
  
  ## Read in ARTIS sciname data
  sciname_taxa <- read_csv(paste0(dir, "sciname.csv"), show_col_types = FALSE) %>%
    build_taxa_levels()
  
  ## Read in ecoregion data
  ### Load in ecoregion data (i.e., average mcp values across ecoregion data)
  
  ### Read in ecoregions for largest soverign shape
  ecoregion_unique <- read.csv("../data/exposure/gap_filling/spatially_joined_ecoregion_to_EEZ/ecoregion_clean.csv") %>%
    select(-FREQUENCY) %>%
    rename(iso3c = "ISO_SOV1",
           shape_area = "MAX_Shape_Area")
  
  # Read in ecoregion data (all shapes associated with soverign)
ecoregion_full <- read.csv("../data/exposure/gap_filling/spatially_joined_ecoregion_to_EEZ/ecoregion.csv") %>%
  rename("iso3c" = "ISO_SOV1",
         ecoregion = "ECOREGION",
         province = "PROVINCE",
         realm = "REALM",
         shape_area = "Shape_Area")
  
  ### Join largest shape for the respective soverign country to their ecoregion names
  ecoregion <- left_join(ecoregion_unique, ecoregion_full, by = c("iso3c", "shape_area")) %>%
    filter(realm != "")
  
  ### Assign countries not associated with a marine ecoregion to "landlocked" code

  landlocked_countries <- consumption %>%
    distinct(consumer_iso3c) %>%
    drop_na() %>%
    add_region(col = "consumer_iso3c", region.col.name = "region") %>%
    filter(!consumer_iso3c %in% ecoregion$iso3c) %>%
    mutate(ecoregion = paste0("Landlocked, ", region),
           province = paste0("Landlocked, ", region),
           realm = paste0("Landlocked, ", region)) %>%
    select(-region) %>%
    rename(iso3c = consumer_iso3c)
  
  # Clean ecoregion data
  ecoregion <- bind_rows(ecoregion, landlocked_countries) %>%
    arrange(iso3c) %>%
    select(-shape_area)
  
  
  #   Correct Juliano's unaligned species names to Fishbase / SeaLifeBase naming
  ##  Problem: mcp data has (1) updated taxa names; and (2) out of date taxa names
  ##  We can't just correct all Juliano's data through Fishbase / SeaLifeBase
  ##  naming as they also have out of date items that are unaligned with ARTIS.
  ### Step 1: See which of Juliano's data when joined to 2019 ARTIS consumption data,
  ### is not joined

  #### Get list of unique ARTIS scientific names
  consumption_species_list <- consumption %>% # Filter to only species names
  distinct(sciname)

  #### Which of Juliano's species are not matching to ARTIS?
  unaligned_names <- correct$taxon_name[!str_to_lower(correct$taxon_name) %in%
                                          consumption_species_list$sciname]
  
  # Obtain fishbase / sealifebase corrections tables 
  if (!file.exists("../data/exposure/fb_slb_taxa_corrections/fishbase_corrections.rds")) {
    for (server in c("fishbase", "sealifebase")) {
    # Acquire full fishbase taxonomy
    taxa <- load_taxa(server = server, version = "25.04")
    
    # Get synonyms table
    syn <- synonyms(server = server, version = "25.04") %>%
      filter(Status == "accepted name" |
             Status == "synonym")
    
    # Join synonyms to corrected taxa names
    naming_corrections <- left_join(
      syn %>% filter(!(SpecCode == 0 | SpecCode == 1)) %>% select(synonym, SpecCode),
      taxa %>% select(SpecCode, Species)
    )
    
    # Save to RDS
    saveRDS(naming_corrections, paste0("../data/exposure/fb_slb_taxa_corrections/",server,"_corrections.rds"))
    
    }
  }
  
  # Load in Fishbase / Sealifebase taxa corrections
  fb_corrections <- readRDS("../data/exposure/fb_slb_taxa_corrections/fishbase_corrections.rds")
  slb_corrections <- readRDS("../data/exposure/fb_slb_taxa_corrections/sealifebase_corrections.rds")
  
  fb_slb_corrections <- bind_rows(fb_corrections, slb_corrections) # Join fb/slb
  
  # Join unaligned names to corrections
  corrections <- sapply(unaligned_names, function (x) unique(fb_slb_corrections$Species[x == fb_slb_corrections$synonym])) %>% unname()
  
  #### Create a separate dataframe of the catch projections that contain corrected names.
  #### We will use the Juliano's original naming as well as 
  #### the corrected one for two joins to ARTIS
  #### the first join will use Juliano's presently correct data, the second join
  #### will join Juliano's previously incorrect names (now corrected)
  
  ##### Create lookup table
  ##### (str_to_lower had to be done here becaue fb/slb lookup tables are uppercase)
  correction_lookups <- setNames(corrections, unaligned_names %>% str_to_lower())
  
  ##### Create data separate data frame
  climate_by_spp_eez_corrected <- climate_by_spp_eez %>%
  mutate(sciname = ifelse(
    sciname %in% names(correction_lookups),
    correction_lookups[sciname],
    NA
  )) %>%
    drop_na(sciname)

  ## Problem: mcp data has (1) updated taxa names; and (2) out of date taxa names
  ### Combine Juliano's pre and post-corrected naming data.
  climate_by_spp_eez_combined <- bind_rows(climate_by_spp_eez, climate_by_spp_eez_corrected) %>%
    group_by(eez_iso3c, sciname) %>%
    summarise(across(c(ssp126, ssp585), ~mean(.x, na.rm = TRUE))) %>%
    ungroup()
  
  ### For each species, average mcp projections across territories with an EEZ
  
  ### Join consumption data to catch projections
    consumption_mpc <- consumption %>%
    left_join(sciname_taxa, by = c("sciname" = "sciname")) %>%
      # Clean only artis sciname that didn't have data in sciname table
    left_join(climate_by_spp_eez_combined, by = c("eez_iso3c", "sciname")) %>%
      mutate(
        genus = case_when(
          sciname == "crassostrea gigas" ~ "crassostrea", TRUE ~ genus
        ),
        subfamily = case_when(
          sciname == "crassostrea gigas" ~ "crassostreinae", TRUE ~ subfamily
        ),
        family = case_when(
          sciname == "crassostrea gigas" ~ "ostreidae", TRUE ~ family
        ),
        order = case_when(
          sciname == "crassostrea gigas" ~ "ostreidae", TRUE ~ order
        ),
        class = case_when(
          sciname == "crassostrea gigas" ~ "bivalvia", TRUE ~ class
        ),
        phylum = case_when(
          sciname == "crassostrea gigas" ~ "mollusca", 
          sciname == "petromyzon marinus" ~ "chordata",
          TRUE ~ phylum,
        ),
        kingdom = case_when(
          sciname == "crassostrea gigas" ~ "animalia", TRUE ~ kingdom
        )
      )
    
  #### Use gap filling if user input says to, otherwise avoid
  if (gap_filling_method == "none") {
    
    x <- consumption_mpc
    
  } else {
    
    # fixit: maybe don't need lookup_table as a input here? but in the embedded
    # function
    x <- fill_with_higher_taxa(df = consumption_mpc,
                               gap_filling_method = gap_filling_method,
                               averaging = averaging,
                               ecoregion = ecoregion,
                               num_species = num_species,
                               assumed_sd = assumed_sd)
    
    
  }

  ### Add taxa levels to data
  interpolated <- x %>%
    mutate(
      taxa_level = case_when(
        sciname == kingdom    ~ "kingdom",
        sciname == phylum     ~ "phylum",
        sciname == superclass ~ "superclass",
        sciname == class      ~ "class",
        sciname == order      ~ "order",
        sciname == family     ~ "family",
        sciname == subfamily  ~ "subfamily",
        sciname == genus      ~ "genus",
        str_detect(sciname, " ") ~ "species"
      ),
      species = if_else(
        str_count(sciname, " ") == 1,
        word(sciname, 2),
        NA_character_
      )
    )
  
# Save interpolated consumption data to csv if gapfilling was done
if (gap_filling_method != "none") {
  if (gap_filling_method == "bayesian") {
    out_path <- paste0("../output/consumption_projections_gap_filled/bayesian/",as.character(assumed_sd),"_sd/consumption_projections_gap_filled_", portfolio, "_", averaging, "_", gap_filling_method, "_gapfill.parquet")
  } else {
   if (!file.exists(paste0("../output/consumption_projections_gap_filled/", num_species, "_species/consumption_projections_gap_filled_", portfolio, "_", averaging, "_", gap_filling_method, "_gapfill.parquet"))) {
    # Create directory if doesn't already exist
    if (!dir.exists(paste0("../output/consumption_projections_gap_filled/", num_species, "_species/"))) {
      dir.create(paste0("../output/consumption_projections_gap_filled/", num_species, "_species/"))
    }
    out_path <- paste0("../output/consumption_projections_gap_filled/", num_species, "_species/consumption_projections_gap_filled_", portfolio, "_", averaging, "_", gap_filling_method, "_gapfill.parquet")
  } 
  }
  write_parquet(interpolated, out_path)
}
  
  # Pivot to longer
  future_long <- interpolated %>%
    pivot_longer(c(ssp126, ssp585), names_to = "scenario", values_to = "future_change_catch")

  # Portfolio-specific row filter + catch-change multiplier
  # Note: the two portfolios intentionally use different multiplier formulas —
  # the foreign portfolio handles positive/negative/zero separately to preserve
  # directionality; the total portfolio uses a single linear conversion.
  stock_pre <- if (portfolio == "foreign_portfolio") {
    future_long %>%
      filter(
        consumer_iso3c != producer_iso3c,
        !is.na(live_weight_t),
        !is.na(future_change_catch)
      ) %>%
      mutate(future_change_catch = case_when(
        future_change_catch > 0 ~ (future_change_catch / 100) + 1,
        future_change_catch < 0 ~ 1 - ((future_change_catch / 100) * -1),
        future_change_catch == 0 ~ 1
      ))
  } else {
    future_long %>%
      filter(!is.na(live_weight_t), !is.na(future_change_catch)) %>%
      mutate(future_change_catch = 1 + (future_change_catch / 100))
  }

  # For each country, calculate the % change in aquatic food supply by calculating
  # a weighted average of each consumed taxa's projected percentage change in 
  # catch potential catch multiplied by by their respective proportional live
  # weight tonnage of their total consumed weight.
  
  # catch projection percentages by their proportional live weight tonnage
  # and converting to  
  exposure <- stock_pre %>%
    mutate(change_in_stock = live_weight_t * future_change_catch) %>%
    group_by(consumer_iso3c, scenario) %>% # FIXIT: changed from consumer_iso3c to producer_iso3c - switch back
    summarise(
      change_in_stock = sum(change_in_stock),
      live_weight_t   = sum(live_weight_t)
    ) %>%
    ungroup() %>%
    mutate(pct_change = 100 * ((change_in_stock - live_weight_t) / live_weight_t)) # FIXIT: Use a pipe here if I am testing the expsoure
    # On producer_iso3c
    # rename(consumer_iso3c = producer_iso3c) # FIXIT: Remove this as well when done testing
  
  # Return exposure dataframe
  return(exposure)
}
```

---

# Step 3 — Sensitivity (per portfolio)

Computes aquatic animal reliance as a share of total animal protein supply.

- `"foreign"` — sums across all consumption sources (domestic + foreign) to
  get total reliance.
- `"total"` — retains only the `foreign` consumption source row (no
  aggregation needed).

```{r build-sensitivity}
build_sensitivity <- function(dir) {

  supply_importance <- read_csv(
    paste0(dir, "sensitivity/fao_aquatic_reliance_source.csv"),
    show_col_types = FALSE
  )

  base <- supply_importance %>%
    filter(year == 2019, habitat == "marine", method == "capture", food_group == "aquatic")
  
  # Calculate reliance on foreign imported marine capture foods
  foreign <- base %>%
      filter(consumption_source == "foreign") %>%
      select(iso3c, foreign_reliance_pct = prop_animal_protein)
  
  # Calculate reiance on marine capture
  mc <- base %>%
      group_by(iso3c) %>%
      summarise(mc_reliance_pct = sum(prop_animal_protein)) %>%
      ungroup()
  
  # combine foreign / marine capture reliance data
  foreign_mc <- left_join(foreign, mc, by = "iso3c")
  
  return(foreign_mc)

}
```

---

# Step 4 — Adaptive capacity (portfolio-agnostic)

Loads all ten AC indicators across the four IPCC dimensions (Assets,
Flexibility, Learning, Social Organization), standardises column names, and
returns a single joined data frame keyed on `consumer_iso3c`.

The `consuming_countries` argument (a vector of ISO3C codes) is passed in
explicitly so this function has no hidden dependency on objects created in
earlier steps.

```{r build-adaptive-capacity}
build_adaptive_capacity <- function(dir, consuming_countries) {

  cc_df <- tibble(consumer_iso3c = consuming_countries, artis = 1)

  # country_coverage() returns the ARTIS countries that are missing from a
  # given AC dataset — useful for diagnosing coverage gaps
  country_coverage <- function(data, join_by) {
    full_join(cc_df, data %>% mutate(covered = 1), by = c("consumer_iso3c" = join_by)) %>%
      filter(is.na(covered), artis == 1) %>%
      pull(consumer_iso3c)
  }

  # --- Assets ---
  sanitation <- read_csv(paste0(dir, "adaptive capacity/assets/sanitation.csv"), show_col_types = FALSE) %>%
    filter(Indicator_Code == "SH_STA_BASS_ZS", Time_Period == 2019) %>%
    rename(sanitation = Value, consumer_iso3c = Geography_Code) %>%
    select(consumer_iso3c, sanitation)

  gdp <- read_csv(paste0(dir, "adaptive capacity/assets/gdp.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, gdp = `2019`) %>%
    filter(!is.na(gdp))

  trade_gdp <- read_csv(paste0(dir, "adaptive capacity/assets/trade_gdp.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Code` == "NE.TRD.GNFS.ZS") %>%
    select(consumer_iso3c = `Country Code`, gdp_trade = `2019`) %>%
    filter(!is.na(gdp_trade))

  # --- Flexibility ---
  life_expectancy <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/life_expectancy_at_birth.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator Code` == "SP.DYN.LE00.IN") %>%
    select(consumer_iso3c = `Country Code`, life_expectancy = `2019`) %>%
    filter(!is.na(life_expectancy))

  supermarkets <- read_csv(
    paste0(dir, "adaptive capacity/flexibility/supermarkets_per_100000.csv"),
    show_col_types = FALSE
  ) %>%
    mutate(consumer_iso3c = countrycode(Region, origin = "country.name", destination = "iso3c")) %>%
    select(consumer_iso3c, supermarkets = `2019`) %>%
    filter(!is.na(supermarkets))

  total_labor   <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_labor_force.csv"), show_col_types = FALSE)
  total_pop     <- read_csv(paste0(dir, "adaptive capacity/flexibility/total_population.csv"), show_col_types = FALSE) %>%
    select(consumer_iso3c = `Country Code`, total_pop_2019 = `2019`)

  prop_labor <- total_labor %>%
    left_join(total_pop, by = c("Country Code" = "consumer_iso3c")) %>%
    select(consumer_iso3c = `Country Code`, labor_2019 = `2019`, total_pop_2019) %>%
    mutate(prop_labor = labor_2019 / total_pop_2019) %>%
    filter(!is.na(prop_labor)) %>%
    select(consumer_iso3c, prop_labor)

  # --- Learning ---
  hci <- read_csv(paste0(dir, "adaptive capacity/learning/HCIData.csv"), show_col_types = FALSE) %>%
    filter(`Indicator Name` == "Human Capital Index (HCI) (scale 0-1)") %>%
    select(consumer_iso3c = `Country Code`, hci = `2018`) %>%
    filter(!is.na(hci))

  # --- Social organization ---
  gov_effectiveness <- read_csv(
    paste0(dir, "adaptive capacity/social organization/government_effectiveness_percentile.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Indicator ID` == "WB.WWGI.GE.PER.RNK") %>%
    select(consumer_iso3c = `Economy ISO3`, gov_effectiveness = `2019`) %>%
    filter(!is.na(gov_effectiveness))

  fsc <- read_csv(
    paste0(dir, "adaptive capacity/social organization/food-systems-dashboard-2025-03-04.csv"),
    show_col_types = FALSE
  ) %>%
    filter(`Start Year` == 2019 | `End Year` == 2019) %>%
    rename(fsc = Value, consumer_iso3c = ISO3) %>%
    select(consumer_iso3c, fsc)

  rol <- read_csv(
    paste0(dir, "adaptive capacity/social organization/WB-WWGI.csv"),
    show_col_types = FALSE
  ) %>%
    filter(Indicator == "Rule of Law: Percentile Rank") %>%
    select(consumer_iso3c = `Economy ISO3`, rol = `2019`)

  # Join all AC indicators
  gdp %>%
    left_join(trade_gdp,        by = "consumer_iso3c") %>%
    left_join(sanitation,       by = "consumer_iso3c") %>%
    left_join(supermarkets,     by = "consumer_iso3c") %>%
    left_join(life_expectancy,  by = "consumer_iso3c") %>%
    left_join(prop_labor,       by = "consumer_iso3c") %>%
    left_join(hci,              by = "consumer_iso3c") %>%
    left_join(gov_effectiveness, by = "consumer_iso3c") %>%
    left_join(fsc,              by = "consumer_iso3c") %>%
    left_join(rol,              by = "consumer_iso3c")
}
```

---

# Orchestrator — `build_e_s_ac_data()`

Calls `build_exposure()`, `build_sensitivity()`, and
`build_adaptive_capacity()` for a given portfolio type, joins the results, and
writes the output parquet.

```{r build-e-s-ac-data}
build_e_s_ac_data <- function(portfolio = c("foreign_portfolio", "total_portfolio"),
                              dir,
                              gap_filling_method = c("none", "region", "realm", "thorough", "hierarchical", "diagonal", "bayesian"),
                              averaging = c("unweighted", "weighted", "none"),
                              consumption,
                              num_species,
                              assumed_sd) {
  
  gap_filling_method <- match.arg(gap_filling_method)
  averaging <- match.arg(averaging)
  portfolio <- match.arg(portfolio)
  cat(sprintf("\n--- Building E/S/AC data: %s portfolio ---\n", portfolio))
  cat(sprintf("%s gap filling, %s averaging; %d minimum species\n", gap_filling_method, averaging, num_species))

  exposure     <- build_exposure(dir, portfolio, gap_filling_method = gap_filling_method, averaging = averaging, consumption = consumption, num_species = num_species, assumed_sd = assumed_sd)
  sensitivity  <- build_sensitivity(dir)

  consuming_countries <- exposure %>% distinct(consumer_iso3c) %>% pull()
  adaptive_cap <- build_adaptive_capacity(dir, consuming_countries)

  e_s_ac <- adaptive_cap %>%
    right_join(sensitivity,  by = c("consumer_iso3c" = "iso3c")) %>%
    left_join(exposure,      by = "consumer_iso3c")
  
  # Create directory if doesn't already exist
  if (!dir.exists(paste0("../output/e_s_ac/", num_species, "_species/"))) {
    
    dir.create(paste0("../output/e_s_ac/", num_species, "_species/"))
    
  }
  
  if (gap_filling_method == "none") {
    out_path <- paste0("../output/e_s_ac/", num_species, "_species/e_s_ac_", portfolio, "_", gap_filling_method, "_gapfill.csv")
  } else {
    if (gap_filling_method == "bayesian") {
      out_path <- paste0("../output/e_s_ac/", as.character(assumed_sd), "_sd/e_s_ac_", portfolio, "_", gap_filling_method, "_gapfill.csv")
    } else {
      out_path <- paste0("../output/e_s_ac/",num_species,"_species/e_s_ac_", portfolio, "_", averaging, "_", gap_filling_method, "_gapfill.csv")
    }
  }

  
  write_csv(e_s_ac, out_path)
  cat(sprintf("✅ Saved %s\n", out_path))

  invisible(e_s_ac)
}
```
