---
title: "02_Results"
author: "Connor Quiroz"
date: "2026-05-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(gtools)
```


# Load in Data

## Load in ARTIS data

```{r load artis data}
# Read in consumption data
# Set parquet database folder location
# db_folder <- "/Users/gephart/Library/CloudStorage/Dropbox/Research/ARTIS/ARTIS_1.1.0_FAO_2025_08_02/"

# Desktop folder
# db_folder <- paste0(here(),"/data/ARITS_1.1.0_FAO_2025_08_02")

# Laptop folder
db_folder <- "C:/Users/cjqui/Desktop/SGL/Data/ARITS_1.1.0_FAO_2025_08_02"

artis_ds <- arrow::open_dataset(file.path(db_folder, "datasets", "ARTIS_consumption_v1.1.0_FAO_mid_2025-08-02.parquet"))

# Summarize consumption file by source country
artis_ts_result <- artis_ds %>%
  filter(
    # FBS data of interest is only for direct human consumption
    end_use == "direct human consumption",
    # Filter to custom time series
    (hs_version == "HS96" & year <= 2003) |
      (hs_version == "HS02" & year >= 2004 & year <= 2009) |
      (hs_version == "HS07" & year >= 2010 & year <= 2012) |
      (hs_version == "HS12" & year >= 2013 & year <= 2020)
  ) %>%
  # Calculate consumption total by year, consumer, source, habitat and method 
  group_by(year, consumer_iso3c, consumption_source, habitat, method) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

consumption_full <- artis_ts_result %>%
  collect() %>%
  # Simplify consumption source to just indicate foreign vs domestic
  mutate(consumption_source = str_replace(consumption_source, " step 1", ""),
         consumption_source = str_replace(consumption_source, " step 2", ""))
```

# Introduction

## Finding: Which countries have increased their aquaculture consumption?

> Source: ARTIS

```{r see which countries have increasing aquauclture consumption}
for (i in c("Africa", "North America", "South America", "Europe", "Oceania")) {
  print(consumption_full %>% 
  group_by(consumer_iso3c, year, method) %>%
  summarize(sum = sum(consumption_live_t)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(region == i,
         method == "aquaculture") %>%
  ggplot(aes(x = year, y = sum)) +
  geom_line() +
  facet_wrap(~ consumer_iso3c))
}
```

> Countries with increasing aquauculture consumption include Australia, Nigeria, Brazil, Russia, and the United States.

# Geographic sourcing contribution to climate vulnerability

## Look at overall across regions

```{r}
theme_reliance <- function(base_size = 8,
                           axis_title_x = TRUE,
                           axis_title_y = FALSE,
                           legend       = TRUE) {
  base <- theme_minimal(base_size = base_size) +
    theme(
      text                    = element_text(family = "sans"),
      plot.background         = element_rect(fill = "transparent", color = NA),
      panel.background        = element_rect(fill = "transparent", color = NA),
      panel.grid.minor        = element_blank(),
      panel.grid.major.y      = element_line(
        color     = scales::alpha("black", 0.06),
        linewidth = 0.3
      ),
      panel.grid.major.x      = element_line(
        color     = scales::alpha("black", 0.25),
        linewidth = 0.3
      ),
      axis.text               = element_text(size = base_size - 3),   # 5 pt
      axis.title.x            = element_text(
        size  = base_size,
        face  = "bold",
        vjust = -0.5,
        margin = margin(t = 4)
      ),
      axis.title.y            = element_blank(),
      strip.text              = element_text(size = base_size - 1, face = "bold"),
      legend.position         = if (legend) "bottom" else "none",
      legend.title            = element_text(size = base_size - 1, face = "bold"),
      legend.text             = element_text(size = base_size - 2),
      legend.key.size         = unit(0.3, "cm"),
      legend.box              = "vertical",
      legend.spacing          = unit(0.1, "cm"),
      legend.margin           = margin(t = 1, b = 0),
      legend.box.margin       = margin(t = -2, b = -2),
      legend.background       = element_rect(fill = "transparent", color = NA),
      legend.box.background   = element_rect(fill = "transparent", color = NA),
      plot.margin             = margin(3, 3, 3, 3)
    )

  if (!axis_title_x) {
    base <- base + theme(axis.title.x = element_blank())
  }

  base
}

exposed_foods <- read_parquet("../output/future_catch_joined_consumption/interpolated_future_data_total_portfolio_weighted_diagonal_gapfill.parquet")

plot_viz <- function(data) {
  x <- data %>%
    ggplot(aes(x = pct_contribution, y = consumer_iso3c, color = consumption_type)) +
  # overall country change
  geom_point(
    aes(x = pct_change, y = consumer_iso3c),
    inherit.aes = FALSE,
    color = "black",
    fill = "white",
    shape = 21,
    stroke = 1.2,
    size = 3.8
  ) +
  # contribution points
  geom_point(size = 2.5,
             alpha = 0.8,
             position = position_jitter(height = 0.08)) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "grey50",
    linewidth = 0.4
  ) +
  facet_wrap(~ region, scales = "free_y") +
  labs(x = "Contribution to total percent change", y = NULL, color = "Consumption type") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom",
    axis.text = element_text(size = 5.5)
  ) +
  scale_color_manual(values = c("#1f77b4","#ff7f0e"))
  
  return(x)
}

plot_data <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  
  group_by(consumer_iso3c, consumption_type) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(consumer_iso3c) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  mutate(pct_change = sum(pct_contribution)) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(consumer_iso3c = fct_reorder(consumer_iso3c, -pct_change))

top_panel <- plot_data %>%
  distinct(consumer_iso3c, pct_change) %>%
  count(condition = pct_change > 0) %>%
  ggplot(aes(y = "", x = n, fill = condition)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = c(`FALSE` = "#B2182B", `TRUE` = "#2166AC"),
                    labels = c(`FALSE` = "Decrease in catch", `TRUE` = "Increase in catch")) +
  theme_reliance() +
  theme(
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(size = 6, face = "bold"),
    panel.grid.major.y = element_blank(),
    aspect.ratio = 0.05,
    legend.title = element_blank()
  )



p_top <- plot_data %>%
  filter(region %in% c("Africa", "Asia", "Europe")) %>%
  plot_viz() +
  guides(color = "none") +
  theme_reliance() +
  theme(axis.title.x = element_blank()) 

p_bottom <- plot_data %>%
  filter(!region %in% c("Africa", "Asia", "Europe")) %>%
  plot_viz() +
  theme_reliance()

bottom_row_narrowed <- cowplot::plot_grid(
  NULL, p_bottom, NULL,
  ncol       = 3,
  rel_widths = c(0.1, 0.75, 0.1)
)

bottom_panel <- plot_grid(p_top,bottom_row_narrowed, nrow = 2, rel_heights = c(1.25,1))

plot <- cowplot::plot_grid(
  top_panel,
  bottom_panel,
  nrow          = 2,
  rel_heights   = c(0.35, 2),
  labels        = c("A", "B"),
  label_size    = 9.5
)

# Save plot
ggsave("../images/exploratory/regional_contribution.jpg", plot = plot, device = "jpg", width = 8, height = 6, dpi = 300)
```


```{r}
plot_data %>%
  ggplot(aes(y = consumption_type, x = pct_contribution, color = consumption_type)) +
  geom_violin() +
  theme_reliance() +
  geom_jitter(alpha = 0.3) +
  labs(x ="Contribution to percent change in aquatic food consumption", color = "Consumption source")

plot_data %>%
  ggplot(aes(x = pct_contribution, y = consumer_iso3c, color = consumption_type)) +
  geom_point() +
  theme_reliance() +
  theme(axis.text.y = element_blank()) +
  labs(x ="Contribution to percent change in aquatic food consumption", color = "Consumption source")

x <- plot_data %>%
  filter(consumption_type == "domestic") %>%
  pull(pct_contribution)

y <- plot_data %>%
  filter(consumption_type == "foreign") %>%
  pull(pct_contribution)
  

t.test(x,y, alternative = "less")
```

## Look within regions

```{r}
theme_si <- function(base_size = 8,
                     show_x_title = TRUE,
                     show_y_title = TRUE,
                     show_legend  = TRUE) {

  base <- theme_minimal(base_size = base_size) +
    theme(
      text                  = element_text(family = "sans", color = "black"),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      panel.grid.minor      = element_blank(),
      panel.grid.major.y    = element_line(
        color = scales::alpha("black", 0.06), linewidth = 0.3
      ),
      panel.grid.major.x    = element_line(
        color = scales::alpha("black", 0.25), linewidth = 0.3
      ),
      axis.text             = element_text(size = base_size - 2.5, color = "black"),
      axis.title.x          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(t = 3)
      ),
      axis.title.y          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(r = 3)
      ),
      strip.text            = element_text(size = base_size - 1, face = "bold"),
      legend.position       = if (show_legend) "bottom" else "none",
      legend.title          = element_text(size = base_size - 1.5, face = "bold"),
      legend.text           = element_text(size = base_size - 2.5),
      legend.key.size       = unit(0.25, "cm"),
      legend.box            = "vertical",
      legend.spacing        = unit(0.05, "cm"),
      legend.margin         = margin(t = 0, b = 0),
      legend.box.margin     = margin(t = -2, b = -2),
      legend.background     = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.margin           = margin(3, 4, 3, 4)
    )

  if (!show_x_title) base <- base + theme(axis.title.x = element_blank())
  if (!show_y_title) base <- base + theme(axis.title.y = element_blank())
  base
}


regions <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")
for (i in regions) {
  
  p1 <- plot_data %>%
  filter(region == i) %>%
  distinct(consumer_iso3c, pct_change) %>%
  count(condition = pct_change > 0) %>%
  ggplot(aes(y = "", x = n, fill = condition)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = c(`FALSE` = "#B2182B", `TRUE` = "#2166AC"),
                    labels = c(`FALSE` = "Decrease in catch", `TRUE` = "Increase in catch")) +
  theme_reliance() +
  theme(
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(size = 6, face = "bold"),
    panel.grid.major.y = element_blank(),
    aspect.ratio = 0.05,
    legend.title = element_blank()
  ) +
    labs(x = "Number of countries")


both_forgn_dom <- plot_data %>%
  filter(region == i) %>%
  count(consumer_iso3c) %>%
  filter(n == 2) %>%
  pull(consumer_iso3c)

p2 <- plot_data %>%
  filter(region == i,
         consumer_iso3c %in% both_forgn_dom) %>%
  group_by(consumer_iso3c) %>%
  mutate(condition = prop_consumption == max(prop_consumption)) %>%
  ungroup() %>%
  count(consumption_type, condition) %>%
  filter(consumption_type == "domestic") %>%
  ggplot(aes(y = "", x = n, fill = condition)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = c(`FALSE` = "#ff7f0e", `TRUE` = "#1f77b4"),
                    labels = c(`FALSE` = "Foreign dominant", `TRUE` = "Domestic dominant")) +
  theme_reliance() +
  theme_reliance() +
  theme(
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(size = 6, face = "bold"),
    panel.grid.major.y = element_blank(),
    aspect.ratio = 0.05,
    legend.title = element_blank()
  ) +
  labs(x = "Number of countries")

bottom_panel <- plot_data %>%
  filter(region == i) %>%
  plot_viz() +
  theme(strip.text = element_blank())

top_panel <- plot_grid(p1,p2)

  title_panel <- ggplot() +
    labs(title = i) +
    theme_void() +
    theme(
      plot.title = element_text(
        hjust = 0.5, face = "bold", size = 11
      ),
      plot.margin = margin(15, 0, -5, 0),
      panel.spacing = unit(0, "pt")
    )
  

(final_plot <- plot_grid(title_panel, top_panel, bottom_panel,
                        nrow = 3,
                        rel_heights = c(0.005,0.1,0.45)))

ggsave(paste0("../images/exploratory/contribution_",i,".jpg"), plot = final_plot, device = "jpg", width = 5, height = 7, dpi = 300)
  
}
```

# Taxa contribution to climate vulnerability

```{r}
# By foreign / domestic & exposed / not exposed
exp_not_exp_sourcing_shannon <- exposed_foods %>%
  mutate(food_is_exposed = case_when(
    ssp126 > 0 ~ TRUE,
    ssp126 < 0 ~ FALSE
  )) %>%
group_by(consumer_iso3c, consumption_type, food_is_exposed, sciname) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(consumer_iso3c, consumption_type, food_is_exposed) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(consumer_iso3c, consumption_type, food_is_exposed) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  mutate(type = paste0(consumption_type, food_is_exposed)) %>%
  mutate(type = case_when(
    type == "foreignFALSE" ~ "Foreign, not exposed",
    type == "foreignTRUE" ~ "Foreign, exposed",
    type == "domesticFALSE" ~ "Domestic, not exposed",
    type == "domesticTRUE" ~ "Domestic, exposed"
  )) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  group_by(consumer_iso3c) %>%
  mutate(total_div = sum(shannon)) %>%
  ungroup() %>%
  mutate(consumer_iso3c = fct_reorder(consumer_iso3c, -total_div)) %>%
  filter(!is.na(food_is_exposed)) %>%
  select(consumer_iso3c, type, shannon)

exp_not_exp_sourcing %>%
  ggplot(aes(x = consumer_iso3c, y = shannon, fill = type)) +
  geom_col(position = "fill") +
  facet_wrap(~ region, scales = "free")

# By exposed / not exposed
exp_not_exp <- exposed_foods %>%
  mutate(food_is_exposed = case_when(
    ssp126 > 0 ~ TRUE,
    ssp126 < 0 ~ FALSE
  )) %>%
group_by(consumer_iso3c, food_is_exposed, sciname) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(consumer_iso3c, food_is_exposed) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(consumer_iso3c, food_is_exposed) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  mutate(type = food_is_exposed) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  group_by(consumer_iso3c) %>%
  mutate(total_div = sum(shannon)) %>%
  ungroup() %>%
  mutate(consumer_iso3c = fct_reorder(consumer_iso3c, -total_div))

exp_not_exp %>%
  ggplot(aes(x = consumer_iso3c, y = shannon, fill = type)) +
  geom_col(position = "fill") +
  facet_wrap(~ region, scales = "free") +
  geom_hline(yintercept = 0.5, linetype = "dashed")


# See how much shannon relates to the level of exposure contribution by dom / for / is exposed / not
exp_not_exp_sourcing_contribution <- exposed_foods %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  mutate(food_is_exposed = case_when(
    ssp126 >= 0 ~ TRUE,
    ssp126 < 0 ~ FALSE
  )) %>%
  group_by(consumer_iso3c, food_is_exposed, consumption_type) %>%
  summarise(
    change_in_stock = sum(change_in_stock),
    live_weight_t   = sum(live_weight_t),
    .groups = "drop"
  ) %>%
  group_by(consumer_iso3c) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t)) %>%
  mutate(pct_change = sum(pct_contribution)) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t)) %>%
  ungroup() %>%
  filter(!is.na(food_is_exposed)) %>%
  mutate(type = paste0(consumption_type, food_is_exposed)) %>%
  mutate(type = case_when(
    type == "foreignFALSE" ~ "Foreign, not exposed",
    type == "foreignTRUE" ~ "Foreign, exposed",
    type == "domesticFALSE" ~ "Domestic, not exposed",
    type == "domesticTRUE" ~ "Domestic, exposed"
  )) %>%
  select(consumer_iso3c, type, pct_contribution)

exp_not_exp_sourcing_contribution %>%
  filter(type == "Foreign, exposed") %>%
  left_join(exp_not_exp_sourcing_shannon, by = c("consumer_iso3c", "type")) %>%
  ggplot(aes(x = shannon, y = pct_contribution, color = type)) +
  geom_point() +
  geom_smooth(method = "lm")
  

for (i in c("Foreign, not exposed",
            "Foreign, exposed",
            "Domestic, not exposed",
            "Domestic, exposed")) {
  
  x <- exp_not_exp_sourcing_contribution %>%
  left_join(exp_not_exp_sourcing_shannon, by = c("consumer_iso3c", "type")) %>%
  filter(type == i)
  
  lm <- lm(pct_contribution ~ shannon, data = x)
  
  print(summary(lm))
  
}


# Follow up question: what does the diversity of countries' consumption look like when compared to their exposure?
```
> The level of exposure on exposed foreign supplies decreases # Fixit probably delete above analysis

## How does the diversity of countries' consumption compare to their level of exposure?

```{r}
exposure <- exposed_foods %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(consumer_iso3c) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) /
           ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1) %>%
  select(consumer_iso3c, exposure)

# By exposed / not exposed
shannon <- exposed_foods %>%
  group_by(consumer_iso3c, sciname) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(consumer_iso3c) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(consumer_iso3c) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  group_by(consumer_iso3c)

exposure %>%
  left_join(shannon, by = "consumer_iso3c") %>%
  ggplot(aes(x = shannon, y = exposure)) +
  geom_point() +
  facet_wrap(~ region)

  
```


## Is the level of exposure related to the diversity of marine taxa consumption?

```{r fixit probably delete}
# Join shannon to exposure data
shannon_exp <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  
  group_by(consumer_iso3c) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(consumer_iso3c) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  ungroup() %>%
             mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) / ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1) %>%
  left_join(shannon %>% select(consumer_iso3c, shannon)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI")

shannon_exp %>%
  ggplot(aes(x = shannon, y = exposure, color = region)) +
  geom_point() +
  geom_smooth(method = "lm", alpha = 0.15) +
  facet_wrap(~ region) +
  theme_reliance() +
  scale_color_manual(values = artis_palette(6))

lm <- lm(exposure ~ shannon * region, data = shannon_exp)


summary(lm)
```

## Globally, which are the most and least exposed species

```{r}
# Countries most exposed conusmed taxa
taxa <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(sciname) %>%
  summarise(
    change_in_stock = sum(change_in_stock),
    live_weight_t   = sum(live_weight_t),
    .groups = "drop"
  ) %>%
  ungroup() %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t, na.rm = TRUE)) %>%
  mutate(pct_change = sum(pct_contribution, na.rm = TRUE)) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t, na.rm = TRUE)) %>%
  ungroup()

taxa %>%
  filter(pct_contribution < 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(-pct_contribution, n = 5) %>%
  ggplot(aes(x = sciname, y = pct_contribution)) +
           geom_col()

taxa %>%
  filter(pct_contribution > 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(-pct_contribution, n = 5) %>%
  ggplot(aes(x = sciname, y = pct_contribution)) +
           geom_col()

taxa_by_eez <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(consumer_iso3c, sciname) %>%
  summarise(
    change_in_stock = sum(change_in_stock),
    live_weight_t   = sum(live_weight_t),
    .groups = "drop"
  ) %>%
  group_by(sciname) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t, na.rm = TRUE)) %>%
  mutate(pct_change = sum(pct_contribution, na.rm = TRUE)) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t, na.rm = TRUE)) %>%
  ungroup()


most_exposed_species <- taxa %>%
  filter(pct_contribution < 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(-pct_contribution, n = 5) %>%
  pull(sciname)

least_exposed_species <- taxa %>%
  filter(pct_contribution > 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(pct_contribution, n = 5) %>%
  pull(sciname)

taxa_by_eez <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(eez_iso3c, sciname) %>%
  summarise(
    change_in_stock = sum(change_in_stock),
    live_weight_t   = sum(live_weight_t),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t, na.rm = TRUE))
  
  
  # Check that the percent contributions to global exposure across EEZs for each species
  # adds back up to the global estimate
all.equal(
  taxa_by_eez %>%
  group_by(sciname) %>%
  summarize(pct_change = sum(pct_contribution)) %>%
  filter(sciname %in% most_exposed_species),
  
  taxa %>%
  filter(pct_contribution < 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(-pct_contribution, n = 5) %>%
    select(sciname, pct_change = pct_contribution)
  )

# See where most exposed species are being sourced from

taxa_by_eez %>%
  filter(sciname %in% most_exposed_species) %>%
  group_by(sciname) %>%
  top_n(-pct_contribution, n = 3) %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  ggplot(aes(x = sciname, y = pct_contribution, fill = eez_iso3c)) +
  geom_col()


taxa_by_eez %>%
  filter(sciname %in% least_exposed_species) %>%
  group_by(sciname) %>%
  top_n(pct_contribution, n = 3) %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  ggplot(aes(x = sciname, y = pct_contribution, fill = eez_iso3c)) +
  geom_col()


taxa %>%
  filter(pct_contribution > 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(pct_contribution, n = 5) %>%
  ggplot(aes(x = sciname, y = pct_contribution)) +
           geom_col()

library(dplyr)
library(ggplot2)
library(ggsankey)
library(networkD3)

taxa_by_eez %>%
  filter(sciname %in% most_exposed_species) %>%
  group_by(sciname) %>%
  slice_max(-pct_contribution, n = 2) %>%
  ungroup() %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(sciname, eez_name, value = pct_contribution) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")

taxa_by_eez %>%
  filter(sciname %in% least_exposed_species) %>%
  group_by(sciname) %>%
  slice_max(pct_contribution, n = 2) %>%
  ungroup() %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(sciname, eez_name, value = pct_contribution) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
```

## What are the most consumed taxa that regions are consuming?

```{r}
taxa_by_region <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(region, sciname) %>%
  summarise(
    change_in_stock = sum(change_in_stock),
    live_weight_t   = sum(live_weight_t),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t, na.rm = TRUE))

prop_consumed_in_eez <- exposed_foods %>%
    select(-region) %>%
    add_region("consumer_iso3c", "region") %>%
    group_by(region, eez_iso3c, sciname) %>%
    summarize(total_weight = sum(live_weight_t)) %>%
    group_by(region, sciname) %>%
    mutate(prop_in_eez = total_weight / sum(total_weight)) %>%
    drop_na() %>%
    select(region, eez_iso3c, sciname, prop_in_eez)

# Most exposed species
taxa_by_region %>%
  group_by(region) %>%
  slice_max(-pct_contribution, n = 2) %>%
  ungroup() %>%
  left_join(prop_consumed_in_eez, by = c("region", "sciname")) %>%
  group_by(region, sciname) %>%
  slice_max(prop_in_eez, n = 2) %>%   # top source EEZ per species per region
  ungroup() %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(region, sciname, eez_name, value = prop_in_eez) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  scale_x_discrete(labels = c("Consuming region", "Consumed taxa", "EEZ caught in")) +
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")

# Least exposed species
taxa_by_region %>%
  group_by(region) %>%
  slice_max(pct_contribution, n = 2) %>%
  ungroup() %>%
  left_join(prop_consumed_in_eez, by = c("region", "sciname")) %>%
  group_by(region, sciname) %>%
  slice_max(prop_in_eez, n = 2) %>%   # top source EEZ per species per region
  ungroup() %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(region, sciname, eez_name, value = prop_in_eez) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  scale_x_discrete(labels = c("Consuming region", "Consumed taxa", "EEZ caught in")) +
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
```

### Where are these species being sourced from primarily?

```{r fixit figure out what this does}
exp_not_exp %>%
  group_by(region) %>%
  select(consumer_iso3c, food_is_exposed, shannon, region) %>%
  group_by(consumer_iso3c) %>%
  mutate(prop = shannon / sum(shannon)) %>%
  group_by(food_is_exposed) %>%
  count(condition = prop > 0.5)
```


