---
title: "02_Results"
author: "Connor Quiroz"
date: "2026-05-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(gtools)
library(rstan)

# Set working directory for pulling in data easily path here
setwd("C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/code")
```


# Load in Data

## Load in ARTIS data

```{r load artis data}
# Read in consumption data
# Set parquet database folder location
# db_folder <- "/Users/gephart/Library/CloudStorage/Dropbox/Research/ARTIS/ARTIS_1.1.0_FAO_2025_08_02/"

# Desktop folder
# db_folder <- paste0(here(),"/data/ARITS_1.1.0_FAO_2025_08_02")

# Laptop folder
db_folder <- "C:/Users/cjqui/Desktop/SGL/Data/ARITS_1.1.0_FAO_2025_08_02"

artis_ds <- arrow::open_dataset(file.path(db_folder, "datasets", "ARTIS_consumption_v1.1.0_FAO_mid_2025-08-02.parquet"))

# Summarize consumption file by source country
artis_ts_result <- artis_ds %>%
  filter(
    # FBS data of interest is only for direct human consumption
    end_use == "direct human consumption",
    # Filter to custom time series
    (hs_version == "HS96" & year <= 2003) |
      (hs_version == "HS02" & year >= 2004 & year <= 2009) |
      (hs_version == "HS07" & year >= 2010 & year <= 2012) |
      (hs_version == "HS12" & year >= 2013 & year <= 2020)
  ) %>%
  # Calculate consumption total by year, consumer, source, habitat and method 
  group_by(year, consumer_iso3c, consumption_source, habitat, method) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

consumption_full <- artis_ts_result %>%
  collect() %>%
  # Simplify consumption source to just indicate foreign vs domestic
  mutate(consumption_source = str_replace(consumption_source, " step 1", ""),
         consumption_source = str_replace(consumption_source, " step 2", ""))
```

# Introduction

## Finding: Which countries have increased their aquaculture consumption?

> Source: ARTIS

```{r see which countries have increasing aquauclture consumption}
for (i in c("Africa", "North America", "South America", "Europe", "Oceania")) {
  print(consumption_full %>% 
  group_by(consumer_iso3c, year, method) %>%
  summarize(sum = sum(consumption_live_t)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(region == i,
         method == "aquaculture") %>%
  ggplot(aes(x = year, y = sum)) +
  geom_line() +
  facet_wrap(~ consumer_iso3c))
}
```

> Countries with increasing aquauculture consumption include Australia, Nigeria, Brazil, Russia, and the United States.

# Geographic sourcing contribution to climate vulnerability

## Look at overall across regions

```{r}
theme_reliance <- function(base_size = 8,
                           axis_title_x = TRUE,
                           axis_title_y = FALSE,
                           legend       = TRUE,
                           bold = TRUE,
                           rotate = TRUE) {
  base <- theme_minimal(base_size = base_size) +
    theme(
      text                    = element_text(family = "sans"),
      plot.background         = element_rect(fill = "transparent", color = NA),
      panel.background        = element_rect(fill = "transparent", color = NA),
      panel.grid.minor        = element_blank(),
      panel.grid.major.y      = element_line(
        color     = scales::alpha("black", 0.06),
        linewidth = 0.3
      ),
      panel.grid.major.x      = element_line(
        color     = scales::alpha("black", 0.25),
        linewidth = 0.3
      ),
      axis.text               = element_text(size = base_size - 3, angle = if (rotate) 45 else 0),   # 5 pt
      axis.title.x            = element_text(
        size  = base_size,
        face  = if (bold) "bold" else "plain",
        vjust = -0.5,
        margin = margin(t = 4)
      ),          
      strip.text              = element_text(size = base_size - 1, face = "bold"),
      legend.position         = if (legend) "bottom" else "none",
      legend.title            = element_text(size = base_size - 1, face = "bold"),
      legend.text             = element_text(size = base_size - 2),
      legend.key.size         = unit(0.3, "cm"),
      legend.box              = "vertical",
      legend.spacing          = unit(0.1, "cm"),
      legend.margin           = margin(t = 1, b = 0),
      legend.box.margin       = margin(t = -2, b = -2),
      legend.background       = element_rect(fill = "transparent", color = NA),
      legend.box.background   = element_rect(fill = "transparent", color = NA),
      plot.margin             = margin(3, 3, 3, 3)
    )

  if (!axis_title_x) {
    base <- base + theme(axis.title.x = element_blank())
  }

  if (!axis_title_y) {
    base <- base + theme(axis.title.y = element_blank())
  }

  base
}

  exposed_foods <- read_parquet("../output/consumption_projections_gap_filled/2_species/consumption_projections_gap_filled_total_portfolio_weighted_diagonal_gapfill.parquet")

  plot_viz <- function(data) {
  x <- data %>%
    ggplot(aes(x = pct_contribution, y = consumer_iso3c, fill = consumption_type)) +
  # overall country change
  # contribution points
  geom_col(alpha = 0.8) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "grey50",
    linewidth = 0.4
  ) +
    geom_point(
    aes(x = pct_change, y = consumer_iso3c),
    shape  = 21,
    fill   = "grey20",
    color  = "white",
    size   = 1.2,
    stroke = 0.3
  ) +
  facet_wrap(~ region, scales = "free_y") +
  labs(x = "Contribution to total % change in consumption", y = NULL, fill = "Consumption type") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom",
    axis.text = element_text(size = 4)
  ) +
  scale_color_manual(values = c("#1f77b4","#ff7f0e"))
  
  return(x)
}

library(ggplot2)
library(cowplot)
library(dplyr)
library(forcats)
library(scales)

# ── Unified colour palette ─────────────────────────────────────────────────────
col_domestic  <- "#1f77b4"
col_foreign   <- "#ff7f0e"
col_increase  <- "#2166AC"
col_decrease  <- "#B2182B"

# ── Data prep ─────────────────────────────────────────────────────────────────
plot_data <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100),
         change_in_stock     = live_weight_t * future_change_catch) %>%
  group_by(consumer_iso3c, consumption_type) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t,   na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(consumer_iso3c) %>%
  mutate(
    pct_change       = 100 * (change_in_stock - live_weight_t) / live_weight_t,
    pct_contribution = pct_change * live_weight_t / sum(live_weight_t)
  ) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  mutate(
    pct_change       = sum(pct_contribution),
    prop_consumption = live_weight_t / sum(live_weight_t)
  ) %>%
  ungroup() %>%
  mutate(
    consumer_iso3c = fct_reorder(consumer_iso3c, -pct_change),
    consumption_type = factor(consumption_type, levels = c("foreign", "domestic"))
  )

# ── Panel A: proportion of countries with increase vs decrease ─────────────────
panel_a <- plot_data %>%
  distinct(consumer_iso3c, pct_change) %>%
  count(condition = pct_change > 0) %>%
  mutate(fill_label = ifelse(condition, "Increase in catch", "Decrease in catch"),
         fill_label = factor(fill_label, levels = c("Increase in catch", "Decrease in catch"))) %>%
  ggplot(aes(y = "", x = n, fill = fill_label)) +
  geom_col(position = "fill") +
  scale_x_continuous(labels = percent_format()) +
  scale_fill_manual(
    values = c("Increase in catch" = col_increase, "Decrease in catch" = col_decrease),
    name   = NULL
  ) +
  guides(fill = guide_legend(nrow = 1)) +
  theme_reliance() +
  theme(
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    axis.title.x       = element_text(size = 6, face = "bold"),
    panel.grid.major.y = element_blank(),
    aspect.ratio       = 0.08,
    legend.position    = "bottom",
    legend.text        = element_text(size = 7),
    legend.key.size    = unit(0.35, "cm")
  ) +
  labs(x = "Percentage of countries", y = NULL)

# ── Panel B: proportion of countries with domestic / foreign dominant consumption ──────────────────────────────
panel_b <- plot_data %>%
    filter(consumer_iso3c %in% both_forgn_dom,
           consumption_type == "domestic") %>%
    mutate(condition = prop_consumption > 0.5) %>%
    count(condition) %>%
    ggplot(aes(y = "", x = n, fill = condition)) +
    geom_col(position = "fill") +
    scale_x_continuous(labels = scales::percent_format(), expand = c(0, 0)) +
    scale_fill_manual element_blank(),
      axis.text.y        = element_blank(),
      axis.ticks.y       =(
      values = c(`FALSE` = col_foreign, `TRUE` = col_domestic),
      labels = c(`FALSE` = "Foreign dominant consumption", `TRUE` = "Domestic dominant consumption"),
      na.translate = FALSE
    ) +
    theme_reliance() +
    theme(
      legend.title       = element_blank(),
      axis.title.x       = element_text(size = 6, face = "bold"),
      panel.grid.major.y = element_blank(),
      aspect.ratio       = 0.12
    ) +
    labs(x = "Percentage of countries", y = NULL)

# ── Panel C: violin + jitter of pct_contribution ──────────────────────────────
panel_c <- plot_data %>%
  ggplot(aes(y = consumption_type, x = pct_contribution, color = consumption_type)) +
  geom_violin(alpha = 0.5) +
  geom_jitter(alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_color_manual(
    values = c(foreign = col_foreign, domestic = col_domestic),
    name   = NULL
  ) +
  guides(color = "none") +
  theme_reliance(bold = FALSE) +
  labs(x = "Contribution to total % change in consumption", y = NULL)

# ── Panel D: regional plots ────────────────────────────────────────────────────
p_top <- plot_data %>%
  filter(region %in% c("Africa", "Asia", "Europe")) %>%
  plot_viz() +
  scale_fill_manual(values = c(col_foreign, col_domestic)) +
  guides(fill = "none") +
  theme(legend.title       = element_blank(),
      axis.title.x       = element_text(size = 6, face = "bold"),
      panel.grid.major.y = element_blank()) +
        labs(y = "Country")

p_bottom_inner <- plot_data %>%
  filter(!region %in% c("Africa", "Asia", "Europe")) %>%
  plot_viz() +
  scale_fill_manual(values = c(col_foreign, col_domestic)) +
  theme(
      panel.grid.major.y = element_blank()
    ) +
      labs(y = "Country")

p_bottom <- plot_grid(NULL, p_bottom_inner, NULL,
                      ncol = 3, rel_widths = c(0.1, 0.75, 0.1))

panel_d <- plot_grid(p_top, p_bottom, nrow = 2, rel_heights = c(1.25, 1))

# ── Assemble ──────────────────────────────────────────────────────────────────
final_plot <- plot_grid(
  plot_grid(panel_a, panel_b, ncol = 2, align = "h",
  labels = c("A","B"), label_size  = 9.5),
  panel_c,
  panel_d,
  nrow        = 3,
  rel_heights = c(0.35, 0.45, 2),
  labels      = c("","C","D"),
  label_size  = 9.5
)

# ── Save ──────────────────────────────────────────────────────────────────────
ggsave(
  "../images/exploratory/regional_contribution.jpg",
  plot   = final_plot,
  device = "jpg",
  width  = 7,
  height = 8,
  dpi    = 300
)
```

# Are domestic or foreign foods the most exposed?

```{r}
plot_data <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100),
         change_in_stock     = live_weight_t * future_change_catch) %>%
  group_by(consumer_iso3c, consumption_type) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t,   na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(consumer_iso3c) %>%
  mutate(
    pct_change       = 100 * (change_in_stock - live_weight_t) / live_weight_t,
    pct_contribution = pct_change * live_weight_t / sum(live_weight_t)
  ) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  mutate(
    pct_change       = sum(pct_contribution),
    prop_consumption = live_weight_t / sum(live_weight_t)
  ) %>%
  ungroup() %>%
  mutate(
    consumer_iso3c = fct_reorder(consumer_iso3c, -pct_change),
    consumption_type = factor(consumption_type, levels = c("foreign", "domestic"))
  )

x <- plot_data %>%
  filter(consumption_type == "domestic") %>%
  pull(pct_contribution)

y <- plot_data %>%
  filter(consumption_type == "foreign") %>%
  pull(pct_contribution)

t.test(x,y, alternative = "less")
```

## Look at domestic / foreign exposure within regions

```{r}
theme_si <- function(base_size = 8,
                     show_x_title = TRUE,
                     show_y_title = TRUE,
                     show_legend  = TRUE) {

  base <- theme_minimal(base_size = base_size) +
    theme(
      text                  = element_text(family = "sans", color = "black"),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      panel.grid.minor      = element_blank(),
      panel.grid.major.y    = element_line(
        color = scales::alpha("black", 0.06), linewidth = 0.3
      ),
      panel.grid.major.x    = element_line(
        color = scales::alpha("black", 0.25), linewidth = 0.3
      ),
      axis.text             = element_text(size = base_size - 2.5, color = "black"),
      axis.title.x          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(t = 3)
      ),
      axis.title.y          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(r = 3)
      ),
      strip.text            = element_text(size = base_size - 1, face = "bold"),
      legend.position       = if (show_legend) "bottom" else "none",
      legend.title          = element_text(size = base_size - 1.5, face = "bold"),
      legend.text           = element_text(size = base_size - 2.5),
      legend.key.size       = unit(0.25, "cm"),
      legend.box            = "vertical",
      legend.spacing        = unit(0.05, "cm"),
      legend.margin         = margin(t = 0, b = 0),
      legend.box.margin     = margin(t = -2, b = -2),
      legend.background     = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.margin           = margin(3, 4, 3, 4)
    )

  if (!show_x_title) base <- base + theme(axis.title.x = element_blank())
  if (!show_y_title) base <- base + theme(axis.title.y = element_blank())
  base
}

library(ggplot2)
library(cowplot)
library(dplyr)
library(forcats)
library(countrycode)
library(purrr)

# ── Unified colour palette ─────────────────────────────────────────────────────
col_domestic  <- "#1f77b4"
col_foreign   <- "#ff7f0e"
col_increase  <- "#2166AC"
col_decrease  <- "#B2182B"

regions <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")

# ── Helper: build one region plot ─────────────────────────────────────────────
make_region_plot <- function(data, region_name) {

  d <- data %>%
    filter(region == region_name) %>%
    mutate(
      consumer_label = countrycode(consumer_iso3c, "iso3c", "country.name"),
      consumption_type = factor(consumption_type, levels = c("foreign", "domestic"))
    )

  # ── p1: direction of change (proportional) ──────────────────────────────────
  p1 <- d %>%
    distinct(consumer_iso3c, pct_change) %>%
    count(condition = pct_change > 0) %>%
    ggplot(aes(y = "", x = n, fill = condition)) +
    geom_col(position = "fill") +
    scale_x_continuous(labels = scales::percent_format(), expand = c(0, 0)) +
    scale_fill_manual(
      values = c(`TRUE` = col_increase, `FALSE` = col_decrease),
      labels = c(`TRUE` = "Increase in catch", `FALSE` = "Decrease in catch"),
      na.translate = FALSE
    ) +
    theme_reliance() +
    theme(
      legend.title       = element_blank(),
      axis.text.y        = element_blank(),
      axis.ticks.y       = element_blank(),
      axis.title.x       = element_text(size = 6, face = "bold"),
      panel.grid.major.y = element_blank(),
      aspect.ratio       = 0.12
    ) +
    labs(x = "Percentage of countries", y = NULL)

  # ── p2: domestic vs foreign dominance (proportional) ────────────────────────
  both_forgn_dom <- d %>%
    count(consumer_iso3c) %>%
    filter(n == 2) %>%
    pull(consumer_iso3c)

  p2 <- d %>%
    filter(consumer_iso3c %in% both_forgn_dom,
           consumption_type == "domestic") %>%
    mutate(condition = prop_consumption > 0.5) %>%
    count(condition) %>%
    ggplot(aes(y = "", x = n, fill = condition)) +
    geom_col(position = "fill") +
    scale_x_continuous(labels = scales::percent_format(), expand = c(0, 0)) +
    scale_fill_manual(
      values = c(`FALSE` = col_foreign, `TRUE` = col_domestic),
      labels = c(`FALSE` = "Foreign dominant consumption", `TRUE` = "Domestic dominant consumption"),
      na.translate = FALSE
    ) +
    theme_reliance() +
    theme(
      legend.title       = element_blank(),
      axis.text.y        = element_blank(),
      axis.ticks.y       = element_blank(),
      axis.title.x       = element_text(size = 6, face = "bold"),
      panel.grid.major.y = element_blank(),
      aspect.ratio       = 0.12
    ) +
    labs(x = "Percentage of countries", y = NULL)

  # ── Shared legend (harvested from a temporary plot) ──────────────────────────
  legend_source <- bind_rows(
    tibble(label = "Increase in catch",   fill = col_increase,  group = "Direction"),
    tibble(label = "Decrease in catch",   fill = col_decrease,  group = "Direction"),
    tibble(label = "Domestic dominant",   fill = col_domestic,  group = "Source"),
    tibble(label = "Foreign dominant",    fill = col_foreign,   group = "Source")
  )

  shared_legend <- get_legend(
    ggplot(legend_source, aes(x = label, fill = label)) +
      geom_col(aes(y = 1)) +
      scale_fill_manual(values = setNames(legend_source$fill, legend_source$label),
                        name = NULL) +
      guides(fill = guide_legend(nrow = 2)) +
      theme_void() +
      theme(legend.position = "bottom",
            legend.text     = element_text(size = 7),
            legend.key.size = unit(0.35, "cm"))
  )

  top_panel <- plot_grid(
    plot_grid(p1, p2, ncol = 2, align = "h"),
    shared_legend,
    nrow = 2,
    rel_heights = c(1, 0.35)
  )

  # ── Middle panel: proportion by country (no x-axis text — shared with bottom) 
  middle_panel <- d %>%
    ggplot() +
    geom_col(aes(x = fct_reorder(consumer_iso3c, pct_change),
                 y = prop_consumption * 100,
                 fill = consumption_type)) +
    scale_fill_manual(values = c(col_foreign, col_domestic)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_reliance(axis_title_y = TRUE, axis_title_x = FALSE, legend = FALSE, rotate = FALSE) +
    theme(
      axis.ticks.x = element_blank(),
      axis.text.x = element_text(size = 3)
    ) +
    labs(y = "Percentage of consumption", x = NULL)

  # ── Bottom panel ─────────────────────────────────────────────────────────────
  bottom_panel <- d %>%
    mutate(consumption_type = factor(consumption_type, levels = c("domestic", "foreign"))) %>%
    plot_viz() +
    labs(y = "Country") +
    theme(strip.text = element_blank()) +
    scale_fill_manual(values = c(col_domestic, col_foreign)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))

  # ── Title ────────────────────────────────────────────────────────────────────
  title_panel <- ggplot() +
    labs(title = region_name) +
    theme_void() +
    theme(
      plot.title  = element_text(hjust = 0.5, face = "bold", size = 12),
      plot.margin = margin(15, 0, -5, 0)
    )

  # ── Assemble ─────────────────────────────────────────────────────────────────
  plot_grid(
    title_panel,
    top_panel,
    middle_panel,
    bottom_panel,
    nrow        = 4,
    align       = "v",
    axis        = "lr",
    rel_heights = c(0.07, 0.22, 0.33, 1.0)
  )
}

# ── Render & save all regions ─────────────────────────────────────────────────
walk(regions, function(r) {
  p <- make_region_plot(plot_data, r)
  ggsave(
    filename = paste0("../images/exploratory/contribution_", r, ".jpg"),
    plot     = p,
    device   = "jpg",
    width    = 7,       # double-column width
    height   = 9,
    dpi      = 300
  )
  message("Saved: ", r)
})
```

# Taxa contribution to climate vulnerability

```{r}
# By foreign / domestic & exposed / not exposed
exp_not_exp_sourcing_shannon <- exposed_foods %>%
  mutate(food_is_exposed = case_when(
    ssp126 > 0 ~ TRUE,
    ssp126 < 0 ~ FALSE
  )) %>%
group_by(consumer_iso3c, consumption_type, food_is_exposed, sciname) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(consumer_iso3c, consumption_type, food_is_exposed) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(consumer_iso3c, consumption_type, food_is_exposed) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  mutate(type = paste0(consumption_type, food_is_exposed)) %>%
  mutate(type = case_when(
    type == "foreignFALSE" ~ "Foreign, not exposed",
    type == "foreignTRUE" ~ "Foreign, exposed",
    type == "domesticFALSE" ~ "Domestic, not exposed",
    type == "domesticTRUE" ~ "Domestic, exposed"
  )) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  group_by(consumer_iso3c) %>%
  mutate(total_div = sum(shannon)) %>%
  ungroup() %>%
  mutate(consumer_iso3c = fct_reorder(consumer_iso3c, -total_div)) %>%
  filter(!is.na(food_is_exposed)) %>%
  select(consumer_iso3c, type, shannon)

exp_not_exp_sourcing %>%
  ggplot(aes(x = consumer_iso3c, y = shannon, fill = type)) +
  geom_col(position = "fill") +
  facet_wrap(~ region, scales = "free")

# By exposed / not exposed
exp_not_exp <- exposed_foods %>%
  mutate(food_is_exposed = case_when(
    ssp126 > 0 ~ TRUE,
    ssp126 < 0 ~ FALSE
  )) %>%
group_by(consumer_iso3c, food_is_exposed, sciname) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(consumer_iso3c, food_is_exposed) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(consumer_iso3c, food_is_exposed) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  mutate(type = food_is_exposed) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  group_by(consumer_iso3c) %>%
  mutate(total_div = sum(shannon)) %>%
  ungroup() %>%
  mutate(consumer_iso3c = fct_reorder(consumer_iso3c, -total_div))

exp_not_exp %>%
  ggplot(aes(x = consumer_iso3c, y = shannon, fill = type)) +
  geom_col(position = "fill") +
  facet_wrap(~ region, scales = "free") +
  geom_hline(yintercept = 0.5, linetype = "dashed")


# See how much shannon relates to the level of exposure contribution by dom / for / is exposed / not
exp_not_exp_sourcing_contribution <- exposed_foods %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  mutate(food_is_exposed = case_when(
    ssp126 >= 0 ~ TRUE,
    ssp126 < 0 ~ FALSE
  )) %>%
  group_by(consumer_iso3c, food_is_exposed, consumption_type) %>%
  summarise(
    change_in_stock = sum(change_in_stock),
    live_weight_t   = sum(live_weight_t),
    .groups = "drop"
  ) %>%
  group_by(consumer_iso3c) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t)) %>%
  mutate(pct_change = sum(pct_contribution)) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t)) %>%
  ungroup() %>%
  filter(!is.na(food_is_exposed)) %>%
  mutate(type = paste0(consumption_type, food_is_exposed)) %>%
  mutate(type = case_when(
    type == "foreignFALSE" ~ "Foreign, not exposed",
    type == "foreignTRUE" ~ "Foreign, exposed",
    type == "domesticFALSE" ~ "Domestic, not exposed",
    type == "domesticTRUE" ~ "Domestic, exposed"
  )) %>%
  select(consumer_iso3c, type, pct_contribution)

exp_not_exp_sourcing_contribution %>%
  filter(type == "Foreign, exposed") %>%
  left_join(exp_not_exp_sourcing_shannon, by = c("consumer_iso3c", "type")) %>%
  ggplot(aes(x = shannon, y = pct_contribution, color = type)) +
  geom_point() +
  geom_smooth(method = "lm")
  

for (i in c("Foreign, not exposed",
            "Foreign, exposed",
            "Domestic, not exposed",
            "Domestic, exposed")) {
  
  x <- exp_not_exp_sourcing_contribution %>%
  left_join(exp_not_exp_sourcing_shannon, by = c("consumer_iso3c", "type")) %>%
  filter(type == i)
  
  lm <- lm(pct_contribution ~ shannon, data = x)
  
  print(summary(lm))
  
}


# Follow up question: what does the diversity of countries' consumption look like when compared to their exposure?
```
> The level of exposure on exposed foreign supplies decreases # Fixit probably delete above analysis

## How does the diversity of countries' consumption compare to their level of exposure?

```{r}
exposure <- exposed_foods %>%
  filter(gapfill_method == "Gapfill not needed") %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(consumer_iso3c) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) /
           ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1) %>%
  select(consumer_iso3c, pct_change)

shannon <- exposed_foods %>%
  group_by(consumer_iso3c, sciname) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(consumer_iso3c) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(consumer_iso3c) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  group_by(consumer_iso3c)

test_delete <- exposure %>%
  left_join(shannon)

test_delete %>%
  ggplot(aes(x = shannon, y = pct_change)) +
  geom_point() +
  theme_reliance(bold = FALSE, axis_title_y = TRUE) +
  labs(x = "Shannon diversity", y = "% change in portfolio")
```

```{r}

exposed_foods <- read_parquet("../output/consumption_projections_gap_filled/2_species/consumption_projections_gap_filled_total_portfolio_weighted_diagonal_gapfill.parquet")

# exposure <- exposed_foods %>%
#   filter(gapfill_method == "Gapfill not needed") %>%
#   mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
#   mutate(change_in_stock = live_weight_t * future_change_catch) %>%
#   group_by(consumer_iso3c) %>%
#   summarise(
#     change_in_stock = sum(change_in_stock, na.rm = TRUE),
#     live_weight_t   = sum(live_weight_t, na.rm = TRUE),
#     .groups = "drop"
#   ) %>%
#   mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
#   mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) /
#            ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
#            exposure = -exposure + 1) %>%
#   select(consumer_iso3c, exposure)



exposure_eez <- exposed_foods %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  # filter(gapfill_method == "Gapfill not needed") %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(eez_iso3c) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) /
           ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1) %>%
  select(eez_iso3c, exposure)

# By exposed / not exposed
shannon <- exposed_foods %>%
  group_by(eez_iso3c, sciname) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(eez_iso3c) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(eez_iso3c) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  add_region("eez_iso3c", "region") %>%
  filter(eez_iso3c != "NEI") %>%
  group_by(eez_iso3c)

exposure_eez <- exposed_foods %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  # filter(gapfill_method == "Gapfill not needed") %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(eez_iso3c) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) /
           ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1) %>%
  select(eez_iso3c, exposure)

# Shannon by eez, consumption_type
shannon <- exposed_foods %>%
  mutate(food_is_exposed = case_when(
    ssp126 > 0 ~ FALSE,
    ssp126 < 0 ~ TRUE
  )) %>%
  group_by(consumer_iso3c, sciname, food_is_exposed, consumption_type) %>%
  summarize(sciname_total = sum(live_weight_t)) %>%
  group_by(consumer_iso3c, food_is_exposed, consumption_type) %>%
  mutate(genus_total_country = sum(sciname_total)) %>%
  ungroup() %>%
  mutate(pi = sciname_total / genus_total_country, pi_lnpi = pi * log(pi)) %>%
  group_by(consumer_iso3c, food_is_exposed, consumption_type) %>%
  summarize(shannon = -sum(pi_lnpi)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI") %>%
  group_by(consumer_iso3c)

# Exposure contribution 
contrib_to_exposure_eez <- exposed_foods %>%
  mutate(food_is_exposed = case_when(
    ssp126 > 0 ~ FALSE,
    ssp126 < 0 ~ TRUE
  )) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  # filter(gapfill_method == "Gapfill not needed") %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(consumer_iso3c, food_is_exposed, consumption_type) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  group_by(consumer_iso3c) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t)) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t)) %>%
  mutate(pct_change = sum(pct_contribution))

######################################################

test_df <- contrib_to_exposure_eez %>%
  drop_na() %>%
  left_join(shannon, by = c("consumer_iso3c", "food_is_exposed", "consumption_type")) %>%
  mutate(food_is_exposed = case_when(
    food_is_exposed == TRUE ~ "Foods with projected declines",
    food_is_exposed == FALSE ~ "Foods with projected increases",
  )) %>%
    filter(consumer_iso3c != "NEI") %>%
  ungroup %>%
  mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) / ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1)

  test_df %>%
    filter(!(consumption_type == "foreign" & food_is_exposed == "Foods with projected declines")) %>%
      ggplot(aes(x = shannon, y = prop_consumption, color = consumption_type)) +
      geom_point() +
      geom_smooth(method = "lm", alpha = 0.15) +
      facet_wrap(~ food_is_exposed) +
      scale_color_manual(values = c("#1f77b4", "#ff7f0e")) +
      theme_reliance(axis_title_y = TRUE, bold = FALSE) +
      labs(x = "Shannon diversity", y = "Proportion of consumption")

test_df %>%
  ggplot(aes(x = shannon, y = pct_change, color = consumption_type)) +
  geom_point() +
  facet_wrap(~ food_is_exposed) +
  geom_smooth(method = "lm", alpha = 0.15) +
  theme_reliance(axis_title_y = TRUE, bold = FALSE) +
  labs(x = "Shannon diversity", y = "% change in portfolio") +
  scale_color_manual(values = c("#1f77b4", "#ff7f0e"))

x <- test_df %>%
  filter(food_is_exposed == "Foods with projected declines",
  consumption_type == "foreign") %>%
    select(consumer_iso3c, pct_change, V1 = shannon) %>%
    left_join(test_df %>%
  filter(food_is_exposed == "Foods with projected increases",
  consumption_type == "foreign") %>%
    select(consumer_iso3c, V2 = shannon), by = "consumer_iso3c") %>%
    left_join(test_df %>%
  filter(food_is_exposed == "Foods with projected declines",
  consumption_type == "domestic") %>%
    select(consumer_iso3c, V3 = shannon), by = "consumer_iso3c") %>%
    left_join(test_df %>%
  filter(food_is_exposed == "Foods with projected increases",
  consumption_type == "domestic") %>%
    select(consumer_iso3c, V4 = shannon), by = "consumer_iso3c")

lm(pct_change ~ V1 + V2 + V3 + V4,data = x) %>% summary()

df1 <- test_df %>%
  filter(food_is_exposed == "Foods with projected declines",
  consumption_type == "foreign") %>%
    select(consumer_iso3c, prop_consumption, shannon, pct_change, exposure) %>%
    exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "region")

df2 <- test_df %>%
  filter(food_is_exposed == "Foods with projected increases",
  consumption_type == "foreign") %>%
    select(consumer_iso3c, prop_consumption, shannon, pct_change, exposure) %>%
    exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "region") 

df3 <- test_df %>%
  filter(food_is_exposed == "Foods with projected declines",
  consumption_type == "domestic") %>%
    select(consumer_iso3c, prop_consumption, shannon, pct_change, exposure) %>%
    exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "region")

df4 <- test_df %>%
  filter(food_is_exposed == "Foods with projected increases",
  consumption_type == "domestic") %>%
    select(consumer_iso3c, prop_consumption, shannon, pct_change) %>%
    exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "region") 
   

lm(prop_consumption ~ shannon, data = df1) %>% summary()
lm(prop_consumption ~ shannon, data = df2) %>% summary()
lm(prop_consumption ~ shannon, data = df3) %>% summary()
lm(prop_consumption ~ shannon, data = df4) %>% summary()

# create_map(data = df1, fill = "shannon", country.col.name = "consumer_iso3c")
# create_map(data = df2, fill = "shannon", country.col.name = "consumer_iso3c")
# create_map(data = df3, fill = "shannon", country.col.name = "consumer_iso3c")
# create_map(data = df4, fill = "shannon", country.col.name = "consumer_iso3c")

```

## Bayesian Model

```{r hierarchical bayesian model}
dfs <- list(df1, df2, df3, df4)
for (i in seq_along(dfs)) {
  # Input data (region)
    region <- dfs[[i]]$region %>% factor() %>% as.integer()

    num_regions <- region %>% unique() %>% length()
    
    # Response — z-scored
    y <- scale(dfs[[i]] %>% pull(prop_consumption))[, 1]

    # Stan data
    stan_data <- list(
      N                = length(region),
      y                = y,
      region           = region,
      num_regions      = num_regions
    )

    # Initial values matching Stan model parameters
    inits <- function() {
      list(
        mu_region    = rep(0, 6),
        sigma        = 1
      )
    }

    # Fit Stan model
    fit <- stan(
      file   = "../utils/shannon_bayesian.stan",
      data   = stan_data,
      init   = inits,
      chains = 3,
      iter   = 4000,
      warmup = 1000
    )
    
mcmc_output <- extract(fit, permuted = T)

# Get regions names via facator level

# Convert factor levels back into country names
region_factor_numeric <- df1$region %>% factor() %>% as.integer()

region_names_ordered <- df1 %>%
  mutate(factor_level = region_factor_numeric) %>%
  distinct(region, factor_level) %>%
  arrange(factor_level) %>%
  pull(region)

colnames(mcmc_output$mu_region) <- region_names_ordered

# Interval colours — three lightness steps per model colour
if (i %in% c(1,3)) {
  color = "red"
} else {
  color = "blue"
}
col_95 <- colorspace::lighten(color, 0.65)
col_80 <- colorspace::lighten(color, 0.35)
col_50 <- colorspace::lighten(color, 0.05)

plot <- data.frame(mcmc_output$mu_region) %>%
  pivot_longer(cols = everything()) %>%
  group_by(name) %>%
  summarise(
      lo95 = quantile(value, 0.025),
      lo80 = quantile(value, 0.10),
      lo50 = quantile(value, 0.25),
      med  = quantile(value, 0.50),
      hi50 = quantile(value, 0.75),
      hi80 = quantile(value, 0.90),
      hi95 = quantile(value, 0.975),
      .groups = "drop"
    ) %>%
  ggplot(aes(y = name)) +
  geom_segment(aes(x = lo95, xend = hi95),
               color = col_95, linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo80, xend = hi80),
                 color = col_80, linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo50, xend = hi50),
                 color = col_50, linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_point(aes(x = med), shape = 21, size = 1.6,
               fill = "white", color = "grey20", stroke = 0.4) +

    geom_vline(xintercept = 0, linetype = "dashed",
               linewidth = 0.35, color = "grey40") +
    scale_color_identity() +
    # scale_x_continuous(limits = c(-x_limit, x_limit)) +
    # scale_y_discrete(labels = var_labels) +
    # labs(title = title) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title         = element_text(size = 3.5, face = "bold", hjust = 0.5,
                                        margin = margin(b = 2)),
      plot.tag           = element_text(size = 7, face = "bold"),
      # plot.tag.position  = tag_pos,
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      axis.text.x        = element_text(size = 5.5, color = "black"),
      axis.title         = element_blank(),
      plot.margin        = margin(4, 6, 4, 6)
      # axis.text.y        = element_blank(),
      # axis.ticks.y       = element_blank()
    )

print(plot)

}   
df1
```

## Is the level of exposure related to the diversity of marine taxa consumption?

```{r fixit probably delete}
# Join shannon to exposure data
shannon_exp <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  
  group_by(eez_iso3c_iso3c) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(consumer_iso3c) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  ungroup() %>%
             mutate(exposure = (pct_change - min(pct_change, na.rm = TRUE) ) / ( max(pct_change, na.rm = TRUE) - min(pct_change, na.rm = TRUE) ),
           exposure = -exposure + 1) %>%
  left_join(shannon %>% select(consumer_iso3c, shannon)) %>%
  add_region("consumer_iso3c", "region") %>%
  filter(consumer_iso3c != "NEI")

shannon_exp %>%
  ggplot(aes(x = shannon, y = exposure, color = region)) +
  geom_point() +
  geom_smooth(method = "lm", alpha = 0.15) +
  facet_wrap(~ region) +
  theme_reliance() +
  scale_color_manual(values = artis_palette(6))

lm <- lm(exposure ~ shannon * region, data = shannon_exp)
summary(lm)
```

## Globally, which are the most and least exposed species

```{r}
# Countries most exposed conusmed taxa
taxa <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(sciname) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ungroup() %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t, na.rm = TRUE)) %>% # Calculate the percent contribution that each taxa has toward the total global percent change in consumption
  mutate(pct_change = sum(pct_contribution, na.rm = TRUE)) %>% # Global percent projected change in consumption
  mutate(prop_consumption = live_weight_t / sum(live_weight_t, na.rm = TRUE)) %>%
  ungroup()

# Calculate the percentage that global consumption is projected to change 
projected_global_change <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  pull(pct_change)




most_exposed_species <- taxa %>%
  filter(pct_contribution < 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(-pct_contribution, n = 5) %>%
  pull(sciname)

least_exposed_species <- taxa %>%
  filter(pct_contribution > 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(pct_contribution, n = 5) %>%
  pull(sciname)

taxa_by_eez <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(eez_iso3c, sciname) %>%
  summarise(
    change_in_stock = sum(change_in_stock, na.rm = TRUE),
    live_weight_t   = sum(live_weight_t, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(pct_change = sum(pct_change * (live_weight_t / sum(live_weight_t, na.rm = TRUE)), na.rm = TRUE)) ## 

# Ensure that country percents add up to global average (should be 4.99267% if using 2 species diagonal weighted gapfill)
unique(taxa_by_eez$pct_change) %>% round(5) == projected_global_change %>% round(5)
  
  
  # Check that the percent contributions to global exposure across EEZs for each species
  # adds back up to the global estimate
all.equal(
  taxa_by_eez %>%
  group_by(sciname) %>%
  summarize(pct_change = sum(pct_contribution)) %>%
  filter(sciname %in% most_exposed_species),
  
  taxa %>%
  filter(pct_contribution < 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(-pct_contribution, n = 5) %>%
    select(sciname, pct_change = pct_contribution)
  )

# See where most exposed species are being sourced from

taxa_by_eez %>%
  filter(sciname %in% most_exposed_species) %>%
  group_by(sciname) %>%
  top_n(-pct_contribution, n = 3) %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  ggplot(aes(x = sciname, y = pct_contribution, fill = eez_iso3c)) +
  geom_col()


taxa_by_eez %>%
  filter(sciname %in% least_exposed_species) %>%
  group_by(sciname) %>%
  top_n(pct_contribution, n = 3) %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  ggplot(aes(x = sciname, y = pct_contribution, fill = eez_iso3c)) +
  geom_col()


taxa %>%
  filter(pct_contribution > 0) %>%
  # group_by(consumer_iso3c) %>%
  top_n(pct_contribution, n = 5) %>%
  ggplot(aes(x = sciname, y = pct_contribution)) +
           geom_col()

library(dplyr)
library(ggplot2)
library(ggsankey)
library(networkD3)

# Most exposed taxa
taxa_by_eez %>%
  filter(sciname %in% most_exposed_species) %>%
  group_by(sciname) %>%
  slice_max(-pct_contribution, n = 2) %>%
  ungroup() %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(sciname, eez_name, value = pct_contribution) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")

# Least exposed taxa
taxa_by_eez %>%
  filter(sciname %in% least_exposed_species) %>%
  group_by(sciname) %>%
  slice_max(pct_contribution, n = 2) %>%
  ungroup() %>%
  select(eez_iso3c, sciname, pct_contribution) %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(sciname, eez_name, value = pct_contribution) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
```

## What are the most consumed taxa that regions are consuming?

```{r}
taxa_by_region <- exposed_foods %>%
  select(-ssp585, -cv_ssp585) %>%
  mutate(future_change_catch = 1 + (ssp126 / 100)) %>%
  mutate(change_in_stock = live_weight_t * future_change_catch) %>%
  group_by(region, sciname) %>%
  summarise(
    change_in_stock = sum(change_in_stock),
    live_weight_t   = sum(live_weight_t),
    .groups = "drop"
  ) %>%
  mutate(pct_change = 100 * ((change_in_stock - live_weight_t)) / live_weight_t) %>%
  mutate(pct_contribution = pct_change * live_weight_t / sum(live_weight_t, na.rm = TRUE))

prop_consumed_in_eez <- exposed_foods %>%
    select(-region) %>%
    add_region("consumer_iso3c", "region") %>%
    group_by(region, eez_iso3c, sciname) %>%
    summarize(total_weight = sum(live_weight_t)) %>%
    group_by(region, sciname) %>%
    mutate(prop_in_eez = total_weight / sum(total_weight)) %>%
    drop_na() %>%
    select(region, eez_iso3c, sciname, prop_in_eez)

# Most exposed species
taxa_by_region %>%
  group_by(region) %>%
  slice_max(-pct_contribution, n = 2) %>%
  ungroup() %>%
  left_join(prop_consumed_in_eez, by = c("region", "sciname")) %>%
  group_by(region, sciname) %>%
  slice_max(prop_in_eez, n = 2) %>%   # top source EEZ per species per region
  ungroup() %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(region, sciname, eez_name, value = prop_in_eez) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  scale_x_discrete(labels = c("Consuming region", "Consumed taxa", "EEZ caught in")) +
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")

# Least exposed species
taxa_by_region %>%
  group_by(region) %>%
  slice_max(pct_contribution, n = 2) %>%
  ungroup() %>%
  left_join(prop_consumed_in_eez, by = c("region", "sciname")) %>%
  group_by(region, sciname) %>%
  slice_max(prop_in_eez, n = 2) %>%   # top source EEZ per species per region
  ungroup() %>%
  mutate(eez_name = countrycode(eez_iso3c, "iso3c", "country.name")) %>%
  make_long(region, sciname, eez_name, value = prop_in_eez) %>%
  ggplot(aes(x = x,
             next_x = next_x,
             node = node,
             next_node = next_node,
             value = value,
             fill = node)) +
  geom_sankey(flow.alpha = 0.7, node.color = "black") +
  geom_sankey_label(aes(label = node), size = 3, color = "black") +
  theme_minimal() +
  scale_x_discrete(labels = c("Consuming region", "Consumed taxa", "EEZ caught in")) +
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
```

### Where are these species being sourced from primarily?

```{r fixit figure out what this does}
exp_not_exp %>%
  group_by(region) %>%
  select(consumer_iso3c, food_is_exposed, shannon, region) %>%
  group_by(consumer_iso3c) %>%
  mutate(prop = shannon / sum(shannon)) %>%
  group_by(food_is_exposed) %>%
  count(condition = prop > 0.5)
```


