---
title: "01 — Climate Risk Analysis"
author: "Connor Quiroz, Juliano Palacios-Abrantes, and  Jessica A. Gephart"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(gtools)
library(here)
library(showtext)


# Set up global font family for plots
# Register a specific static weight from the variable font file
font_add(family = "Bahnschrift", regular = "/path/to/bahnschrift.ttf")
showtext_auto()
showtext_opts(dpi = 96)  # match whatever dpi your ggsave/rsvg pipeline uses

# font_import()
family = "Bahnschrift"
```

# Objectives

This notebook:

1. Runs `01_Analysis.Rmd` to load all data-building functions into the
   environment.
2. Builds E/S/AC datasets for two portfolio types: a country's **foreign
   imports** and its **entire consumption** portfolio.
3. Calculates climate risk for each portfolio.
4. Runs a sensitivity analysis on adaptive capacity components *(work in
   progress)*.

---

# Setup

> **User input:** Set `file_path` to the root of your input data folder.
> Everything else runs automatically.

## Calculate the risk index based on a country’s exposure, sensitivity, and adaptive capacity components.

```{r load-working-directory-and-functions}
# Set working directory for pulling in data easily path here
# setwd("C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/code")

## Connor's laptop path
file_path <- "C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/data/"

# Knitting 00_Preprocess_Data.Rmd loads all helper functions for generating the exposure, sensitiviyty, adaptive capacity.
knitr::knit("01_Analysis.Rmd", output = tempfile())


# Load helper scripts
sapply(
  list.files(file.path("../utils"),
             pattern = "\\.R$",
             full.names = TRUE,
             recursive = TRUE),
  source
) %>%
  invisible()

dir <- file_path
```

```{r}
# Calculate vulnerabilities

for (i in c("thorough", "diagonal", "hierarchical")) {
  for (j in c("unweighted", "weighted")) {
    for (k in c(1, 2, 5)) {
      vuln <- calculate_vulnerability(
        portfolio = "total_portfolio",
        dir = file_path,
        gap_filling_method = i,
        averaging_type = j,
        num_species = k,
        drop_na = FALSE
      )
      
    }
    
  }
  
}

# Run correlations among different exposure calculations
files <- list.files("../output/vulnerability", pattern = "\\.csv$",
                    full.names = TRUE, recursive = TRUE)

# Read in all exposure files
data_list <- lapply(files, read.csv)

# Derive the comparisons we need to run against
comparison_matrix <- number_unique_combs(18)

# Add on a third column to the matrix for correlation
cor_matrix <- matrix(NA, nrow = nrow(comparison_matrix), ncol = 1)

cor_comparison_matrix <- cbind(comparison_matrix, cor_matrix)

for (i in 1:nrow(comparison_matrix)) {
  
  ds1_idx <- comparison_matrix[i,1]
  ds2_idx <- comparison_matrix[i,2]
  ds1 <- data_list[[ds1_idx]]
  ds2 <- data_list[[ds2_idx]]
  
  # Run a continuous-variable comparison
  cor <- cor(na.omit(ds1$exposure), na.omit(ds2$exposure), method = "kendall")
  cor_comparison_matrix[i,3] <- cor

  
      
}

cor_comparison_matrix %>%
  data.frame() %>%
  arrange(-X3) %>%
  mutate(comparison = paste0(X1, "_", X2)) %>%
  rename(cor = X3) %>%
  ggplot(aes(x = fct_reorder(comparison, -cor), cor)) +
  geom_col(width = 1.5) +
  geom_point(size = 0.5) +
  theme_light() +
  labs(x = "Comparison", y = "Kendall tau correlation") +
  theme(panel.grid.major.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title = element_text(size = 15),
        axis.text = element_text(color = "black"))
```

---

# Check gap filled consumption data coverage

```{r}
consumption_gap_filled_1 <- read_parquet(
  "../output/consumption_catch_proj_gap_filled/1_species/consumption_total_portfolio_unweighted_diagonal_gapfill.parquet")

consumption_gap_filled_2 <- read_parquet(
  "../output/consumption_catch_proj_gap_filled/2_species/consumption_total_portfolio_unweighted_diagonal_gapfill.parquet")

consumption_gap_filled_5 <- read_parquet(
  "../output/consumption_catch_proj_gap_filled/5_species/consumption_total_portfolio_unweighted_diagonal_gapfill.parquet")

consumption_gap_fills <- list(
  consumption_gap_filled_1,
  consumption_gap_filled_2,
  consumption_gap_filled_5
)

species_gap_fills <- c("1 species","2 species", "5 species")

for (i in seq_along(species_gap_fills)) {
  
  data <- consumption_gap_fills[[i]]
  
  plot <- cons_filled %>%
  mutate(gapfill_method = case_when(
    !(gapfill_method == "No fill needed" |
      gapfill_method == "Unsuccessful gapfill") ~
      substring(gapfill_method, first = 1, last = nchar(gapfill_method)-8),
    TRUE ~ gapfill_method
  )) %>%
  mutate(gapfill_performed = paste0(gapfill_method, ", ", gapfill_level),
         gapfill_performed = case_when(
           gapfill_performed ==
             "No fill needed, No fill needed" ~ "No fill needed",
           gapfill_performed ==
             "Unsuccessful gapfill, Unsuccessful gapfill" ~
             "Unsuccessful gapfill",
           TRUE ~ gapfill_performed
           )) %>%
  group_by(fill_type) %>%
  mutate(total_weight = sum(live_weight_t)) %>%
  group_by(gapfill_performed) %>%
  summarize(live_weight_t = sum(live_weight_t), .groups = "drop") %>%
  mutate(prop_consumption = live_weight_t / total_weight) %>%
  mutate(gapfill_performed = factor(
    gapfill_performed, levels = c(
      "No fill needed",
      "eez, genus", "eez, family", "realm, genus",
      "eez, order", "realm, family", "region, genus",
      "eez, class", "realm, order", "region, family",
      "global, genus", "eez, phylum", "realm, class",
      "region, order", "global, family", "eez, kingdom",
      "realm, phylum", "region, class", "global, order",
      "realm, kingdom", "region, phylum", "global, class",
      "region, kingdom", "global, phylum", "global, kingdom",
      "Unsuccessful gapfill")
  )) %>%
  ggplot(
  aes(
    x = 100 * prop_consumption,
    y = fct_rev(gapfill_performed)
  )
) +
  geom_col(
    fill = "#3182BD",
    width = 0.7
  ) +
  geom_text(
    aes(label = paste0(round(100 * prop_consumption, 1), "%")),
    hjust = -0.15,
    fontface = "bold",
    size = 3.5) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    x = "Percentage of ARTIS consumption accounted for",
    y = "Gapfill performed",
    title = paste0(species_gap_fills[i], " minimum")
  ) +
  theme_classic(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5)
  ) +
    facet_wrap(~ fill_type)

print(plot)
  
}

```

# See where gapfills are coming from

```{r}
consumption_proj <- read_parquet("../output/consumption_catch_proj_gap_filled/2_species/consumption_total_portfolio_unweighted_diagonal_gapfill.parquet")

consumption_proj %>%
  filter(str_d)
  group_by(sciname_hs_modified, eez_iso3c, gapfill_method, gapfill_level) %>%
  summarize(live_weight_t = sum(live_weight_t), .groups = "drop") %>%
  mutate(pct_consumption = round(live_weight_t / sum(live_weight_t) * 100, 10)) %>%
  filter(gapfill_method == "eez_gapfill", gapfill_level == "genus") %>%
  arrange(-pct_consumption)


  lookup_table <- readRDS("../data/exposure/lookup_tables/gap_fill/1_species/lookup_table.rds")
```


# Map out values on continuous & discrete color scale

```{r}
vuln_1 <- calculate_vulnerability(portfolio = "total_portfolio",
                                  dir = file_path,
                                  gap_filling_method = "diagonal",
                                  averaging_type = "unweighted",
                                  num_species = 1,
                                  drop_na = FALSE
                                  )

vuln_2 <- calculate_vulnerability(portfolio = "total_portfolio",
                                  dir = file_path,
                                  gap_filling_method = "diagonal",
                                  averaging_type = "weighted",
                                  num_species = 2,
                                  drop_na = FALSE
                                  )

vuln_5 <- calculate_vulnerability(portfolio = "total_portfolio",
                                  dir = file_path,
                                  gap_filling_method = "diagonal",
                                  averaging_type = "unweighted",
                                  num_species = 5,
                                  drop_na = FALSE
                                  )
# Continuous color scale
vuln_1 %>% create_map(fill = "exposure", country.col.name = "consumer_iso3c")
vuln_2 %>% create_map(fill = "exposure", country.col.name = "consumer_iso3c")
vuln_5 %>% create_map(fill = "exposure", country.col.name = "consumer_iso3c")


# Discrete color scale
vuln_dfs <- list(vuln_1, vuln_2, vuln_5)

species_gap_fills <- c("1 species","2 species", "5 species")


for (i in seq_along(species_gap_fills)) {
  
  vuln_df <- vuln_dfs[[i]]

  

  # Create bounds that will be "low vulnerability", "moderate vulnerability",
  # and "high vulnerability"
mod_vuln_lb <- vuln_df$exposure %>% quantile(0.33)
  high_vuln_lb <- vuln_df$exposure %>% quantile(0.66)
  
  plot <- vuln_df %>%
    mutate(
      vuln_type = case_when(
        exposure <= mod_vuln_lb ~ "low exposure",
        exposure >= mod_vuln_lb &
          exposure <= high_vuln_lb ~ "moderate exposure",
        exposure >= high_vuln_lb ~ "high exposure"
      ),
      vuln_type = factor(vuln_type, levels = c("low exposure",
                                            "moderate exposure",
                                            "high exposure"))
    ) %>%
    create_map(
      fill = "vuln_type",
      country.col.name = "consumer_iso3c"
    ) +
    scale_fill_manual(
      values = c(
        "low exposure" = "#FFF3B0",
        "moderate exposure" = "#F4A261",
        "high exposure" = "#C44536"
      ),
      breaks = c(
        "low exposure",
        "moderate exposure",
        "high exposure"
      ),
      labels = c(
        "low climate exposure",
        "moderate climate exposure",
        "high climate exposure"
      )
    ) +
    labs(fill = "",
         title = ) +
    theme(legend.key.spacing.x = unit(1, "cm"),
          plot.title = element_text(hjust = 0.5))
  
  print(plot)
  
}


```

# Correlation between different methods of calculating exposure

```{r fixit delete}
vuln_2_unwgt <- calculate_vulnerability(portfolio = "total_portfolio",
                                  dir = file_path,
                                  gap_filling_method = "diagonal",
                                  averaging = "unweighted",
                                  num_species = 2,
                                  drop_na = TRUE
                                  )

vuln_2_wgt <- calculate_vulnerability(portfolio = "total_portfolio",
                                  dir = file_path,
                                  gap_filling_method = "diagonal",
                                  averaging = "weighted",
                                  num_species = 2,
                                  drop_na = TRUE
                                  )


plot(vuln_2_unwgt$exposure, vuln_2_unwgt$vulnerability,  method = "kendall")

lm(vulnerability ~ adaptive_capacity, data = vuln_2_unwgt) %>%
  summary()

pairs(vuln_2_unwgt[,2:5])
cor(vuln_2_unwgt[,2:5])

consumption_gap_filled %>%
  count(gapfill_method, gapfill_level)
```


# Check adaptive c
# Check adaptive capacity data variable coverage

```{r}
e_s_ac <- read_csv("../output/e_s_ac/1_species/e_s_ac_total_portfolio_unweighted_diagonal_gapfill.csv")

e_s_ac %>%
  group_by(gdp, gdp_trade, sanitation,
           supermarkets, life_expectancy, prop_labor,
           hci, gov_effectiveness, fsc,	rol) %>%
  summarise_each(list(na_count = ~sum(is.na(.))))
gdp	gdp_trade	sanitation	supermarkets	life_expectancy	prop_labor	hci	gov_effectiveness	fsc	rol

e_s_ac %>%
  group_by(consumer_iso3c) %>%
  mutate(
    n_NA = rowSums(is.na(across(c(
      gdp, gdp_trade, sanitation, supermarkets,
      life_expectancy, prop_labor, hci,
      gov_effectiveness, fsc, rol
    ))))
  ) %>%
  mutate(n_NA = factor(n_NA)) %>%
  create_map(fill = "n_NA", country.col.name = "consumer_iso3c") +
  scale_fill_viridis_d(option = "viridis")
```

# Sensitivity analyis

Here, we compare use Kendall tau correlation to compare differences in rank across different methods for (1) different methods for calculating exposure; (3) different contributing measures to sensitivity; (3) the individual variables contributing measures within adaptive capacity; and (4) the grouped measures that contributing to adaptive capacity (i.e., learning, flexibility, assets, social organization), where each grouped measure can contain multiple variables of adaptive capacity. We first run sensitivity analysis on each of the measures

## Analyze the sensitivity of different exposure measures

```{r exposure measurement: sensitivity analysis}
# Helper function to dynamically select either exposure or vulnerability columns
# to run sensitivity analysis
run_sensitivity_analysis <- function(dfs, comparison = c("exposure", "vulnerability")) {
  comparison <- match.arg(comparison)

  list <- lapply(dfs, function(df) df[[comparison]])   # or `pull()` if you want a vector
  
  # Run sensitivity analysis
  dir_probs_matrix <- exp_sensitivity_analysis(exp_list = list)
  
  return(dir_probs_matrix)
}

# Create a list out of the different exposure calculations
dfs <- list(vuln_total_no_gapfill,
             vuln_total_unwgt_region_gapfill,
             vuln_total_wgt_region_gapfill,
             vuln_total_unwgt_thorough_gapfill,
             vuln_total_wgt_thorough_gapfill,
             vuln_total_unwgt_hierarchical_gapfill,
             vuln_total_wgt_hierarchical_gapfill,
             vuln_total_unwgt_diagonal_gapfill,
             vuln_total_wgt_diagonal_gapfill)

# Run sensitvity analysis
dir_probs_matrix <- run_sensitivity_analysis(dfs, comparison = "exposure")


# compute min r_kendall per combination
mins <- dir_probs_matrix %>%
  drop_na() %>%
  group_by(combination) %>%
  summarise(
    min_y = min(r_kendall, na.rm = TRUE) # x‑position of the min
  )

# Step 2: main plot
plot <- data.frame(dir_probs_matrix) %>%
  drop_na() %>%
  mutate(combination = factor(combination)) %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = combination)) +
  geom_line() +
  facet_wrap(~ combination, nrow = 6) +
  labs(
    y = "Kendall tau correlation",
    x = "Dirichlet-assigned weight for focal exposure method",
    title = "Sensitivity of exposure score"
  ) +
  theme_cowplot() +
  guides(color = "none") +
  theme(
    axis.text = element_text(size = 8),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +

  # Step 3: add the min labels
  geom_text(
    data = mins,
    aes(x = 0.6, y = 0.6, label = sprintf("Min: %.3f", min_y)),
    color = "black",
    size = 2.5,
    vjust = -0.5,
    inherit.aes = FALSE
  )


# Save plot
ggsave("../images/exploratory/sensitiity_analysis_exposure.jpg", plot = plot, device = "jpg", width = 10, height = 6.5, dpi = 300)
```

## Sensitivity analysis of sensitivity measure

```{r sensitivity measurement: sensitivity analysis}
# Run sensitivity analysis
dir_probs_matrix_sensitivity <- sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                                                             analysis = "sensitivity")

plot <- dir_probs_matrix_sensitivity %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var),
         var = case_when(var == "1" ~ "Foreign reliance (%)",
                         var == "2" ~ "Marine capture reliance (%)")) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal sensitivity variable", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_sensitivity.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```

## Sensitivity analysis on adaptive capacity variables

### Variables within adaptive capacity

```{r adaptive capacity measurement: sensitivity analysis}
dir_probs_matrix_ac_variables <- sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                                                             analysis = "adaptive_capacity_variables")

plot <- dir_probs_matrix_ac_variables %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var),
         var = case_when(var == "1" ~ "GDP per capita",
                         var == "2" ~ "Trade (% of GDP)",
                         var == "3" ~ "Sanitation",
                         var == "4" ~ "Supermarkets",
                         var == "5" ~ "Life expectancy",
                         var == "6" ~ "Proportion labor force",
                         var == "7" ~ "Human capacity index",
                         var == "8" ~ "Government effectiveness",
                         var == "9" ~ "Food safety capacity",
                         var == "10" ~ "Rule of law")) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
    theme(legend.position = "none",
          axis.text = element_text(size = 8),
          strip.text = element_text(size = 7),
          plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal adaptive capacity variable", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_adaptive_capacity_variables.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```

### Components within adaptive capacity

```{r}
dir_probs_matrix_ac_components <- 
  sens_ac_sensitivity_analysis(input_df = e_s_ac_total_no_gapfill,
                               analysis = "adaptive_capacity_components")

plot <- dir_probs_matrix_ac_components %>%
  filter(r_kendall != 1) %>%
  mutate(var = factor(var)) %>%
  mutate(var = case_when(
         var == "1" ~ "Assets",
         var == "2" ~ "Flexibility",
         var == "3" ~ "Social Organization",
         var == "4" ~ "Learning"),
         dirichlet_prob = case_when(
           var != "Learning" ~ dirichlet_prob * 3,
           TRUE ~ dirichlet_prob
         )) %>%
  drop_na() %>%
  ggplot(aes(x = dirichlet_prob, y = r_kendall, color = var)) +
  geom_line() +
  theme_cowplot() +
  facet_wrap(~ var) +
  scale_color_viridis_d(end = 0.8) +
    theme(legend.position = "none",
          axis.text = element_text(size = 8),
          strip.text = element_text(size = 7),
          plot.title = element_text(hjust = 0.5)) +
  labs(x = "Dirichlet-assigned weight for focal adaptive capacity component", y = "Kendall tau correlation", title = "Sensitivity of vulnerability score")

ggsave("../images/exploratory/sensitivity_analysis_adaptive_capacity_components.jpg", plot = plot, device = "jpg", width = 6, height = 4, dpi = 300)
```
