---
title: "03_Results"
author: "Connor Quiroz, Juliano Palacios-Abrantes, and  Jessica A. Gephart"
date: "2026-05-11"
output: html_document
---

# Load in packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)
library(gtools)
library(rstan)
library(showtext)
library(grid)


# Set up global font family for plots
# Register a specific static weight from the variable font file
font_add(family = "Bahnschrift", regular = "/path/to/bahnschrift.ttf")
showtext_auto()
showtext_opts(dpi = 96)  # match whatever dpi your ggsave/rsvg pipeline uses

# font_import()
family = "Bahnschrift"

# Set working directory for pulling in data easily path here
setwd("C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/code")
```

# Load in functions
```{r}
source("../utils/create_map.R") ## For creating maps
source("../utils/calculate_vulnerability.R") ## For calculating vulnerability
source("../utils/calculate_simulated_vulnerability.R") ## For calculating simulated vulnerability (sensitivity analysis)
# source("../utils/sens_sensitivity_analysis.R")  ## Sensitivity analysis on the sensitivity measure (FIXIT: work in progress)
source("../utils/exp_sensitivity_analysis.R") ## Sensitivity analysis on the exposure measure
source("../utils/sens_ac_sensitivity_analysis.R") ## Sensitivity analysis on the exposure measure
source("../utils/number_unique_combs.R")  # Calculate the different unique combinations given the number of features (Used in sensitivity analysis)
```

# Instantiate plot themes

```{r}
# Shared base theme (mirrors theme_realm_panel style)
theme_bar <- function(base_size = 8,
                           axis_title_x = TRUE,
                           axis_title_y = FALSE,
                           legend       = TRUE) {
  base <- theme_minimal(base_size = base_size) +
    theme(
      text                    = element_text(family = family),
      plot.background         = element_rect(fill = "transparent", color = NA),
      panel.background        = element_rect(fill = "transparent", color = NA),
      panel.grid.minor        = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.3), linewidth = 0.15
      ),
      panel.grid.major.y = element_line(
        color = scales::alpha("black", 0.06), linewidth = 0.15
      ),
      axis.text               = element_text(
        size = base_size - 3, family = family
        ),
      axis.title.x            = element_text(
        size  = base_size,
        face  = "bold",
        vjust = -0.5,
        margin = margin(t = 4),
        family = family
      ),
      axis.title.y            = element_blank(),
      strip.text              = element_text(
        size = base_size - 1, face = "bold", family = family, color = "black"
        ),
      legend.position         = if (legend) "bottom" else "none",
      legend.title            = element_text(
        size = base_size - 1, face = "bold", family = family
        ),
      legend.text             = element_text(
        size = base_size - 2, family = family
        ),
      legend.key.size         = unit(0.3, "cm"),
      legend.box              = "vertical",
      legend.spacing          = unit(0.1, "cm"),
      legend.margin           = margin(t = 1, b = 0),
      legend.box.margin       = margin(t = -2, b = -2),
      legend.background       = element_rect(fill = "transparent", color = NA),
      legend.box.background   = element_rect(fill = "transparent", color = NA),
      plot.margin             = margin(3, 3, 3, 3)
    )

  if (!axis_title_x) {
    base <- base + theme(axis.title.x = element_blank())
  }

  base
}
```

# Geographic contribution to climate vulnerability

## Map of vulnerability

```{r}
# Continuous color scale
calculate_vulnerability(e_s_ac) %>%
  create_map(fill = "vulnerability", country.col.name = "consumer_iso3c") +
  scale_fill_gradient2(
    low = "#5E3C99",    # Purple — low vulnerability
    mid = "#F7F7F7",    # Near-white — midpoint
    high = "#E66101",   # Orange — high vulnerability
    midpoint = 0,
    name = "Climate vulnerability"
  )
  
# discrete color scale

# Create bounds that will be "low vulnerability", "moderate vulnerability",
# and "high vulnerability"
mod_vuln_lb <- vuln_df$vulnerability %>% quantile(0.33)
high_vuln_lb <- vuln_df$vulnerability %>% quantile(0.66)

calculate_vulnerability(e_s_ac) %>%
  mutate(
    vuln_type = case_when(
      vulnerability <= mod_vuln_lb ~ "low vulnerability",
      vulnerability >= mod_vuln_lb &
        vulnerability <= high_vuln_lb ~ "moderate vulnerability",
      vulnerability >= high_vuln_lb ~ "high vulnerability"
    ),
    vuln_type = factor(vuln_type, levels = c("low vulnerability",
                                          "moderate vulnerability",
                                          "high vulnerability"))
  ) %>%
  create_map(
    fill = "vuln_type",
    country.col.name = "consumer_iso3c"
  ) +
  scale_fill_manual(
    values = c(
      "low vulnerability" = "#FFF3B0",
      "moderate vulnerability" = "#F4A261",
      "high vulnerability" = "#C44536"
    ),
    breaks = c(
      "low vulnerability",
      "moderate vulnerability",
      "high vulnerability"
    ),
    labels = c(
      "low climate vulnerability",
      "moderate climate vulnerability",
      "high climate vulnerability"
    )
  ) +
  labs(fill = "") +
  theme(legend.key.spacing.x = unit(1, "cm"))
```

## How much does sensitivity, exposure, and adaptive capacity contribute to the cliamte vulnerability?

```{r load in bayesian 15sd gapfill}
# Read in e/s/ac (full portfolio)
e_s_ac <- read_csv("../output/e_s_ac/15_sd/e_s_ac_total_portfolio_bayesian_gapfill.csv")

# Read in e/s/ac (just foreign imports)
e_s_ac_foreign <- read_csv("../output/e_s_ac/15_sd/e_s_ac_total_portfolio_bayesian_gapfill_foreign.csv")

# Calculate the percentage that foreign consumption contributes to the total exposure
for_pct_contribution <- e_s_ac %>%
  filter(scenario == "ssp126") %>%
  left_join(
    
    e_s_ac_foreign %>%
      filter(scenario == "ssp126") %>%
      select(
        consumer_iso3c,
        pct_change
        ),
    by = c("consumer_iso3c")
    
  ) %>%
  # Calculate exposure from domestic sources
  mutate(domestic_prop_change = pct_change.x * 2 - pct_change.y,
         foreign_prop_change = pct_change.y) %>% 
  # Calculate the proportion contribution of the foreign exposure to the total exposure
  mutate(foreign_prop_contribution = (abs(foreign_prop_change) / (abs(foreign_prop_change) + abs(domestic_prop_change))),
         foreign_prop_contribution = round(foreign_prop_contribution, 1)) %>%
  select(consumer_iso3c, foreign_prop_contribution)

# Check to ensure that contribution ranges between 0 and 100%
all(
  for_pct_contribution$foreign_prop_contribution %>% round(1) %>% range() >= 0 &
      for_pct_contribution$foreign_prop_contribution %>% round(1) %>% range() <= 1
  )

# Calculate vulnerability, adaptive capacity, sensitivity, and exposure
# of foreign imports and domestic marine capture consumption
vuln_df <- calculate_vulnerability(e_s_ac) %>%
  # join on foreign contribution to total exposure
  left_join(for_pct_contribution) %>%
  mutate(foreign_exposure = exposure * foreign_prop_contribution,
         domestic_exposure = exposure * (1 - foreign_prop_contribution)) %>%
  mutate(adaptive_capacity = -adaptive_capacity) %>%
  pivot_longer(cols = c(foreign_exposure, domestic_exposure, sensitivity, adaptive_capacity))

# Helper to build a bar panel
make_bar_panel <- function(regions,
                           axis_x_label = FALSE,
                           legend_show  = FALSE,
                           text_size    = 4.75,
                           mod_vuln_lb  = -0.57942,
                           high_vuln_lb = -0.0830816) {

  df <- vuln_df %>% filter(region %in% regions)

  threshold_df <- tibble::tibble(
    xintercept = c(mod_vuln_lb, high_vuln_lb),
    threshold  = factor(c("Moderate vulnerability threshold",
                           "High vulnerability threshold"),
                         levels = c("Moderate vulnerability threshold",
                                    "High vulnerability threshold"))
  )

  ggplot(df) +
    geom_col(aes(x = value, y = fct_reorder(consumer_iso3c, -vulnerability), fill = name)) +
    geom_vline(data = threshold_df,
               aes(xintercept = xintercept, linetype = threshold),
               color = "grey30", linewidth = 0.4) +
    geom_point(aes(x = vulnerability, y = consumer_iso3c, color = "hi"), shape = 21,
               fill = "grey20", size = 1.2, stroke = 0.3) +
    scale_linetype_manual(
      name   = "",
      values = c("Moderate vulnerability threshold" = "dotted",
                 "High vulnerability threshold"      = "dashed")
    ) +
    scale_fill_manual(
      values = c(
        "domestic_exposure" = "#C44E52",
        "foreign_exposure"  = "#E6A23C",
        "sensitivity"       = "#756BB1",
        "adaptive_capacity" = "#4C9F9A"
      ),
      labels = c(
        "domestic_exposure" = "Exposure (domestic consumption)",
        "foreign_exposure"  = "Exposure (foreign consumption)",
        "sensitivity"       = "Sensitivity",
        "adaptive_capacity" = "Adaptive Capacity"
      )
    ) +
    labs(x = "Index value", y = "", fill = "Index", color = "") +
    facet_wrap(~ region, scales = "free_y", nrow = 1) +
    theme_bar(axis_title_x = axis_x_label, legend = legend_show) +
    scale_color_manual(
      values = "white",
      label = "Climate vulnerability",
      guide = guide_legend(
        order = 2,
        override.aes = list(
          fill = list(
            pattern(
              grobTree(circleGrob(x = 0.5, y = 0.5, r = 0.3, gp = gpar(fill = "grey20"))),
              extend = "repeat",
              gp = gpar(col = "grey20", fill = "white", lwd = 0.25)
            )
          ),
          color = "grey20",
          linewidth = 0.2,
          size = 3.5
        )
      )
    )
}


# Build panels - they are separated into two rows because including all in one
# would make the countries too small to see
all_regions   <- unique(vuln_df$region)
top_regions   <- all_regions[!all_regions %in% c("Oceania", "South America")]
bot_regions   <- c("Oceania", "South America")

# Top row of bars (i.e.. North America, Asia, Africa, Europe)
top_row <- make_bar_panel(
  top_regions,
  axis_x_label = FALSE,
  legend_show = FALSE,
  text_size = 4.75) +
  theme(plot.margin = margin(t = 0,  # Top margin
                             r = 5,  # Right margin
                             b = 0,  # Bottom margin
                             l = 5))

# Top row of bars (i.e.. Africa and Oceania)
bottom_row <- make_bar_panel(
  bot_regions,
  axis_x_label = TRUE,
  legend_show = TRUE,
  text_size = 5.15)

# Since there are less plots in the bottom row, add spacers in to make
# the aspect ratio of the bottom plots look similar in size to the top plots
bottom_row_narrowed <- cowplot::plot_grid(
  NULL, bottom_row, NULL,
  ncol       = 3,
  rel_widths = c(0.1, 0.3, 0.1)
) %>% suppressWarnings()

# combine the bottom and top bar plots (this will make up the bottom panel of
# the plot)
(final_plot <- cowplot::plot_grid(
  top_row,
  bottom_row_narrowed,
  nrow        = 2,
  rel_heights = c(2, 1.25)
) %>%
  suppressWarnings())

# Save plot
ggsave(
  filename = "../images/contribution_to_vulnerability.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 8,
  height   = 6
)
```

## What proportion of regions' consumption is consumed from for/dom and what proportion of that is projected to increase / decrease in catch?

```{r}
# Load in gap filled consumption data
proj_consumption <- read_parquet("../output/consumption_projections_gap_filled/bayesian/15_sd/consumption_projections_gap_filled_total_portfolio_unweighted_bayesian_gapfill.parquet")

# Proportion of regions consumption contributed to climate vulnerability
prop_for_dom_region <- proj_consumption %>%
  group_by(region, consumption_type) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  group_by(region) %>%
  mutate(prop_for_dom = live_weight_t / sum(live_weight_t)) %>%
  select(region, consumption_type, prop_for_dom)

# What proportion of regions' for/dom is projected to increase / decrease
# in catch?
prop_exposed <- proj_consumption %>%
  mutate(is_exposed = ssp126 < 0) %>%
  group_by(region, consumption_type, is_exposed) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  drop_na(is_exposed) %>%
  group_by(region, consumption_type) %>%
  mutate(prop_is_exposed = live_weight_t / sum(live_weight_t))  %>%
  left_join(prop_for_dom_region)

# Prepare data for the alluvial plot
sankey_data <- prop_exposed %>%
  mutate(
    source = case_when(
      consumption_type == "domestic" ~ "Domestic",
      consumption_type == "foreign"  ~ "Foreign"
    ),
    exposure = if_else(is_exposed, "Decrease", "Increase"),
    region = factor(
      region,
      levels = c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")
    ),
    source = factor(source, levels = c("Domestic", "Foreign")),
    exposure = factor(exposure, levels = c("Decrease", "Increase"))
  ) %>%
  group_by(region) %>%
  mutate(
    region_total_t = sum(live_weight_t),          # keep the real total for labelling
    y_plot = live_weight_t / region_total_t        # each region now sums to 1
  ) %>%
  ungroup()

sankey_data <- sankey_data %>%
  group_by(region) %>%
  mutate(region_label = paste0(
    as.character(region), "\n(",
    scales::label_number(scale = 1e-6, suffix = " Mt", accuracy = 0.1)(first(region_total_t)),
    ")"
  )) %>%
  ungroup() %>%
  mutate(region_fill = region)

sankey_long <- to_lodes_form(
  sankey_data,
  axes = c("region", "source", "exposure"),
  id = "alluvium"
)

# Publication-friendly colour palette
region_cols <- c(
  "Africa" = "#741A32",
  "Asia" = "#B34232",
  "Europe" = "#D38F35",
  "North America" = "#D4B95F",
  "Oceania" = "#4FA2A2",
  "South America" = "#114F59"
  
)


ggplot(
  sankey_long,
  aes(x = x, stratum = stratum, alluvium = alluvium, y = y_plot, fill = region_fill)
) +
  geom_flow(
    width = 0.18,
    alpha = 0.65,
    colour = "white",     # thin white border instead of NA — invisible against white background
    linewidth = 0.1,
    aes.bind = "flows"
  ) +
  geom_stratum(width = 0.23, fill = "white", colour = "grey30", linewidth = 0.35) +   # was 0.18
geom_text(stat = "stratum", aes(label = str_wrap(after_stat(stratum), width = 10)), size = 3.2) +
  scale_x_discrete(
    limits = c("region", "source", "exposure"),
    labels = c("Region", "Consumption source", "Projected percent\nchange in catch"),
    expand = c(0.08, 0.08)
  ) +
  scale_fill_manual(values = region_cols) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 11, colour = "black", margin = margin(t = 8)),
    axis.text.y = element_blank(),
    legend.position = "none",
    plot.margin = margin(10, 20, 10, 10)
  ) +
labs(x = "", y = "")
```

# Taxa contribution to vulnerability

## See what proportion of countries' consumption is exposed and not exposed

```{r}
# Load in gap filled consumption data
proj_consumption <- read_parquet("../output/consumption_projections_gap_filled/bayesian/15_sd/consumption_projections_gap_filled_total_portfolio_unweighted_bayesian_gapfill.parquet")

# Get the % change in catch by species
proj_change <- proj_consumption %>%
  distinct(sciname, eez_iso3c, ssp126)

proj_consumption %>%
  group_by(consumer_iso3c, sciname, eez_iso3c) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  left_join(proj_change) %>%
  mutate(is_exposed = ssp126 < 0) %>%
  group_by(consumer_iso3c, is_exposed) %>%
  drop_na() %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  group_by(consumer_iso3c) %>%
  mutate(perc_exposed = (live_weight_t / sum(live_weight_t)) * 100) %>%
  filter(is_exposed == TRUE) %>%
  create_map(fill = "perc_exposed", country.col.name = "consumer_iso3c") +
  labs(fill = "Percentage of consumption\nwith projected\ndeclines in catch ") +
  scale_fill_gradient2(low = "#FFF3B0",
                       mid = "#F4A261",
                       high = "#C44536",
                       midpoint = 50,
                       labels = scales::label_percent(scale = 1)) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(hjust = 0.5))
```


## See global top species and whether they are "exposed" or "not"

```{r}
proj_consumption <- read_parquet("../output/consumption_projections_gap_filled/bayesian/15_sd/consumption_projections_gap_filled_total_portfolio_unweighted_bayesian_gapfill.parquet")

# Check globally
# Get the pct_change by sciname
proj_change <- proj_consumption %>%
  distinct(sciname, eez_iso3c, ssp126)

# Find the top 5 taxa consumed
top_10_species_consumed <- proj_consumption %>%
  group_by(sciname) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  top_n(5) %>%
  pull(sciname)

# # For each species, calculate the proportion that one species is consumed
  # out of the top five species
prop_cons <- proj_consumption %>%
  group_by(sciname) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  top_n(5) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t)) %>%
  select(sciname, prop_consumption)

# See which EEZ
proj_consumption %>%
  group_by(sciname, eez_iso3c) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  left_join(proj_change) %>%
  filter(sciname %in% top_10_species_consumed) %>%
  mutate(is_exposed = ssp126 < 0) %>%
  group_by(sciname, is_exposed) %>%
  summarize(live_weight_t = sum(live_weight_t)) 

# See the proportion that countries' species are exposed or aren't exposed
proj_consumption %>%
  group_by(sciname, eez_iso3c) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  left_join(proj_change) %>%
  filter(sciname %in% top_10_species_consumed) %>%
  mutate(is_exposed = ssp126 < 0) %>%
  group_by(sciname, is_exposed) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  group_by(sciname) %>%
  mutate(prop_exposed = live_weight_t / sum(live_weight_t)) %>%
  # Join on consumption props by species (from prop_cons)
  left_join(prop_cons) %>%
  mutate(
    exposure = if_else(is_exposed, "Exposed", "Not exposed"),
    # Amount of total consumption flowing into each exposure category
    flow = prop_consumption * prop_exposed
  ) %>%

ggplot(
  aes(
    axis1 = sciname,
    axis2 = exposure,
    y = flow
  )
) +
  geom_alluvium(
    aes(fill = exposure),
    width = 0.15,
    alpha = 0.7
  ) +
  geom_stratum(
    width = 0.15,
    fill = "grey90",
    color = "grey30"
  ) +
  geom_text(
    stat = "stratum",
    aes(label = after_stat(stratum)),
    size = 3
  ) +
  scale_y_continuous(
    labels = percent_format()
  ) +
  scale_fill_manual(
    values = c(
      "Exposed" = "#D55E00",
      "Not exposed" = "#0072B2"
    )
  ) +
  labs(
    x = NULL,
    y = "Proportion of total consumption",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank()
  )
```

```{r}
proj_consumption <- read_parquet("../output/consumption_projections_gap_filled/bayesian/15_sd/consumption_projections_gap_filled_total_portfolio_unweighted_bayesian_gapfill.parquet")

# Check globally
# Get the pct_change by sciname
proj_change <- proj_consumption %>%
  distinct(sciname, region, eez_iso3c, ssp126)

# Find the top 5 taxa consumed
top_taxa_consumed <- proj_consumption %>%
  group_by(region, sciname) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  arrange(region, desc(live_weight_t), sciname) %>%
  group_by(region) %>%
  slice_head(n = 2) %>%
  ungroup() %>%
  select(region, sciname)  

# # For each species, calculate the proportion that one species is consumed
  # out of the top five species
prop_cons <- proj_consumption %>%
  group_by(region, sciname) %>%
  filter(sciname %in% top_taxa_consumed) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t)) %>%
  select(region, sciname, prop_consumption)

# See the proportion that countries' species are exposed or aren't exposed
sankey_data <- proj_consumption %>%
  group_by(region, sciname, eez_iso3c) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  left_join(proj_change) %>%
  semi_join(top_taxa_consumed, by = c("region", "sciname")) %>%   # <-- was filter(sciname %in% ...)
  mutate(is_exposed = ssp126 < 0) %>%
  group_by(region, sciname, is_exposed) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  group_by(region, sciname) %>%
  mutate(prop_exposed = live_weight_t / sum(live_weight_t)) %>%
  left_join(prop_cons) %>%
  mutate(
    exposure = if_else(is_exposed, "Decrease", "Increase"),
    flow = prop_consumption * prop_exposed
  )

region_cols <- c(
  "Africa" = "#741A32",
  "Asia" = "#B34232",
  "Europe" = "#D38F35",
  "North America" = "#D4B95F",
  "Oceania" = "#4FA2A2",
  "South America" = "#114F59"
)

sankey_data <- sankey_data %>%
  mutate(
    region = factor(
      region,
      levels = c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")
    ),
    region_fill = region      # duplicate BEFORE melting, so it survives to_lodes_form()
  ) %>%
  arrange(region, sciname, exposure)   # controls sub-ribbon stacking, avoids crossing artifacts

sankey_long <- to_lodes_form(
  sankey_data,
  axes = c("region", "sciname", "exposure"),
  id = "alluvium"
)

final_plot <- ggplot(
  sankey_long,
  aes(x = x, stratum = stratum, alluvium = alluvium, y = flow, fill = region_fill)
) +
  geom_flow(
    width = 0.18,
    alpha = 0.65,
    colour = "white",
    linewidth = 0.1,
    aes.bind = "flows"
  ) +
  geom_stratum(width = 0.23, fill = "white", colour = "grey30", linewidth = 0.35) +
  geom_text(
    stat = "stratum",
    aes(label = str_wrap(after_stat(stratum), width = 10)),
    size = 3.2
  ) +
  scale_x_discrete(
    limits = c("region", "sciname", "exposure"),
    labels = c("Region", "Top taxa conusmed", "Projected percent\nchange in catch"),
    expand = c(0.08, 0.08)
  ) +
  scale_fill_manual(values = region_cols) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 11, colour = "black", margin = margin(t = 8)),
    axis.text.y = element_blank(),
    legend.position = "none",
    plot.margin = margin(10, 20, 10, 10)
  ) +
  labs(x = "", y = "")

# Save plot
ggsave(
  filename = "../images/sankey_region_species.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 12,
  height   = 11
)
```

# Where are these species primarily coming from?

```{r}
top_taxa_consumed <- top_species_consumed %>%
  pull(sciname)

species_eez_data <- proj_consumption %>%
  filter(sciname %in% top_taxa_consumed) %>%
  group_by(sciname, eez_iso3c) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  arrange(-live_weight_t) %>%
  group_by(sciname) %>%
  slice_head(n = 2) %>%
  mutate(prop_consumption = live_weight_t / sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_fill = sciname) %>%      # duplicate BEFORE melting, so it survives to_lodes_form()
  arrange(sciname, desc(prop_consumption))


species_eez_long <- to_lodes_form(
  species_eez_data,
  axes = c("sciname", "eez_iso3c"),
  id = "alluvium"
)

# Sanity check before plotting
species_eez_long %>% filter(is.na(stratum) | is.na(prop_consumption)) %>% nrow()  # should be 0
artis_
# Colour palette — one colour per species (swap for your preferred palette)
species_cols <- c(
  "actinopterygii" = "#741A32",
  "dosidicus gigas" = "#A83B32",
  "engraulis japonicus" = "#C87534",
  "engraulis ringens" = "#D3A349",
  "gadus chalcogrammus" = "#A7B175",
  "macruronus novaezelandiae" = "#449495",
  "sardina pilchardus" = "#114F59"
  # add remaining species here
)

final_plot <- ggplot(
  species_eez_long,
  aes(x = x, stratum = stratum, alluvium = alluvium, y = prop_consumption, fill = sciname_fill)
) +
  geom_flow(width = 0.18, alpha = 0.65, colour = "white", linewidth = 0.1, aes.bind = "flows") +
  geom_stratum(width = 0.23, fill = "white", colour = "grey30", linewidth = 0.35) +
  # Species labels: default centered behavior, no hjust override
  geom_text(
    data = species_eez_long %>% filter(x == "sciname"),
    stat = "stratum",
    aes(label = str_wrap(after_stat(stratum), width = 12)),
    size = 3.2
  ) +
  # EEZ labels: pushed right, outside the block, to avoid overlap on small strata
  geom_text(
    data = species_eez_long %>% filter(x == "eez_iso3c"),
    stat = "stratum",
    aes(label = after_stat(stratum)),
    size = 3.2,
    hjust = -0.15
  ) +
  scale_x_discrete(
    limits = c("sciname", "eez_iso3c"),
    labels = c("Top species", "EEZ"),
    expand = c(0.1, 0.18)   # less padding on left (species side), more on right (EEZ side)
  ) +
  scale_fill_manual(values = species_cols) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 11, colour = "black", margin = margin(t = 8)),
    axis.text.y = element_blank(),
    legend.position = "none",
    plot.margin = margin(10, 40, 10, 10)   # extra room on the right for outward EEZ labels
  ) +
  labs(x = "", y = "")

# Save plot
ggsave(
  filename = "../images/sankey_species_eez.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 12,
  height   = 11
)
```

