---
title: "Validation Checks"
author: "Connor Quiroz"
date: "2026-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)
library(devtools)
library(tidytext)
library(ggsankey)
library(exploreARTIS)

# remotes::install_github("ropensci/rfishbase")
library(rfishbase)
```

```{r load in functions}
# Loads functions needed to import SAU dataall helper functions into this session
knitr::knit("00_Preprocess_Data.Rmd", output = tempfile())
```

```{r user-input}
# Set your path here
file_path <- "C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/data/"

# Jessica's path:
# file_path <- ""

dir <- file_path
```

# Purpose

This file is for testing and analyzing data for quality checks that may not go in the final code

# Sea Around Us (SAU) maximum catch potential (mcp) projection data assessment

## Is Juliano's data complete? And if not, what's missing from the data

```{r}
# Load in file names and species names acquired from the sau mcp data
file_names <- readRDS("../data/exposure/utils/file_names.rds")
species_names <- readRDS("../data/exposure/utils/species_names.rds")

# ---- File integrity diagnostic ----

filename_keys <- as.integer(str_extract(basename(file_names), "\\d+"))

# Scan all files once, collecting all needed info in a single pass
file_scan <- map_dfr(file_names, function(f) {
  df_i <- fread(f, colClasses = list(character = c("eez_name", "ssp")))
  filename_key <- as.integer(str_extract(basename(f), "\\d+"))
  
  if (nrow(df_i) == 0) {
    return(tibble(filename_key = filename_key, actual_key = NA_integer_, n_rows = 0L, found_in_file = basename(f)))
  }
  
  tibble(filename_key = filename_key, actual_key = unique(df_i$taxon_key), n_rows = nrow(df_i), found_in_file = basename(f))
})

# No file at all
no_file <- species_names %>% 
  filter(!taxon_key %in% filename_keys) %>%
  as_tibble() %>%
  mutate(issue = "no file", found_in_file = NA_character_)

# Empty files
empty_files <- file_scan %>%
  filter(n_rows == 0) %>%
  left_join(as_tibble(species_names), by = c("filename_key" = "taxon_key")) %>%
  mutate(issue = "empty file") %>%
  select(taxon_key = filename_key, taxon_name, issue, found_in_file)

# Mismatched keys — also check if no_file species are hiding in these files
mismatched <- file_scan %>%
  filter(n_rows > 0, filename_key != actual_key) %>%
  left_join(as_tibble(species_names), by = c("actual_key" = "taxon_key")) %>%
  mutate(issue = "mismatched key") %>%
  select(taxon_key = actual_key, taxon_name, issue, found_in_file)

# Check if any no_file species are hiding inside mismatched files
no_file_found <- no_file %>%
  filter(taxon_key %in% mismatched$taxon_key) %>%
  left_join(mismatched %>% select(taxon_key, found_in_file), by = "taxon_key") %>%
  mutate(issue = "no file — found in mismatched file")

no_file_not_found <- no_file %>%
  filter(!taxon_key %in% mismatched$taxon_key)

# Get the files which have no data themselves and contain files from another species id
mismatched_origin_file <- gsub(c("percentage_change_"), "", mismatched$found_in_file) 
mismatched_origin_file <- as.numeric (gsub(".csv", "", mismatched_origin_file))

mismatched_origin_file <- species_names %>%
  filter(
    taxon_key %in% 
           species_names$taxon_key[species_names$taxon_key %in% mismatched_origin_file]
  )

# Correct files
correct <- species_names %>%
  as_tibble() %>%
  filter(
    taxon_key %in% filename_keys,
    !taxon_key %in% empty_files$taxon_key,
    !taxon_key %in% mismatched$taxon_key,
    !taxon_key %in% mismatched_origin_file$taxon_key
  ) %>%
  mutate(issue = "correct", found_in_file = NA_character_)

# Summary
cat("Exposure data quality__________________________________________________________________", "\n")
cat("No file (truly missing):                                                           ", nrow(no_file_not_found), "\n")
cat("File unavailable, but data found in other file:                                    ", nrow(mismatched), "\n")
cat("File available, but contains another species data (original data is nowhere else): ", nrow(mismatched_origin_file), "\n")
cat("Empty file:                                                                        ", nrow(empty_files), "\n")
cat("Incorrect:                                                                         ", nrow(no_file_not_found) + nrow(no_file_found) + nrow(empty_files) + nrow(mismatched) + nrow(mismatched_origin_file), "\n")
cat("Correct:                                                                           ", nrow(correct), "\n")
cat("Total accounted for:                                                               ", nrow(no_file_not_found) + nrow(no_file_found) + nrow(empty_files) + nrow(mismatched) + nrow(mismatched_origin_file) + nrow(correct), "\n")
cat("Total in species_names:                                                            ", nrow(species_names), "\n")

# Full breakdown table
# bind_rows(no_file_not_found, no_file_found, empty_files, mismatched, correct) %>%
#   select(taxon_key, taxon_name, issue, found_in_file) %>%
#   arrange(issue, taxon_key)
```

> Currently of the `r nrow(species_names)` species Juliano sent, there are `r nrow(no_file_not_found)` truly missing, `r nrow(empty_files)` with empty files, `r nrow(mismatched)` with mismatched keys, and `r nrow(correct)` species 




```{r Jessica / Connor Meeting }
# Join consumption data to catch projections
  consumption_mpc <- consumption %>%
    left_join(sciname_taxa, by = c("sciname" = "sciname")) %>%
    left_join(climate_by_spp_eez, by = c("eez_iso3c", "sciname" = "sciname"))


consumption_mpc %>% filter(str_detect(sciname, " "),
                           is.na(ssp126))

# Mismatches between species names 
consumption_mpc %>% filter(str_detect(sciname, " "),
                           is.na(ssp126)) %>%
  group_by(sciname, eez_iso3c) %>%
  summarize(sum = sum(live_weight_t)) %>%
  arrange(-sum)

consumption_mpc  %>%
  filter(sciname == "scomber scombrus",
         eez_iso3c == "GBR")
  distinct(sciname, eez_iso3c)
```


## How many of Juliano's current species names match with ARTIS naming?
```{r assess species alignment of SAU data to ARTIS}
# Load in data

## Filter ARTIS consumption data to only species (2019 marine capture) 
consumption <- read_parquet(paste0(file_path, "example_consumption_eez_2024_12_06.parquet")) %>%
    filter(year == 2019, habitat == "marine", method == "capture")

## Read in ARTIS Sciname data
  sciname_taxa <- read_csv(paste0(dir, "sciname.csv"), show_col_types = FALSE)
    # build_taxa_levels()

## Load in future catch projections data
### For each species, average mcp projections across territories with an EEZ
  climate_by_spp_eez <- read_csv("../output/future_climate_joined.csv", show_col_types = FALSE) %>%
    drop_na() %>%
    group_by(eez_iso3c, sciname) %>%
    summarise(across(c(ssp126, ssp585), ~mean(.x, na.rm = TRUE)), .groups = "drop")

## Get list of unique species in 2019 marine capture ARTIS consumption data
consumption_species_list <- consumption %>% # Filter to only species names
  distinct(sciname)

## Get list of unique species in sciname
sciname_species <- sciname_taxa %>% filter(str_detect(sciname, " ")) %>% pull(sciname)


# Data assessment
## Which of Juliano's species are not matching to ARTIS?
unaligned_names <- correct$taxon_name[!str_to_lower(correct$taxon_name) %in% consumption_species_list$sciname]

# Correct Juliano's species names that aren't aligned with Fishbase / Sealifebase

## Fishbase corrections
fb_corrections <- validate_names(species_list = unaligned_names, serve = "fishbase", version = "24.07") %>% suppressMessages()

## SeaLifeBase corrections
slb_corrections <- validate_names(species_list = unaligned_names, serve = "sealifebase", version = "24.07") %>% suppressMessages()

## Join corrections together
corrections <- na.omit(fb_corrections) %>% append(na.omit(slb_corrections))

##### Create lookup table
  ##### (str_to_lower had to be done here becaue fb/slb lookup tables are uppercase)
  correction_lookups <- setNames(corrections %>% str_to_lower(), unaligned_names %>% str_to_lower())

# Correct catch projection data
climate_by_spp_eez_corrected <- climate_by_spp_eez %>%
  mutate(sciname = ifelse(
    sciname %in% names(correction_lookups),
    correction_lookups[sciname],
    NA
  )) %>%
    drop_na(sciname)

climate_by_spp_eez_combined <- bind_rows(
  climate_by_spp_eez, climate_by_spp_eez_corrected
  ) %>%
    group_by(eez_iso3c, sciname) %>%
    summarise(across(c(ssp126, ssp585), ~mean(.x, na.rm = TRUE)), .groups = "drop") ### For each species, average mcp projections across territories with an EEZ

# Assess data coverage improvements post naming corrections

## Which Juliano's *corrected* species are not matching to ARTIS consumption 2019 data?
num_sp_post_corrections_cons <- corrections[!str_to_lower(corrections) %in% consumption$sciname] %>% unique() %>% length()

## Which Juliano's *corrected* species are not matching to ARTIS sciname data?
num_sp_post_corrections_sciname <- corrections[!str_to_lower(corrections) %in% sciname_species] %>% length()

## See which values in my vector are equal to one another
  for (i in 1:length(corrections)) {
     z <- which(corrections[i] == corrections)
     current_idx <- 0
     
     if (length(z) > 1) {
       idx <- z
        if (all(idx != current_idx)) { # only run if there is a new index found (this doesn't work, doesn't need to work, just run if there's a match)
         cat("Species that are matched to the same corrected name: ", unaligned_names[idx])
         cat("\n")
         cat("Species correction destination: ", corrections[idx[1]], "\n")
        }
       # current_idx <- idx
     }
     
    
  }

# Join consumption data to catch projections
  consumption_mpc <- consumption %>%
    left_join(sciname_taxa, by = c("sciname" = "sciname")) %>%
    left_join(climate_by_spp_eez_combined, by = c("eez_iso3c", "sciname" = "sciname"))

# Totals: What After naming correction methods, how many species do Juliano's cover?

## Subtract number of naming errors (i.e., species that were not joined to ARTIS) from the files with upfront "no data value / file errors" to get the number of Juliano's correct species that originally join to ARTIS
cat("Number of Juliano's species that correctly join to ARTIS (before correction): ", nrow(correct) - length(unique(corrections)))

cat("Number of Juliano's species that correctly join to ARTIS (before correction): ", nrow(correct) - num_sp_post_corrections_cons)

cat("Number of species that were added into 2019 consumption data: ", length(unique(corrections)) - num_sp_post_corrections_cons, "\n")

# Total rows: How many are missing?

## Number of missing rows from uncorrected data
num_rows <- nrow(consumption %>% distinct(eez_iso3c, sciname)) - anti_join(consumption, climate_by_spp_eez, by = c("eez_iso3c", "sciname")) %>% distinct(eez_iso3c, sciname) %>% nrow()

cat("Number of rows filled in (no corrections): ", num_rows, "; ", round(num_rows / nrow(consumption %>% distinct(eez_iso3c, sciname)) * 100,1), "% data coverage of ", consumption %>% distinct(sciname, eez_iso3c) %>% nrow(), " total pairs" )

## Number of missing rows from corrected data
num_rows <- nrow(consumption %>% distinct(eez_iso3c, sciname)) - anti_join(consumption, climate_by_spp_eez_combined, by = c("eez_iso3c", "sciname")) %>% distinct(eez_iso3c, sciname) %>% nrow()
 
cat("Number of rows filled in (no corrections): ", num_rows, "; ", round(num_rows / nrow(consumption %>% distinct(eez_iso3c, sciname)) * 100,1), "% data coverage of ", consumption %>% distinct(sciname, eez_iso3c) %>% nrow(), " total pairs" )
```

```{r Testing gapfill function}
# Helper: for each higher taxonomic level, compute mean SSP values per EEZ
# so that species with missing projections can borrow from relatives
build_taxa_lookup <- function(df) {
  tax_levels <- c("genus", "subfamily", "family", "order", "class", "superclass","phylum")

  unique_climate <- df %>%
    select(eez_iso3c, sciname, ssp126, ssp585,
           genus, subfamily, family, order, class, phylum, kingdom, superclass) %>%
    distinct()

  # For each taxa in an EEZ, compute the average ssp126 / 585 value for all species within in the taxa
  
  # Alternate: Build averages by region to gapfill if there is no neighbor found within an EEZ
  eez_gapfills <- map_dfr(tax_levels, function(level) {
    unique_climate %>%
      filter(!is.na(ssp126) | !is.na(ssp585)) %>%
      distinct(eez_iso3c, sciname, .keep_all = TRUE) %>%
      group_by(eez_iso3c, !!sym(level)) %>%
      summarise(
        eez_avg_ssp126 = mean(ssp126, na.rm = TRUE),
        eez_avg_ssp585 = mean(ssp585, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      rename(taxa_name = !!sym(level)) %>%
      mutate(taxa_level = level)
  })
  
  region_gapfills <- map_dfr(tax_levels, function(level) {
    unique_climate %>%
      filter(!is.na(ssp126) | !is.na(ssp585)) %>%
      distinct(eez_iso3c, sciname, .keep_all = TRUE) %>%
      add_region("eez_iso3c", "region") %>%
      group_by(region, !!sym(level)) %>%
      summarise(
        region_avg_ssp126 = mean(ssp126, na.rm = TRUE),
        region_avg_ssp585 = mean(ssp585, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      rename(taxa_name = !!sym(level)) %>%
      mutate(taxa_level = level)
  })
  
  return(list(
    eez_gapfills = eez_gapfills,
    region_gapfills = region_gapfills
  ))
      
}

# Helper: walk up the taxonomic ladder and fill NA SSP values from the nearest
# available ancestor within the same EEZ
fill_with_higher_taxa <- function(df, lookup_table) {
  tax_levels <- c("genus", "subfamily", "family", "order", "class", "superclass", "phylum")

  df <- df %>% mutate(gapfill_level = NA_character_, gapfill_method = NA_character_)
  
  # Run through EEZ averaged gapfilling first
  for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level)

    df <- df %>%
      select(-taxa_level) %>%
      ## Join data to EEZ interpolated data
      left_join(
        lookup_table$eez_gapfills %>% filter(taxa_level == level),
        by = c("eez_iso3c", join_by_map)
      ) %>%
      ## Conduct EEZ gapfills
      ### Only assign the gapfill method if there was actually a gapfill occurring
      ### Only assign the taxonomic level for which the gapfill was performed at
      ### if a gapfill occurred
      mutate(
        gapfill_level = case_when( ## Define at which taxa level gap filled data was gotten from
          is.na(ssp126) & !is.na(eez_avg_ssp126) ~ level, # Data coverage between ssp126/585 are the same, so we only need to check one variable
          TRUE                                              ~ gapfill_level
        ),
        gapfill_method = case_when(
          is.na(ssp126) & !is.na(eez_avg_ssp126) ~ "eez_gapfill",
          TRUE                                              ~ gapfill_method
        ),
        ssp126 = coalesce(ssp126, eez_avg_ssp126),
        ssp585 = coalesce(ssp585, eez_avg_ssp585),
      ) %>%
      select(-eez_avg_ssp126, -eez_avg_ssp585)
  }
  
  # Run through Region averaged gapfilling second
  
  for (level in tax_levels) {
    join_by_map <- setNames("taxa_name", level)
    
    if (level == "genus") { # Only do on first loop iteration
      df <- df %>%
        add_region("eez_iso3c", "region")
    }
    
    df <- df %>%
      select(-taxa_level) %>%
    # Join data to Region interpolated data
      left_join(
        lookup_table$region_gapfills %>% filter(taxa_level == level),
        by = c("region", join_by_map)
      ) %>%
        mutate(
        gapfill_level = case_when( ## Define at which taxa level gap filled data was gotten from
          is.na(ssp126) & !is.na(region_avg_ssp126) ~ level, # Data coverage between ssp126/585 are the same, so we only need to check one variable
          TRUE                                              ~ gapfill_level
        ),
        gapfill_method = case_when(
          is.na(ssp126) & !is.na(region_avg_ssp126) ~ "regional_gapfill",
          TRUE                                              ~ gapfill_method
        ),
        ssp126 = coalesce(ssp126, region_avg_ssp126),
        ssp585 = coalesce(ssp585, region_avg_ssp585),
      ) %>%
      select(-region_avg_ssp126, -region_avg_ssp585)  
  }
  
  # Post gapfilling, delineate whether the gapfilling
  # was not needed or was unable to e performed
  
  df <- df %>%
    mutate(gapfill_method = case_when(
      !is.na(ssp126) & is.na(gapfill_method) ~ "Gapfill not needed",
      is.na(ssp126) ~ "Unsuccessful gapfill",
      TRUE ~ gapfill_method
    ),
    gapfill_level = case_when(
      !is.na(ssp126) & is.na(gapfill_level) ~ "Gapfill not needed",
      is.na(ssp126) ~ "Unsuccessful gapfill",
      TRUE ~ gapfill_level
    )
    )
  
  # Perform final data cleaning
  df %>% select(-taxa_level) %>%
    mutate(gapfill_level = factor(gapfill_level, levels = c("genus", "subfamily", "family", "class", "order", "phylum")))
  
}

# Build test interpolated data
test_interpolated <- fill_with_higher_taxa(consumption_mpc, build_taxa_lookup(consumption_mpc))

# What is the data coverage of the interpolated data?

## Calculate the total number of rows in consumption data
total_rows <- nrow(consumption_mpc)

## Get the number of interpolatd rows
interpolated_rows <- test_interpolated %>%
  # filter(str_detect(sciname, " ")) %>%
  filter(!is.na(ssp126)) %>%
  nrow()


# Decide on gapfill method
# Check compare the variance between individual EEZ's and at the region level before vs after gapfilling?

## Before gapfilling

### EEZ
consumption_mpc %>%
  group_by(eez_iso3c, taxa_level) %>%
  summarize(var = var(ssp126, na.rm = TRUE)) %>%
  add_region("eez_iso3c", "region") %>%
  filter(taxa_level == "species") %>%
  ggplot(aes(x = eez_iso3c, y = var)) +
  geom_col() +
  facet_wrap(~ region, scales = "free") +
  theme(axis.text = element_text(size = 5, angle = 90)) +
  labs(x = "EEZ iso3c", y = "Variance in Maximum catch potential across species")

### Region level
consumption_mpc %>%
  add_region("eez_iso3c", "region") %>%
  group_by(region, taxa_level) %>%
  summarize(var = var(ssp126, na.rm = TRUE)) %>%
  filter(taxa_level == "species") %>%
  ggplot(aes(x = region, y = var)) +
  geom_col() +
  labs(x = "EEZ iso3c", y = "Variance in Maximum catch potential across species")

## Check variances after gapfilling

### Interpolate using regional taxa data
test_interpolated %>%
  filter(gapfill_method == "eez_gapfill") %>%
  group_by(eez_iso3c, gapfill_level) %>%
  summarize(var = var(ssp126, na.rm = TRUE)) %>%
  add_region("eez_iso3c", "region") %>%
  ggplot(aes(x = eez_iso3c, y = var, fill = gapfill_level)) +
  geom_col() +
  facet_wrap(~ region, scales = "free") +
  theme(axis.text = element_text(size = 5, angle = 90)) +
  labs(x = "EEZ iso3c", y = "Variance in Maximum catch potential across species") +
  scale_fill_viridis_d(end = 0.95) +
  facet_wrap(~ region, scales = "free")

### Interpolate using regional taxa data
test_interpolated %>%
  group_by(eez_iso3c, gapfill_level, gapfill_method) %>%
  summarize(var = var(ssp126, na.rm = TRUE)) %>%
  filter(gapfill_method == "regional_gapfill") %>%
  add_region("eez_iso3c", "region") %>%
  ggplot(aes(x = eez_iso3c, y = var, fill = gapfill_level)) +
  geom_col() +
  # facet_wrap(~ region, scales = "free") +
  theme(axis.text = element_text(size = 5, angle = 90)) +
  labs(x = "EEZ iso3c", y = "Variance in Maximum catch potential across species") +
  scale_fill_viridis_d(end = 0.95) +
  facet_wrap(~ region, scales = "free")
  
  # How much gapfilling was used at each taxonomic level?
  (plot <- consumption_mpc_realm %>%
    # filter(gapfill_method %in% c("eez_gapfill", "regional_gapfill")) %>%
    mutate(
      gapfill_method_level = case_when(
        gapfill_method == "eez_gapfill" & gapfill_level == "subfamily" ~ "EEZ: Subfamily",
        gapfill_method == "eez_gapfill" & gapfill_level == "genus" ~ "EEZ: Genus",
        gapfill_method == "eez_gapfill" & gapfill_level == "family" ~ "EEZ: Family",
        gapfill_method == "eez_gapfill" & gapfill_level == "order" ~ "EEZ: Order",
        gapfill_method == "eez_gapfill" & gapfill_level == "class" ~ "EEZ: Class",
        gapfill_method == "eez_gapfill" & gapfill_level == "phylum" ~ "EEZ: Phylum",
        
        gapfill_method == "realm_gapfill" & gapfill_level == "subfamily" ~ "Realm: Subfamily",
        gapfill_method == "realm_gapfill" & gapfill_level == "genus" ~ "Realm: Genus",
        gapfill_method == "realm_gapfill" & gapfill_level == "family" ~ "Realm: Family",
        gapfill_method == "realm_gapfill" & gapfill_level == "order" ~ "Realm: Order",
        gapfill_method == "realm_gapfill" & gapfill_level == "class" ~ "Realm: Class",
        gapfill_method == "realm_gapfill" & gapfill_level == "phylum" ~ "Realm: Phylum",
        gapfill_method == "Gapfill not needed" ~ "No gapfill needed",
        gapfill_method == "Unsuccessful gapfill" ~ "Unsuccessful gapfill"
      ),
      gapfill_method_level = factor(
        gapfill_method_level,
        levels = c(
          "EEZ: Genus",
          "EEZ: Subfamily",
          "EEZ: Family",
          "EEZ: Order",
          "EEZ: Class",
          "EEZ: Phylum",
          "Realm: Genus",
          "Realm: Subfamily",
          "Realm: Family",
          "Realm: Order",
          "Realm: Class",
          "Realm: Phylum",
          "No gapfill needed",
          "Unsuccessful gapfill"
        )
      )
    ) %>%
    group_by(eez_iso3c, gapfill_method_level) %>%
    # summarize(n = n_distinct(sciname), .groups = "drop") %>%
    group_by(eez_iso3c) %>%
    mutate(prop = n / sum(n)) %>%
    add_region("eez_iso3c", "region") %>%
    ggplot(aes(y = gapfill_method_level, x = eez_iso3c, fill = prop)) +
    geom_tile() +
    scale_fill_viridis_c() +
    labs(x = "EEZ iso3c", fill = "Proportion") +
    facet_grid(. ~ region, scales = "free_x", space = "free_x") +
    theme_cowplot() +
    theme(legend.position = "bottom",
            axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 5.4),
          strip.text.x = element_text(size = 8),
         legend.key.size       = unit(.6, "cm"),
         legend))
    
  
  
  # Save plot
    ggsave(
      filename = paste0("../images/exploratory/gapfilling_coverage_region", ".jpg"),
      plot     = plot,
      device   = "jpg",
      width    = 15,
      height   = 6,
      units    = "in",
      dpi = 1000
    )

  
# Add in superclass grouping
### Varianes appear similar, so I will move forward with interpolating via region.

    sciname_taxa %>% names()
    
    
    x <- consumption_mpc_realm %>%
      mutate(
      gapfill_method_level = case_when(
        gapfill_method == "eez_gapfill" & gapfill_level == "subfamily" ~ "EEZ: Subfamily",
        gapfill_method == "eez_gapfill" & gapfill_level == "genus" ~ "EEZ: Genus",
        gapfill_method == "eez_gapfill" & gapfill_level == "family" ~ "EEZ: Family",
        gapfill_method == "eez_gapfill" & gapfill_level == "order" ~ "EEZ: Order",
        gapfill_method == "eez_gapfill" & gapfill_level == "class" ~ "EEZ: Class",
        gapfill_method == "eez_gapfill" & gapfill_level == "phylum" ~ "EEZ: Phylum",
        
        gapfill_method == "realm_gapfill" & gapfill_level == "subfamily" ~ "Realm: Subfamily",
        gapfill_method == "realm_gapfill" & gapfill_level == "genus" ~ "Realm: Genus",
        gapfill_method == "realm_gapfill" & gapfill_level == "family" ~ "Realm: Family",
        gapfill_method == "realm_gapfill" & gapfill_level == "order" ~ "Realm: Order",
        gapfill_method == "realm_gapfill" & gapfill_level == "class" ~ "Realm: Class",
        gapfill_method == "realm_gapfill" & gapfill_level == "phylum" ~ "Realm: Phylum",
        gapfill_method == "Gapfill not needed" ~ "No gapfill needed",
        gapfill_method == "Unsuccessful gapfill" ~ "Unsuccessful gapfill"
      ),
      gapfill_method_level = factor(
        gapfill_method_level,
        levels = c(
          "EEZ: Genus",
          "EEZ: Subfamily",
          "EEZ: Family",
          "EEZ: Order",
          "EEZ: Class",
          "EEZ: Phylum",
          "Realm: Genus",
          "Realm: Subfamily",
          "Realm: Family",
          "Realm: Order",
          "Realm: Class",
          "Realm: Phylum",
          "No gapfill needed",
          "Unsuccessful gapfill"
        )
      )) %>%
      group_by(gapfill_method_level, sciname) %>%
      summarize(cv = mean(ssp126, na.rm = TRUE) / sd(ssp126, na.rm = TRUE)) %>%
      drop_na()
    
    x 
      mutate()
      ggplot(aes(aes(x = gapfill_method_level, y = cv))) +
      geom_col()
```

# Check coverage of interpolated consumption data

```{r}
consumption_mpc_realm <- read_parquet("../output/future_catch_joined_consumption/interpolated_future_data_realm_gapfill.parquet")

consumption_mpc_region <- read_parquet("../output/future_catch_joined_consumption/interpolated_future_data_region_gapfill.parquet")

consumption_mpc_realm %>%
  filter(gapfill_method != "Gapfill not needed") %>%
  group_by(eez_iso3c, sciname, gapfill_method) %>%
  summarize(live_weight_t = sum(live_weight_t), .groups = "drop") %>%
  group_by(eez_iso3c, sciname) %>%
  slice_max(live_weight_t, n = 3) %>%
  arrange(desc(live_weight_t))


consumption_mpc_realm %>%
  group_by(gapfill_method) %>%
  summarize(live_weight_t = sum(live_weight_t)) %>%
  arrange(-live_weight_t)

consumption_mpc_region %>%
  group_by(gapfill_method) %>%
  summarize(live_weight_t = round(sum(live_weight_t),0)) %>%
  arrange(-live_weight_t)
  

```

