---
title: "01_Analysis"
author: "Connor Quiroz"
date: "2025-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# General data cleaning + visualization + analysis packages
library(tidyverse)
library(ggformula)
library(countrycode)
library(data.table)
library(arrow)
library(cowplot)
library(RColorBrewer)

# Map creation
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(dirmult)

# Load in packages needed for ARTIS
library(devtools)
library(tidytext)
# devtools::install_github("davidsjoberg/ggsankey")
library(ggsankey)
# devtools::install_github("Seafood-Globalization-Lab/exploreARTIS@v1.0.0", dependencies = TRUE)
library(exploreARTIS)
```

# Objectives

The purpose of this 01_Analysis is to:

- **Run a data acquisition script** that gathers  
  - Exposure data  
  - Sensitivity data  
  - Adaptive capacity data  

- **Calculate climate risk** for each country's imports and overall seafood portfolio  
  using the combined exposure, sensitivity, and adaptive capacity datasets.

- **Perform a sensitivity analysis** on adaptive capacity components  
  to ensure no single variable has a disproportionate effect on total risk.

> 💡 **User input:**  
> In the code chunk below, replace `file_path` with the path to your input data folder.  
> Once specified, the script will automatically execute all subsequent steps.


```{r initialize functions}
source("00_Preprocess_Data.R") # function that runs a script to acquire exposure, sensitivity, and adaptive capacity data
source("../R/create_map.R") # function for creating maps
source("../R/calculate_risk.R") # function for calculating risk
# source("../R/sensitivity_analysis.R") # function for running a sensitivity analysis on adaptive capacity variables (Work in Progress)
```

## Calculate the risk index based on a country’s exposure, sensitivity, and adaptive capacity components.

> Insert your path to get to your data file (line)

```{r calculate risk}
# Change this line based on user
# Connor's desktop path
# file_path <- "C:/Users/cjqui/OneDrive/Desktop/2024-2025/SGL/Chapter 2/data/"
# Connor's laptop path
file_path <- "C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/data/"

# Jessica's path (insert below)
# file_path <- ""


# Get exposure, sensitivty, and adaptive capacity data for entire portfolio
e_s_ac_data_entire_portfolio <- acquire_e_s_ac_data(foreign = FALSE, directory_path = file_path)

# Get exposure, sensitivty, and adaptive capacity data for foreign imports
e_s_ac_data <- acquire_e_s_ac_data(foreign = TRUE, directory_path = file_path)

# Calculate risk
risk_overall <- calculate_risk(e_s_ac_data_entire_portfolio, coverage = "entire portfolio")

risk_foreign_imports <- calculate_risk(e_s_ac_data, coverage = "foreign imports")
  
# Look at correlations between vulnerability and E + S + AC
pairs(risk_foreign_imports[,2:5])
cor_vals <- cor(risk_foreign_imports[,2:5])
```

## Sensitivity analysis on adaptive capacity variables and components

> Note: This is still a work in progress, so results are likely not be reliable at this stage.

```{r run sensitivity analysis}
# # Run sensitivity analysis on foreign import risk
# sensitivity_analysis(data = e_s_ac_data, risk_data = risk_foreign_imports)
# 
# # Run sensitivity analysis on entire portfolio risk
# sensitivity_analysis(data = e_s_ac_data_entire_portfolio, risk_data = risk_overall)
```


```{r}
# Read in consumption file
consumption <- read_parquet("../data/example_consumption_eez_2024_12_06.parquet") %>%
  filter(year == 2019) %>%
  group_by(consumer_iso3c, sciname_hs_modified) %>%
  summarize(live_weight_t = sum(live_weight_t))

# Read in future change catch % data
future_changes <- read_parquet("../output/interpolated_future_data.parquet")

# Read in sciname data
sciname <- read_csv(paste0(file_path,"sciname.csv"))

consumption %>%
  left_join(future_changes %>%
              select(consumer_iso3c, sciname_hs_modified, ssp126) %>% 
              distinct() %>%
              filter(!is.na(ssp126)),
            by = c("consumer_iso3c", "sciname_hs_modified"))

# Join dataset to generalized species groups names (actinopterygii --> fish)
# joined_dataset <- consumer_stock_change_pre %>%
#   left_join(sciname %>% select(sciname, isscaap), by = c("sciname_hs_modified" = "sciname")) %>%
#   ungroup() %>%
#   mutate(marginal_weight_change = change_in_stock - live_weight_t) %>%
#   group_by(consumer_iso3c, scenario) %>%
#   mutate(total_live_weight_t = sum(live_weight_t))


# Join species % change to total % change to calculate marginal cointribution of each species to total % change in consumption by species
# joined_dataset <- joined_dataset %>%
#               mutate(marginal_prop = marginal_weight_change / total_live_weight_t)
#
# # # Convert negative/positive percents to all positive proportons
# # joined_dataset <- joined_dataset %>%
# #   ungroup() %>%
# #   mutate(pct_change = abs((pct_change / 100)),
# #          total_pct_change = abs((total_pct_change  / 100)),
# #          marginal_change = pct_change / total_pct_change)
#   # Calculate how much each species % change contributes to the total % change
#
#
# joined_dataset %>% filter(consumer_iso3c == "USA") %>%
# ggplot(aes(x = log(live_weight_t), y = marginal_prop)) +
#   geom_point()
#
#
# # Look at rate of change between future change in catch % and countries' consumption by species
# joined_dataset %>%
#   filter(consumer_iso3c == "CHN", scenario == "ssp126") %>%
#   ggplot(aes(x = log(live_weight_t), y = pct_change)) +
#   geom_point(aes(color = isscaap)) +
#   guides(color = "none") +
#   geom_hline(yintercept = 1, linetype = "dashed", alpha = 0.5) +
#   geom_smooth(method = "lm") +
#   theme_light()
#
# vals <- joined_dataset %>%
#   filter(scenario == "ssp126") %>%
#   split(.$consumer_iso3c) %>%
#   map(~lm(pct_change~log(live_weight_t), data = .x)) %>%
#   map_df(broom::tidy, .id = 'consumer_iso3c') %>%
#   filter(term != "(Intercept)")
#
# vals %>%
#   add_region(col = "consumer_iso3c", region.col.name = "region") %>%
#   ggplot(aes(x = fct_reorder(consumer_iso3c, -estimate), y = estimate, fill = region)) +
#   geom_col() +
#   theme_light() +
#   theme(legend.position = "bottom",
#         axis.text.x = element_blank())
#
#
# # Plot exposure, sensitivity, and adaptive capacity data together
# vals %>%
#   left_join(risk_overall, by = "consumer_iso3c") %>%
#   mutate(pos_or_neg = case_when(estimate > 0 ~ "positive",
#                                 estimate < 0 ~ "negative",
#                                 estimate == 0 ~ "neither")) %>%
#   ggplot(aes(x = sensitivity, y = adaptive_capacity, color = estimate)) +
#   geom_point() +
#   scale_color_gradient2(
#     low = "blue",   # blue for negative
#     mid = "white",     # neutral at 0
#     high = "red",  # red for positive
#     midpoint = 0,
#     name = "Rate of change in\naquatic animal reliance"
#   ) +
#   theme_dark()
```
